### 2017-04-21 2017-Xudong-SMT/submit-ICTAC2017.tex
finished ??

% !TEX TS-program = pdflatex

\documentclass{lncs/llncs}

\usepackage[T1]{fontenc}
%\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{a4paper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}

%\usepackage{amsfonts}
%\usepackage{fancyhdr}
%\usepackage{cite}
%\usepackage{ifthen}
%\usepackage{amssymb}
%\usepackage{fancyhdr}
%\usepackage{pifont}
\usepackage{stmaryrd}
\usepackage{mathtools,mathpartir}
\usepackage{proof}
%\usepackage{setspace}
%\usepackage{indentfirst}
\usepackage{amsmath,amssymb,amscd,mathrsfs}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{epsfig,color,subfigure,enumitem}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand{\NOTE}[1]{\textcolor{blue}{\textbf{[NOTE:#1]}}}
\newcommand{\ERIC}[1]{\textcolor{blue}{#1}}
\newcommand{\OTvar}{\texttt}
\newcommand{\OTland}{\;\land\ }
\definecolor{airforceblue}{rgb}{0.26, 0.44, 0.56}
\newcommand{\QIN}[1]{\textcolor{airforceblue}{#1}}
\newcommand{\coloncolon}{{:\hspace{-.2ex}:}}
\makeatletter
\newcommand{\raisemath}[1]{\mathpalette{\raisem@th{#1}}}
\newcommand{\raisem@th}[3]{\raisebox{#1}{$#2#3$}}
\makeatother

\usepackage{macrospNets}

\def\AlgT{\mathcal{T}}
\def\AlgE{\mathcal{E}}
\def\AlgA{\mathcal{A}}
\def\AlgAS{\mathcal{A}_S}
\def\AlgB{\mathcal{B}}
\def\AlgI{\mathcal{I}}

\newcommand{\Pred}{\symb{Pred}}
\newcommand{\Post}{\symb{Post}}
%\usepackage[math]{cellspace}
%\setlength\cellspacetoplimit{ 37pt}
%\setlength\cellspacebottomlimit{18pt}

\pagestyle{plain}

% addition to the mathpartir package for red dotted rules,
% that we use for open-transitions

\makeatletter
\def \dotover {\textcolor{red}{\leavevmode\cleaders\hb@xt@ .22em{\hss $\cdot$\hss}\hfill\kern\z@}}
\def \reddottedrule #1#2{\hbox {\advance \hsize by -0.5em
%\sbox0{$\genfrac{}{}{0pt}{0}{#1}{#2}$} \phantom{\copy0} %
 {\ooalign{\vphantom{$\genfrac{}{}{0pt}{0}{#1}{#2}$}\cr\dotover\cr$\genfrac{}{}{0pt}{0}{#1}{#2}$\cr}}}}

 \def \dottedrule #1#2 {
  {\sbox0{$\genfrac{}{}{0pt}{0}{#1}{#2}$}%
    \vphantom{\copy0}%
    \ooalign{%
      \hidewidth
      $\vcenter{\moveright\nulldelimiterspace
        \hbox to\wd0{%
         \xleaders\hbox{\kern.5pt\vrule height 0.4pt width 1.5pt\kern.5pt}\hfill
          \kern-1.5pt
        }%
      }$
      \hidewidth\cr
    \box0\cr}}
}

\let \defaultfraction \mpr@@fraction
\makeatother

%\newtheorem{theorem}{Theorem}[section]
\newtheorem{prop}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{alg}[theorem]{Algorithm}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{example}[theorem]{Example}
%\newtheorem{problem}[theorem]{Problem}
%\newtheorem{proof}[theorem]{Proof}

% Macros for the SOS rules and proof trees:
%\newcommand\openrule[2]{\redinfer{#1}{#2}}
\newcommand\openrule[2]{\inferrule*[myfraction=\reddottedrule,center]{#1}{#2}}
%\newcommand\openrule[2]{\inferrule*{#1}{#2}}
%\newcommand\ostate[1]{\triangleleft{\;#1\;}\triangleright}
\newcommand\ostate[1]{\triangleleft{#1}\triangleright}
\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
\usepackage{algorithm} 
%\usepackage{algorithmicx} 
\usepackage{algpseudocode}  
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\algnewcommand{\IfThenElse}[3]{\State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}% \IfThenElse{<if>}{<then>}{<else>}

\title{Using SMT engine to generate Symbolic Automata\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{ Xudong Qin\inst{2,3}  \ \ \  Eric Madelaine\inst{1,2}
  \ \ \  Min Zhang\inst{3}} 
\institute{Univ. of Nice Sophia Antipolis, CNRS, UMR 7271, 06900 Sophia Antipolis, France
	\and INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France
\and Shanghai Key Laboratory of Trustworthy Computing, ECNU, China}
\date{}                                           % Activate to display a given date or no date

                          % Activate to display a given date or no date

\begin{document}

\maketitle

%\section{}
%\subsection{}

\begin{abstract}
  %% \TODO{not good for a conference, (to be rewriten by Eric)}
  %% We implement the symbolic semantics of open pNets using their
  %% so-called ``Open Automaton'' behavioral semantics. This involve
  %% building predicates expressing the synchronization conditions
  %% allowing some combination of events in the pNet system. These
  %% predicates are typically built using first order logic, plus some
  %% predicates specific of particular action algebras. To reduce the
  %% complexity of the generated open automata, we use the Z3 SMT engine
  %% to check satisfiability of the predicates, and prune the state
  %% space..

  Open pNets are used to model the behavior of open
  synchronous or asynchronous systems expressed in various calculi
  or languages. They are endowed with a symbolic operational semantics
  in terms of so-called ``Open Automata''. We implement an algorithm
  computing this semantics, and building predicates expressing the
  synchronization conditions allowing some combination of events in
  the pNet system. Checking such predicates requires symbolic
  reasoning over first order logics, and specific functions of each
  action algebra. To reduce the complexity of the generated open
  automata, we use the Z3 SMT engine to check satisfiability of the
  predicates, and minimize the number of significant symbolic
  transitions.

\end{abstract}

\section{Introduction}

In the nineties, several 
works extended the basic behavioral models based on labelled
transition systems to address value-passing or parameterized systems, using
various symbolic encodings of the
transitions~\cite{deSimone85,Larsen87,HennessyLin:TCS95,Linconcur96}. 
In \cite{Linconcur96}, H.M. Lin addressed value-passing calculi, for which he
developed a symbolic behavioral semantics, and proved algebraic properties.
Separately J. Rathke~\cite{HennessyRathke:TCS98} defined another
symbolic semantics for 
a parameterized broadcast calculus, together with strong and weak bisimulation
equivalences, and developed a symbolic model-checker based on a tableau
method for these processes. Thirty years later, no
practical verification approach and no verification platform are
using this kind of approaches to provide proof methods for
value-passing processes or open process expressions.

Parameterized Networks of Synchronized Automata (pNets) were proposed
to give a behavioral specification formalism for distributed
systems, synchronous, asynchronous, or heterogeneous. It is used in
VerCors, a platform for designing and 
verifying distributed systems, as the intermediate language for various
high-level languages. The high-level languages in VerCors formalize
each component of the 
distributed system and gives out the composition of these
components.
pNets provides the core low-level semantic formalism for VerCors, and
is made of a hierarchical composition of (value-passing) automata,
called parameterized labelled transition systems (pLTS), where each
hierarchical level defines the possible synchronization of the lower levels.
Traditionally, pNets have been used to formalize fully
defined systems or softwares. But we want also to define and reason
about incompletely defined systems, like program skeletons, operators,
or open expressions of process calculi.
The open pNet is proposed to solve the
problem of the undefined components contained in the systems using
"holes" as the process parameter dealing with these components. The
hole acts as the placeholder for the uncertain component with a set of
possible behaviors the component might conduct, presenting the "open"
property.  
The pNet model was developed in a series of
papers~\cite{HMZ:PDP15,henrio:Forte2016} and some examples have been
introduced showing its ability to encode the operators from some
other algebras or  program skeletons. The operational semantic of the open pNet has
already been defined, but we only had a sketch about how to compute it and
had no implementation. This operational semantic is defined as an
Open Automaton in which Open Transitions contain logical predicates
expressing the relations between the behavior of the holes, and the
global behavior of the system.

Computing the open transitions from the composition of pNet is the
main part of the algorithm and needs to satisfy the constraints expressed as
the rules of the operational semantics.
Computing open automata by hand is complicated and error prone. And
the algorithm published in \cite{henrio:Forte2016} was not precise
enough for an implementation. Our first goal was then to refine the
algorithm, and implement it in a tool.
In the long goal, we want to be able to check the equivalence between
pNets. The equivalence between pNets is "FH-bisimulation" taking the
predicate of the open transitions into account each time matching such
open transitions. In our long-term goals, the final target is to check
the property of open systems.

This paper presents our open automaton computing algorithm and its
implementation using the pNet API. Comparing the result of the algorithm with
the manual result on our running examples shows that the algorithm
works properly in computing the pNet semantics. The
algorithm contains the method computing the open transitions composing
the transitions from the subnets according to several defined rules
the open transitions need to satisfy. These results are not the final
results in fact, because they contain some unsatisfiable results we need
to eliminate. To eliminate those unsatisfiability, we interact with an
automatic theorem prover for satisfiability checking.

\paragraph{Contribution}
In the article we show how:
\begin{itemize}
\item We have refined the open automaton generation algorithm, to make
  it implementable, and we implemented a full working prototype, within the
  VerCors platform. In the process, we had to improve significantly the
  semantics rules, and add features in the algorithm to deal the full
  model, including management of variables and assignments.
%  \item We extended the meta-model of action expressions, and the
%    associated API, in the VerCors platform, to be able to manage
%    parameterisation over the action algebra operators.
  \item We have implemented the interaction between our algorithm and the Z3
    SMT solver, for checking satisfiability of the logical predicates
    generated by the algorithm.
    \item We validated the whole approach on a small use-case covering all
      important features of the model.
\end{itemize}

\paragraph{Related works}

%% \TODO{say that apart from the old work by J. Rathke, we know no other
%%   research addressing our goals, and especially no attempts to develop
%%   an algorithmic treatment of such symbolic systems by interacting
%%   with automatic theorem provers}.

%% \TODO{Then you can list the ATP/ITP you have mentioned Coq, Isabelle,
%%   B-tools, or others like CVC4, as possible alternative to Z3}

There is not much researches trying to develop symbolic bisimulation approaches
for the value-passing process algebra and languages, as our long-term goals,
especially no algorithm treatment of the symbolic systems developed by
interacting with automatic theorem provers. The closest work is the
one already mentioned from J. Rathke~\cite{HennessyRathke:TCS98},
who developed the symbolic bisimulation for a
calculus of broadcasting system (CBS). CBS is similar with classic
process calculi such as CCS and CSP, but communicating by broadcasting
values, one-to-many communication instead of one-to-one communication
and transmitting values without blocking. That makes the definition of
the symbolic semantic and bisimulation equivalence different from the
classic works.

For other applications, e.g. for the analyses of 
programming languages, there exist dedicated platforms making use of
external automatic theorem 
provers (ATP), or even some automatic tactics from interactive theorem
provers (ITP), to perform symbolic reasoning, and for example to
discharge some subgoals in the proofs.
Tools like Rodin~\cite{deharbe2013,deharbe2014,abrial2007} have
already integrated several provers, like Z3, as modules for proving
the proof obligations generated from the model inputed by user. 
The prover we used, also is Z3, is developed by Microsoft Research
based on the satisfiability modulo 
theories (SMT), mainly applied in extended static checking, test case
generation, and predicate abstraction.
In a similar way, there are several ATPs/ITPs we could consider to use for
the result pruning and bisimulation checking in our algorithm, as an
alternative to Z3, such as CVC4~\cite{barrett:CAV2011},
Coq~\cite{armand:CPP2011},
%Coq~\cite{armand:CPP2011,besson:CPP2011},
Isabelle~\cite{blanchette:FroCoS2011} or others.

\paragraph{Structure.}
In section
\ref{section:pnets} we give a description and a formal definition of
the pNet model, as found in previous publications. Then in
\ref{section:examples} we present a simple example that will
illustrate the rest of the paper.
Section \ref{section:op-semantics} recalls the operational semantics
of pNet, including its structural rules, and a sketch of the algorithm
to compute such semantics, but also explains the enhancement we have
done to these definitions.
Section \ref{section:implementation} explains in details our
implementation within the Vercors platform, and shows the full result of
the semantic computation on the running example.
Finally we conclude and discuss perspectives in section
\ref{section:conclusion}.

\section{Background: pNets definition}
\label{section:pnets}

% \TODO{I have removed some of the technical stuff, adding referemces to Forte'16.
%  We will see if this is enough, and will have to check if this is coherent, and if we can remove more, but for ICTAC, we MUST keep enough formal stuff}

This section introduces pNets and the notations we will use in
this paper. Then it gives the formal definition of pNet structures,
together with an operational semantics for open pNets.

pNets are tree-like structures, where the leaves are either
\emph{parameterized labelled transition systems (pLTSs)}, expressing the
behavior of basic processes, or \emph{holes}, used as placeholders
for unknown processes, of which we only specify their set of possible
actions, named \emph{sort}.
Nodes of the tree (pNet nodes) are synchronizing artifacts, using a
set of \emph{synchronization vectors} that express the possible
synchronization between the parameterized actions of a subset of the
sub-trees.

%\smallskip\noindent
\paragraph*{Notations.}
We extensively use indexed structures
over some countable indexed sets, which are equivalent to mappings over
the countable set. % . The indexes will usually be
% integers, bounded or not. Such an indexed family is
%denoted
%follows:
$a_i^{i\in I}$
%, or equivalently  $(i\mapsto a_i)^{i\in I}$
denotes a family of elements $a_i$ indexed over the
set $I$. % Such a family
% is equivalent to the mapping $(i\mapsto a_i)^{i\in I}$.
% To specify the set over which the structure is indexed,
% indexed structures are always denoted with an exponent of the form $i\in I$
% (arithmetic only appears in the indexes if necessary).
$a_i^{i\in I}$ defines both $I$ the set over which the family is
indexed (called \emph{range}), and $a_i$ the elements of the family.
E.g., $a^{i\in\{3\}}$ is the mapping with a single entry $a$ at index
$3$ ; abbreviated $(3\mapsto a)$ in the following.
When this is not
ambiguous, we shall use notations for sets, and typically write
``indexed set over I'' when formally we should speak of multisets, and
write $x\in a_i^{i\in I}$ to mean $\exists i\in I.\, x=a_i$.  An empty
family is denoted $\emptyset$. We
denote classically $\overline{a}$ a family when the indexing set is
not meaningful.  $\uplus$ is the disjoint union on
indexed sets.

\paragraph*{Term algebra.}
Our models rely on a notion of parameterized actions, that are
symbolic expressions using data types and variables. As our model aims
at encoding the low-level behavior of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on unification between two or more action
expressions. This will be used in the semantics of synchronization
vectors to express various kinds of communication or synchronization mechanisms.

\def\Talg{\mathcal{T}_{\Sigma,\P}}
Formally, we assume the existence of a term algebra $\Talg$,
where $\Sigma$ is the signature of the data and action constructors,
and $\P$ a set of variables. Within $\Talg$, we distinguish a set of
data expressions $\mathcal{E}_\P$, including a set of boolean
expressions $\mathcal{B}_{\P}$ ($\mathcal{B}_{\P}\subseteq\mathcal{E}_\P$).
On top of $\mathcal{E}_\P$ we build the action algebra
$\mathcal{A}_\P$, with $\mathcal{A}_P\subseteq\mathcal{T}_\P,
\mathcal{E}_P\cap\mathcal{A}_P=\emptyset$;
naturally action terms will use data expressions as subterms.
To be able to reason about the data flow between pLTSs, we
distinguish \emph{input variables} of the form $?x$ within terms; the function
$\vars(t)$ identifies the set of variables in a term
$t\in\AlgT$, and $iv(t)$ returns its input variables.

pNets can encode naturally the notion of input actions as found e.g. in value-passing CCS
\cite{Milner89} or of usual point-to-point message passing calculi, but it also allows
for more general mechanisms, like gate negotiation in Lotos, or broadcast
communications.
%\TODO{I removed the CCS input action stuff, check later if we need to
%  replace it here by something different about input action notation}.

%% Formally, value-passing actions \emph{\`a la} CCS, denoted as
%% ``?ch(x)'' and ``!ch(exp)'' in our CCS running example, stand
%% respectively for $CCSact(i(ch),?x)$ and $CCSact(ch,exp)$, where $ch$
%% is a variable denoting a channel name, $?x$ is an input variable
%% denoting the data argument received as input on $ch$, and $exp$ some
%% data expression (containing no input variables). In our example we
%% will use the usual CCS syntax. 
%% %For Lotos-style
%% %distributed synchronization, we use synchronization vectors that encode matching data
%% %offers on a common gate; or 
%% We can also use more complex action structure such as Meije-SCCS action
%% monoids, like in $a.b$, $a^{f(n)}$ (see \cite{deSimone85}). The expressiveness of the synchronization constructs
%% will depend on the action algebra.

\subsection{The (open) pNets Core Model}
\label{section:pNets}

A pLTS is a labelled transition system with variables; variables can be
manipulated, defined, or accessed inside states, actions, guards, and
assignments. Without loss of generality and to simplify the formalization, we suppose 
here that variables are local to 
state: each state has its set of variables disjoint from the others. Transmitting 
variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
Note that we make no assumption on finiteness of the set of states nor
on finite branching of the transition relation.

We first define the set of actions a pLTS can use, let $a$
range over action labels, $\symb{op}$ are operators, and $x_i$ range over
variable names. Action terms are:
\[
\begin{array}[l]{rcl@{\quad}p{5.5cm}}
  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
  p_i&::=& ?x~|~\symb{Expr}&\text{parameters (input variable or expression)}\\
  \symb{Expr}&::=& \symb{Value}~|~x~|~\symb{op}(\symb{Expr}_1,..,\symb{Expr}_n)&\text{Expressions}
\end{array}
\]
The input variables in an action term are those marked with a
$\symb{?}$.
We additionally suppose that each input variable does not
appear somewhere else in the same action term:
$p_i=?x\Rightarrow\forall j\neq i.\, x\notin \vars(p_j)$

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of states.
\item[$\bullet$]
$s_0 \in S$ is the initial state.
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
where $\alpha \in\AlgA$ is a parameterized action, $e_b \in
\AlgB$ is a guard, and the variables $x_j\in P$
are assigned the expressions $e_j\in \AlgE$.
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then $\iv(\alpha)\!\subseteq\! \vars(s')$, 
		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! \vars(s)$, 
		$\vars(e_b)\!\subseteq\! \vars(s')$, and
		$\forall j\!\in\! J .\,\vars(e_j)\!\subseteq\! \vars(s)\land 
		x_j\!\in\!\vars(s')$. %,  and $s= s'\Rightarrow J=\emptyset$. 
\item[$\bullet$]
The $x_j$ here are $state\ variables$ of state $s'$. $State\ variables$ can only be modified by the assignment in the transitions targetting its owner state. A global state variable of a pLTS is a variable $x$ satisfying $\forall s\!\in\! S .\, x\!\in\! \vars(s)$.

\end{itemize}
\end{definition}

Now we define
pNet nodes, as constructors for hierarchical behavioral structures.
A pNet node has a set of sub-pNets that can be either pNets or pLTSs, and a
set of Holes, playing the role of process parameters.

A composite pNet consists of a set of sub-pNets exposing
a set of actions, each of them triggering internal actions in each of
the sub-pNets. The synchronization between global actions and
internal actions is given by  \emph{synchronization vectors}: a
synchronization vector synchronizes one or several internal actions, and
exposes a single resulting global action.
Actions involved at the pNet level (in the synchronization vectors) do
not need to distinguish between input and output
variables. Action terms for pNets are defined as follows:
\[\begin{array}[l]{rcl@{\quad}l}
  \alpha\in \AlgAS &::=&a(Expr_1,\ldots,Expr_n)
\end{array}
\]

\begin{definition}[pNets]\label{def-pnets}
A pNet is a hierarchical structure where leaves are pLTSs and holes:\\
$\pNet\triangleq pLTS~|~\mylangle \pNet_i^{i\in I}, S_j^{j\in J}, \symb{SV}_k^{k\in K}\myrangle$
where
\begin{itemize}
\item[$\bullet$] $I \in \I$ is the set over which sub-pNets are indexed.
\item[$\bullet$] $\pNet_i^{i\in I}$ is the family of sub-pNets.
%  $\pNet_i^{i\in I}$ is a family of sub-pNets where $I\in\I_\P$ is the set over which sub-pNets are indexed.

\item[$\bullet$] $J\in\I_\P$ is the set over which holes are indexed.
$I$ and $J$ are \emph{disjoint}: $I\cap J=\emptyset$,  $I\cup J\neq\emptyset$
\item[$\bullet$] $S_j \subseteq \AlgAS$ is a set of action terms, denoting the $\Sort$ of
hole $j$.

\item[$\bullet$] $\symb{SV}_k^{k\in K}$ is a set of
  synchronization vectors ($K\in\I_\P$). $\forall k\!\in\! K,
  \symb{SV}_k\!=\!\alpha_{l}^{l\in I_k \uplus J_k}\to\alpha'_k$ where
  $\alpha'_k\in \mathcal{A}_\P$, $I_k\subseteq I$, $J_k\subseteq J$,
  $\forall i\!\in\!
  I_k.\,\alpha_{i}\!\in\!\Sort(\pNet_i)$,  $\forall j\!\in\!
  J_k.\,\alpha_{j}\!\in\!S_j$, and $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\Label(\symb{SV}_k) = \alpha'_k$.

\end{itemize}
\end{definition}

By lack of space, we have omitted here the formal (structural)
definition of $Sorts$ (the 
set of actions of a hole or a pNet), $Leaves$ and $Holes$ (all pLTSs
(resp holes) in a pNet hierarchical system. These can be found in \cite{henrio:Forte2016}.
%The preceding definition relies on the auxiliary functions below:

%\begin{definition}[Sorts, Holes, Leaves of pNets]
%  \begin{itemize}
%  \item The sort of a pNet is its signature, i.e. the set of actions it can
%perform. In the definition of sorts, we do not need to distinguish
%input variables (that specify the dataflow within LTSs), so for
%computing LTS sorts, we use a substitution operator\footnote{$\subst{y_k\gets x_k}^{k\in K}$ is the parallel substitution 
%operation.} to remove the
%\emph{input marker} of variables. Formally:
%\[
%\begin{array}{l}
%\Sort(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{x \gets ?x| 
%x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
%		J}\rangle} s'\in \to \} \\ \Sort(\mylangle
%\set{\pNet}, \set{\symb{S}}, \set{\symb{SV}}\myrangle) =\{\alpha'_k |\, \alpha_j^{j\in J_k}\to\alpha'_k\in\set{\symb{SV}}\}
%\end{array}
%\]
%
%\item
%The set of holes of a pNet is defined inductively; the sets of holes
%in a pNet node and its subnets are all disjoint:
%  \[\begin{array}{l}
%\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
%\Holes(\mylangle \pNet_i^{i\in I}\!,S_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) =J\cup{\displaystyle \bigcup_{i\in I}\Holes(\pNet_i)}\\
%\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
%\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
%\end{array}\]
%\item
%The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined inductively as:
%\[\begin{array}{l}
%\Leaves(\mylangle S,s_0, \to\myrangle) \!=\! \{ \mylangle S,s_0, \to\myrangle \}\\
%\Leaves(\mylangle \pNet_i^{i\in I}\!,S_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \bigcup_{i\in I}\Leaves(\pNet_i)}
%\end{array}\]
%\end{itemize}
%\end{definition}
%
%A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
%is said to be \emph{open}.

%% \begin{figure}[t]
%% \begin{minipage}{6.1cm}

%%   \includegraphics[width=\linewidth]{XFIG/Enable1}
%%   \\[1.3ex]
%%  \includegraphics[width=\linewidth]{XFIG/Enable2}
%%   \caption{Two pNet encodings for  Enable }  \label{schema:enable-pnets}
%% \end{minipage}
%%   \hspace{2mm}
%% \begin{minipage}{6cm}
%%   \includegraphics[width=\linewidth]{XFIG/P-QR}
%%   \caption{Composed pNet for ``P>>(Q>>R)''}  \label{schema:enable-composed}
%% \end{minipage}

%% \end{figure}

\section{Running Example}
\label{section:examples}

Several publications \cite{HMZ:PDP15,henrio:Forte2016} already have introduced
many examples of pNets, encoding
operators of various classical process algebras, or more complex
synchronization structures in distributed or parallel languages.
In this paper, we use an example inspired from Lotos, to illustrate
the main features of pNets and the way we deal with them in the symbolic
semantics, and in the satisfiability checking.

\begin{figure}[h]
  \centerline{\includegraphics[width=10cm]{XFIG/LOTOSFormula.eps}}
  \caption{pNet encodings for the LOTOS expression}  \label{schema:lotos-pnet}
\end{figure}

We choose the expression \texttt{P>>(acc(v);Q)}, that includes the
Enable and Prefix 
operators of the LOTOS language. For the readers not familiar with
Lotos, the pNet encodings here should be sufficient to understand the
Enable operator semantics.
In Fig.~\ref{schema:lotos-pnet}, we show the encoding of the formula using pNets.

  This pNet is composed of: a toplevel node PN1 encoding the Enable
  operator, with its controller (pLTS) C1; a hole representing the
  process variable P; another node PN2 representing the Prefix
  operator, with its controller C2, and its hole Q.

In order to show all features of the pNet model, we choose to use two
styles of encodings in the controllers.
The most common style used in process calculi is ``state oriented'',
we use it here for C2: it has two states to
indicate the position of the control point in the
Prefix expression $(acc(v);Q)$. The second one is
``data oriented'': controller C1 has a single state, with a state
variable $s_0$ encoding the change of control point from the left to
the right sub-net.
%% The transitions in the pLTS have the more general syntax:
%% ``<action-expr> [guard] {sequence of assignments}''. Guards and
%% assignments can only use the values of variables of their source state
%% and values from input variables in the action expression; and a
%% transition can only assign variables of its target state.

Now look more specifically on the top pNet node PN1. In addition to
its controller it has 2 sub-pNets: the hole P and the sub-pNet PN2.
Three synchronization vectors are defining the possible
synchronizations between the actions of the subnets.
 The first
vector, transmitting an action $a_1$ of P, can only occur if $a_1$ is not
a $\delta$, the vector concrete syntax is typically ``$< l, a_1 , \_ >
-> a_1$'', in which ``$\_$'' means that the corresponding subnet is
not involved in this synchronization, and the ``l'' action is related
to the corresponding transition of the controller C1: this action is
available only when the state variable $s_0$ is equal to 0.
The second vector composes the
actions from the subnets and gives out a synchronized action
$\underline{\delta(x)}$, where the underlined action stands for an
internal (already synchronized) action, that cannot be further
synchronized at upper levels. Synchronized action is a straightforward
generalization of the notion of internal actions, that will be
convenient for observing internal events during model-checking. Here
the transition occurs only when process
$P$ performs a $\delta(x)$ action, transferring the control to $Q$,
and sending a value $x$ at the same time. This value must be
``accepted''  through an $acc(x)$ action before $Q$ takes the
control. The last vector acts as the first one, but transmitting the
action $a_2$ of Q.  
Similarly the node PN2 has two synchronization vectors to encode
the behavior of the Prefix operator.

It should be easy to see that this pNet system encodes properly the
behavioral semantics of the corresponding Lotos expression. Remark
that it is built in a systematic structural way, one pNet node
encoding each Lotos operator. It is parameterized both at the level of
actions variables (occurring within the controllers and the
synchronization vectors), and at the level of process parameters
(holes).

\section{Operational Semantics for Open pNets}
\label{section:op-semantics}

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications.

\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a set  $(S_i,s_{0 i}, \rightarrow_i)^{i\in
	I}$ of LTSs, a
	set $J$ of holes with sorts $Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
	{s \xrightarrow {v}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and for all
        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and $\xrightarrow{b_j}_j$
        is a transition of the hole $j$, for any action $b_j$ in the
        sort $Sort_j$. \Pred\ is a predicate
	over the different variables of the
	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $\{x_k\gets e_k\}^{k\in K}$ 
	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
	variables of the open transition.
\end{definition}

\begin{definition}[Open automaton]
	\label{def:open-automaton}
	An \emph{open automaton} is a structure\\ $A =
	<LTS_i^{i\in I},J,\mathcal{S},s_0,\mathcal{T}>$ where:
	\begin{itemize}
		\item[$\bullet$]  $I$ and $J$ are  sets of indices,
		\item[$\bullet$]  $LTS_i^{i\in I}$ is a family of LTSs,
		\item[$\bullet$]   $\mathcal{S}$ is a set of states and $s_0$ an initial state
		among $\mathcal{S}$,
		\item[$\bullet$] $\mathcal{T}$ is a set of open transitions and for each
		$t\in \mathcal{T}$ there exist $I'$, $J'$ with $I'\subseteq I$, $J'
		\subseteq J$, such that $t$ is an open transition over $LTS_i^{i\in I'}$, $J'$,
		and  $\mathcal{S}$.

	\end{itemize}
\end{definition}

%
%Then the semantics of a pNet is characterized by a set of {\em open
%transitions}, where the hypotheses on process parameters are
%replaced by 1) transitions of the pLTSs at the leaves, and 2) formal
%hypotheses on the transitions of the holes. A {\em predicate} is used
%to relate the parameters and names appearing in the actions of the
%leaves and the holes involved in the rules, but also appearing in  the resulting action.

When building an Open Automaton as the semantics of a pNet, its
\emph{states}, and the shape of the predicates in its
\emph{transitions} have a specific structure:

\paragraph{States of open pNets:}\label{def-states}
  A state of an open pNet is a tuple (not necessarily finite) of the
  states of its leaves (in which we denote tuples
  in structured states as $\triangleleft\ldots\triangleright$ for better readability).

  For any pNet p, let $\overline{Leaves} = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be the set of pLTS at its leaves,
  then $States(p) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
A pLTS being its own single leave:
  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S\}$.

The initial state is defined as:
$InitState(p) = \triangleleft {{s_i}_0}^{i\in L}  \triangleright$.

%% \begin{example} \emph{State of a pNet}
%%   The states of pNet \texttt{EnableCompL} are:
%%   $\triangleleft 00 \triangleright, \triangleleft 10 \triangleright, \triangleleft 11 \triangleright$
%% \end{example}

\paragraph{Predicates:}
Let
$\mylangle\overline{\pNet},\overline{S},\symb{SV}_k^{k\in K} \myrangle$
be a pNet. Consider a synchronization vector $SV_k$, for $k\in K$. We define a
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\[\Pred(SV_k, a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow
\begin{array}{l}
\exists {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.\, SV_k={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v'
\\~~\land
\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v'
\end{array}\]
In any other case (if the action families do not match or if there is no valuation of
variables such that the above formula can be ensured) the predicate is undefined.

This definition is not constructive but it is easy to build the predicate constructively
by brute-force unification of the sub-pNets
actions with the corresponding vector actions, possibly followed by a simplification
step.

%\QIN{[ change the example OT here ]}
\begin{example}\emph{An open-transition.}
  \label{OT:lotos-composed}

The \texttt{LOTOS} pNet of Fig. \ref{schema:lotos-pnet} has 2
controllers and 2 holes. We will show its full open Automaton in
section \ref{section:full-result}. We give one of its transition here:

 \smallskip
 $ ot_2  = \openrule{
                       \{0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1\}, ~~
                       \{\xrightarrow{hb_{12}}_P\}, ~~ 
                        [s_0=0 \!\land\! hb_{12}=\delta(x)], ~~ 
                       \{s_0 \xleftarrow{} 1\}     
                      }
     {\ostate{00} \xrightarrow{\underline{\delta(x)}} \ostate{01}}$
\end{example}

The open transition $ot_2$ involves the possible actions from the process $P$ and prefix expression $acc(x);Q$. The hole behavior, represented by variable $hb_{12}$ must be equal to the exit action $\delta(x)$, and the accept event $acc(x)$ will capture the same value $x$.

The transition in the conclusion of this rule starts in
state $\ostate{00}$, where both $C_1$ and $C_2$ are in their initial
states. It ends in a different structured state, because in $ot_2$
both pLTS conduct transitions, showing that the control has gone from
processes $P$ to $(acc(x);Q$, consuming actions $\delta(x)$ and
$acc(x)$. Controller $C_1$ has only one state, but its state variable
$s_0$ has changed to 1.

\paragraph{Structural Semantic Rules:}
Now we build the semantics of an open pNet as an open automaton where
LTSs are the pLTSs at 
the leaves of the pNet structure, and the states are structured as in
the previous section. 
%given by Definition~\ref{def-states}.
The open transitions first
 project the global state into states of the leaves, then apply
pLTS transitions on these states, and compose them with the sort of the holes. %The pNet
%structure does not appear in the open-automaton, only the
%set of Holes and the set of Leaves.
The semantics   regularly instantiates \emph{fresh} variables, and uses a
\emph{clone} operator that clones a term replacing each variable with a
fresh one.

The semantic of the open pNets is already defined as an open automaton in \cite{henrio:Forte2016} using open transitions to present the transitions of its global states.
\begin{definition}[Operational semantics of open pNets]
  \label{def:operationalSemantics}
  The semantics of a pNet $p$ is an open automaton $A = <Leaves(p),J,\mathcal{S}, s_0,
  \mathcal{T}>$ where:
  \begin{itemize}
    \item $J$ is the indices of the holes: $Holes(p)= H_j^{j\in J}$.
    %  \item $\overline{L}^L = Leaves(p), \overline{H}^J = Holes(p)$
    \item $\overline{\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
    \item $\mathcal{T}$ is the smallest set of open transitions
    satisfying the rules below:
  \end{itemize}

  The rule for a pLTS $p$  checks that the guard 
  is verified and transforms assignments into post-conditions:

  \begin{description}
    \item[{\bf Tr1:}]
    $\inferrule
    { s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
          J}\rangle} s'\in \to  \\
      {\tt fresh}(v) \\
      Pred = ~e_b \land (\alpha = v)
    }
    { p = \mylangle  S,s_0, \to \myrangle
      \models
      \inferrule*[myfraction=\reddottedrule]
      {\{s \xrightarrow{\alpha}_p s'\} ,\emptyset ,
      Pred,\left\{x_j\gets e_j\right\}^{j\in J}}
      {\ostate{s} \xrightarrow{v} \ostate{s'}}
    }
    $
  \end{description}
% Note that this note is greatly simplified by the fact that variables are local to 
% thread; introducing global state variables or accepting loops to the same 
% state would 
% require to reason 
% on the scope of 
% each variables, and to introduce additional variables to handle the several occurence 
% of the same pLTS variable in the predicates. Indeed the constraints on pLTS 
% transitions 
% ensure that the same variable never appears both on the left and on the right of the 
% equations of a predicate.

  The second rule deals with pNet nodes: for each possible
  synchronization vector applicable to the rule subject, the premisses
  include one {\em open transition} for each sub-pNet involved, one possible
  {\em action} for each Hole involved, and the predicate relating these
  with the resulting action of the vector.
  A key to understand this rule is that the open transitions are
  expressed in terms of the leaves and holes of the pNet structure,
  i.e. a flatten view of the pNet: e.g. $L$ is the index set of the
  Leaves, $L_k$ the index set of the leaves of one subnet, so all $L_k$
  are disjoint subsets of $L$. Thus the states in the open transitions,
  at each level, are tuples including states of all the
  leaves of the pNet, not only those involved in the chosen
  synchronization vector.

  \begin{description}
    \item[{\bf Tr2:}]
  \end{description}

  \noindent
    $\inferrule
    {k\!\in\! K \\ SV\!\!= \!clone(SV_k) \!=\! \alpha_m^{m \in I_k\uplus J_k} \!\to\! 
    \alpha'_k, G_k \\
      Leaves(p) \!=\! \pLTS_l^{l\in L} \\     
      \forall m\in I_k.
      \pNet_m \models
      \inferrule*[myfraction=\reddottedrule]
      {\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
      \{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
      {\ostate{s_{i}^{i \in L_m}} \xrightarrow {v_m}
        \ostate{s_{i}^{\prime\ i \in L_m}}}
      %\land
      %Leaves(\pNet_m) = \overline{pLTS}^{L_k})
      \\
      I' = \biguplus_{m\in I_k}\!\! I_m'
      \\ J' = \biguplus_{m\in I_k}\!\! J'_m \uplus J_k  \\
      \Pred = \bigwedge_{m\in I_k}\!\! \Pred_m \land
      \Pred(SV,v_m^{m\in I_k},b_j^{j\in J_k},v)\\
      \forall j\!\!\in\!\! J_k. {\tt 
        fresh}(b_j) \\ {\tt fresh}(v) \\ 
      \forall i\in
      L\backslash I'.\,s'_i=s_i 
    }
    {p = \mylangle \pNet_i^{i\in I}, S_j^{j\in J}, \symb{SV}_k^{k\in K}\myrangle
      \models
      {\inferrule*[myfraction=\reddottedrule]
        {\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
        \{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
        \Post_m}
        {\ostate{s_i^{i\in L}} \xrightarrow {v}
          \ostate{s_i^{\prime i\in L}}}
      }
    }
    $

  \medskip  
\end{definition}

%
%\begin{figure}[tb]
%\begin{mathpar}
%  \small
%  \inferrule
%    {\inferrule
%      {
%          \inferrule
%              {1 \xrightarrow {l}_{C_1} 1}
%              {C_1
%                \models
%                \openrule
%                      {1 \xrightarrow{l}_{C_1} 1}
%                      {\ostate{1}\xrightarrow{1}\ostate{1}}
%              }
%        }
%        {\textrm{P}
%          \models
%            \openrule{
%              1 \xrightarrow {l}_{C_1} 1,\,
%              \{\xrightarrow{b_P}_P\},\, b_P\neq l}
%                      {\ostate{1}\xrightarrow{b_P}\ostate{1}}
%        }\\
%%      \sm{fresh}v         \\
%      \inferrule%*[right={L_1}]
%        {
%          \inferrule
%              {0 \xrightarrow{b}_{C_2} 1}
%              {C_4
%                \models
%                \openrule
%                      {0 \xrightarrow{b}_{C_2} 0,\, b\neq r}
%                      {\ostate{0}\xrightarrow{b}\ostate{1}}
%              }
%        }
%        {
%          \textrm{b:Q}\models
%              \openrule
%                  { 0 \xrightarrow {b}_{C_2} 1,\, \ b\neq r}
%                  {\ostate{0}\xrightarrow{b}\ostate{1}}
%        }   
%    }
%    {
%     \textrm{PN1}
%     \models
%        \openrule{
%                       1 \xrightarrow{l}_{C_1} 1 ~~
%                       0 \xrightarrow{b}_{C_2} 1 ~~
%                       \{ \xrightarrow{b_P}_{P} \}               ~~
%                 [ a\neq l \land b\neq r \land b_P=?ch(v) \land b=!ch(v) ]              
%                      }
%     {\ostate{10} \xrightarrow{\tau} \ostate{11}}
%      }\vspace{-4ex}
%\end{mathpar}
%  \caption{Proof of  $OT_{11}$ (with interaction of processes $P$ and action $b$) for ``a.P||b.Q''}
%  \label{usingrules:OT11}
%\end{figure}

\begin{example} \emph{Using the operational rules to compute  open-transitions: }
  in Fig. \ref{usingrules:ot2} we show the deduction tree used to construct and prove the 
  open transition $ot_2$.
\end{example}

\begin{figure}[h]
  \vspace{-10mm}

\begin{mathpar}
  \small
  \inferrule
    {
          \inferrule
              {0 \xrightarrow {\delta}_{C_1} 0 ~~
               [ s_0=0 ]
              }
              {C_1
                \models
                \openrule
                      {0 \xrightarrow{\delta}_{C_1} 0 ~~
                       [ s_0=0 ] ~~
                       \{ s_0 \leftarrow 1\}
                      }
                      {\ostate{0}\xrightarrow{\delta}\ostate{0}}
              }
	\\
      \inferrule
        {
          \inferrule
              {0 \xrightarrow{acc(x)}_{C_2} 1}
              {C_2
                \models
                \openrule
                      {0 \xrightarrow{acc(x)}_{C_2} 1}
                      {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
              }
        }
        {
          \textrm{acc(x);Q}\models
              \openrule
                  { 0 \xrightarrow {acc(x)}_{C_2} 1}
                  {\ostate{0}\xrightarrow{acc(x)}\ostate{1}}
        }   
    }
    {
     \textrm{PN1}
     \models
        \openrule{
                        0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1 ~~
                       \{\xrightarrow{hb_{12}}_P\} ~~ 
                        [s_0=0 \!\land\! hb_{12}=\delta(x)] ~~ 
                       \{s_0 \xleftarrow{} 1\}           
                      }
     {\ostate{00} \xrightarrow{\tau} \ostate{01}}
      }\vspace{-4ex}
\end{mathpar}
  \caption{Proof of  $ot_{2}$ (interaction of process $P$ and action $acc(x)$) for ``P>>(acc(x);Q)''}
  \label{usingrules:ot2}
\end{figure}

The proof tree uses TR1 twice, for the $\delta$ transition of $C_1$
and for the $acc(x)$ transition of $C_2$, then uses an action
$hb_{12}$ of hole $P$, and combines the results using the first vector
of the PN2 sub-pNet, and the second vector of the top node according
to TR2. This yields a final $\underline{\delta(x)}$ transition. The
deduction tree in the figure shows how the predicates are generated in
this process.

\paragraph{Fresh variables management.}
The variables in each synchronization vector are considered local:
for a given pNet expression, we must have fresh local variables for
each occurrence of a vector (= each time we instantiate rule
Tr2). Similarly the state variables of each copy of a
given pLTS in the system, must be distinct, and those created for each
application of Tr2 have to be fresh and all distinct. 
% This will be implemented within the open-automaton generation algorithm,
% e.g. using name generation using a global counter as a suffix.

\subsection{Presentation of algebra}
One of the important goal of the Open pNet framework is to provide us
with a model able to express the semantics of many different algebras
or languages. In particular, as we have seen in section
\ref{section:pnets}, the \emph{term algebra} used to express the
action expressions and the data in their parameters is left open in
the pNet formalization.

So, for a given language, we now need to specify its data and action
domains, giving them an abstract syntax (sorts, constructors, operators,
and predicates). We call this a \emph{Presentation}.
We also define a concrete syntax, that will be used for
pretty-printing, but also for translating the presentation, and the
predicates, into Z3 syntax (we shall detail this translation in
section \ref{section:pruning}).

An algebra presentation contains:
\begin{itemize}
	\item Sorts: Constants sets of the algebra or types of the data
          and actions. In every presentation, sorts for integer and
          boolean are implicitly declared.  There \emph{must} be one sort for actions, and eventually 
several others for the types of data parameters. In the LOTOS example,
we have the $Action$ sort: $\OTvar{Action = } \{ l, \delta, r, p, q, acc \} .$

	\item Operators: Operators used to construct action
          expressions (including their data parameters); they are
          constructors of data and action sorts, but also predicates
          over these objects. For instance, in our LOTOS example, we
          could have two constructor functions ``$DELTA$'' and ``$ACC$''
          to construct the action expression $\delta(x)$ and $acc(x)$, declared as:
          $\OTvar{DELTA:}\OTvar{< Data >} \rightarrow \OTvar{Action .}$
        \item Constants: The satisfiability solver need to know which
          actions in the pLTS behaviors are constants (like the $l$, $\delta$, 
          $r$ of the controllers in running example). 
          For a given pNet system, a sort $Action$ will include all
          these constants.
\end{itemize}

%We need introduce operators for building expressions, of both action
%and data sorts, when encoding the 
%algebra into pNets. So that is needed to know about the input and
%output of the operators used in the expressions from those
%algebra. The input and output should belong to the sorts of the
%pNet. When there are $k$ sorts in this pNet, the structure of a
%$n$-ary operator is vector as $< Sort_{i_1}, ... , Sort_{i_n} >
%\rightarrow Sort_j , i_1 , ... , i_n, j \in [0..k] $.
%% We  introduce operators for building expressions, either action
%% expressions or predicate terms, when encoding the algebra into
%% pNets. The operators here are treated as functions, with the sorts of
%% the pNet acting as the types of their input and output. 
%% For instance, in our LOTOS running example, we use a function ``$ACT$'' action to construct the action expression $\delta(x)$ and $acc(x)$. We can define the operator as a binary operator with the format like:

%%  \[\OTvar{ACT:}\OTvar{< Action, Int >} \rightarrow \OTvar{Action .} \]

%\paragraph{Implementation of the expression meta-model}
% \TODO{maybe we could remove this paragraph (and the figure)}
%The VerCors platform uses Ecore and EMF to define the structure of
%both pNets and actions algebras. There exists a predefined set of
%action and data expressions classes, including boolean, integer and
%action expressions, with predefined variables, literal (constants),
%and operators. Operators can be categorized into unary, binary and
%N-ary operators.
%The original version of VerCors did not allow to extend these classes,
%we have modified its architecture so that a user can extend them,
%defining more elements for his specific algebra.
%
%In Fig~\ref{schema:expression-pnets}, we show the Expression class we
%improved, including extensions for CCS and LOTOS.
%
%\begin{figure}[t]
%  \includegraphics[width=\linewidth]{XFIG/Expression}
%  \caption{The architecture of Expression class}  \label{schema:expression-pnets}
%\end{figure}

%\TODO{give an example of user-defined sorts/operators, and tell that
%  you will give it's translation into Z3 in section 5.4}

\subsection{Computing and using open automata}
In this section we present an algorithm to construct the open
automaton representing the behavior of an open pNet, and show
under which conditions this automaton is finite.

\begin{alg}[behavioral semantics of open pNets: Sketch]
This is a standard residual algorithm over a set of open-automaton
states, but where transitions are open transitions
constructively ``proven'' by deduction trees.

1) Start with a set of unexplored states containing the initial state
of the automaton, and an empty set of explored states.

2) While there are unexplored states:

2a) pick one state from the unexplored set and add it to the explored
set. From this state
build all possible deduction trees by application of the structural
rules Tr1 and Tr2, using all applicable combinations
of synchronization vectors.

2b) For each deduction trees, extract the resulting
open-transition.
Its predicate is checked (using Z3) for  satisfiability. If it is
unsatisfiable, the resulting OT is discarded. This will minimize the
number of resulting transitions.

2c) For each open-transition,  add the transition in the outgoing transitions of
the current state, and add the
resulting state in the unexplored set if it is not already in the
explored set.

3) Finally a reachability check could potentially
prune the state space.

\end{alg}

To have some practical interest, it is important to know when this
algorithm terminates. The following theorem shows that an open-pNet
with finite synchronization sets, finitely many leaves and
holes, and each pLTS at leaves having a finite number of states and
(symbolic) transitions, has a finite automaton. The proof can be found
in \cite{henrio:Forte2016}.

\begin{theorem}[Finiteness of open-automata.]\\
Given an open pNet $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in K}\myrangle$ with leaves $pLTS_i^{i\in L}$ and holes $Hole_j^{j\in
  J}$, if the sets $L$ and $J$ are finite, if the synchronization vectors of all pNets 
  included in  $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in K}\myrangle$ 
  are finite, and if
$\forall i \in L.\, finite{(states(pLTS_i))} \text{ and } pLTS_i$
has a finite number of state variables, then Algorithm 1 terminates
and produces an open automaton 
$\mathcal{T}$ with finitely many states and transitions.

% Given an open pNet $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in
%     K}\myrangle$ with leaves $pLTS_i^{i\in L}$ and holes $Hole_j^{j\in
%   J}$,
% build its semantics as in algorithm 1.

% We have:

% $$ finite{(L)} \land finite{(J)} \land \forall i \in L finite{(\{s_i\})}
%   \to finite{(\mathcal{T})}$$
\end{theorem}

\section{Implementation}
\label{section:implementation}
The VerCors platform uses (closed) pNets as the intermediate language
for some high-level language or graphical formalism to be translated
into both input for a 
model-checker and for generating executable code automatically \cite{HKM-FASE16}.
% to ease the
% specification from non-expert users, compiling it
%from those language or graphical formalism in the pNet API.

In our prototype we start directly with open pNets expressed
programmatically, using the pNet API (graphical editors may be added
later). From these, we compute
the open automaton of an open pNet. We explain here our main
implementation choices.
%:
%\begin{itemize}
%  \item Steps 2a-2b are merged, applying Tr2 from the
%premisses OTs from subnets to generate OTs without
%explicitely constructing the deduction tree.
%  \item We define a general naming schema for implementing the fresh 
%and clone functions with more structure than a simple global name
%generator. It ensures unicity, but builds names
%readable enough for debugging purpose.
%  \item Besides that, we also have designed a management of state variables and
%    assignments throughout the whole computation.
%%%   \item Satisfiability check is done only at
%%% top level of the open-transition construction. An alternative would be
%%% to submit the satisfiability check to the SMT solver at each level of
%%% the hierarchical OT construction, potentially reducing the overall number of
%%% combinations. But the submission to the SMT engine is costly, and more
%%% complexity analysis is required before deciding if this would be
%%% worthwhile. 
%\end{itemize}
%
%In the following sections, we detail:
%\begin{itemize}
%\item The structure of fresh variable names.
%\item The details of the implementation for step 2a-2b, refined into:
%  \subitem combining all possibilities of subnets activity/inactivity,
%  \subitem matching each combination with synchronization vectors.
%\item Management of local variables of pLTSs.
%\item Pruning unsatisfiable OTs using the Z3 SMT solver.
%\end{itemize}

\subsection{Fresh variables}

%The variables in each synchronization vector are considered local, so
%does the variables of an open transition such as its hole behaviors
%and its result action. So we want rename the variables to make them
%unique on every occurrence of a vector. Similarly, the name given to
%the hole behaviors for each hole or the result action for each open
%transition must be distinct. At the same time, we also hope them still
%readable. Here we rename the variables with a regular format using the fresh function.

Application of the semantic rules require generating a lot of fresh
names, for different kind of variables. We could use a global name
generator to guarantee unicity, but 
at the same time, we also hope them still
readable. Here we rename the variables with a regular format using the fresh function.
More precisely, the fresh function generates a new name adding a
suffix after the original name. The suffix contains three parts
combined by a colon.

\begin{definition}[Fresh variable]\label{fresh-variable}
The format of \emph{fresh variable} (renaming) is defined as:
``$\emph{prefix : tree\ index\ :\ counter}$''.
\begin{itemize}
   \item[$\bullet$] $prefix$ identifies the kind of the
     variable. current kinds are: sva (SV action), ra (result action), hb (hole behavior).
   \item[$\bullet$] $tree\ index$ is the index of the node containing the variable in the tree-like structure of the pNet.
   \item[$\bullet$] $counter$ is the current value of the corresponding counter.
\end{itemize}
\end{definition}

%\TODO{Option: Add the process and the vector ids ?}

%$Prefix$ avoids the confusion between variables from different structures with the same name by attaching the type of the structure. $Tree\ index$ is given to every node of the pNets to mention which node the variable belongs to as pNets has a tree-like structure. To each node, the tree index is always a number sequence of the tree index of its higher level and the index of the node in this level at the last. $A\ set\ of\ counters$ is used to count the current times the SV, Hole, subnet invoked, or the number of possible OT generated to provide an identity.

For example, in the running example, the first possible behavior of
the hole, it doesn't have a name, only have the prefix ``hb''. The
hole is the second subnets of the root node so its ``tree index'' is
``12''. The fresh variable name is ``$\OTvar{:hb:12:1}$''.

%\TODO{I have not yet been in details through the next section, but we
%  need to shorten them...}

\subsection{Combining and Matching using TR2}

Our implementation does not build explicitly the proof trees for each
open transition, but instead : [Combining] enumerate all possible combinations of
pLTS transition (using TR1) and each synchronization vector choice
(TR2); [Matching] build the corresponding transition, and in particular
its predicate, in a hierarchical manner.
Each time we apply Tr2, we build a predicate as a conjunction of two
parts. The first part, which is the simpler one, composes the
predicate from the 
subnets and the guard from transitions in sub-pLTS or synchronization
vectors. The second part is the conjunction of equations generated
during matching the subnets' behaviors with the synchronization
vectors.

%Note that it is simple to collect required terms in the first
%part while in the second part we need to find out all the possible
%conjunction forms. We present here a algorithm divided in two
%step. Combining enumerates all the possible combinations of the
%working status. Matching synchronizes the subnets according to
%synchronization vectors to generate possible predicates.

%Combining is to generate all the possible matchings of the subnets' OTs (open transitions). The result list of the matched tuples called *combination*.
% 1. Initialize the result list of OT tuples.
% 2. Get one list of OTs. It's same as get OTs of one subnet.
% 3. Add *null* into the list, it means this subnet does not work this time.
% 4. Combine the rest subnets' OTs. Get the combination of the rest OTs.
% 5. If there are OTs of the chosen list not be chosen, choose one OT.
%      1. If there are tuples of the combination not be chosen, choose one tuple.
%          1. Add the chosen OT into the tuple.
%          2. Add the tuple into the result list, then do the recursion.
% 6. Return the result list.

%We first conduct combining. For each subnet, we enumerate all the
%possible cases of its working status. The result action of a pLTS or a
%pNet is the external action of the node shows the working status of
%it.
%
%For holes, their behavior is unspecified, it doesn't have an exact
%external presentation of action. We only give an action variable to each hole
%and treat it as hole behavior to suppose working status of the
%hole.
%% Here should have a constraint that this variable belongs to the sort of hole.
%In every case, the hole behaviors are unchanged, so we
%only make the combinations of subnets' open transitions. 
%
%Algorithm 1 shows the combining algorithm for enumerating all the possible combinations of open transitions. The algorithm initializes an empty list $L_C$. Dealing with the list $TR$ containing several sets of open transitions from different subnets, the algorithm only chooses one of sets $L_{ot}$ and does combining on the remaining part of $TR$ recursively then get a partial result of combining, $L'_C$. Since there could exist the case that the subnet is not working, a $null$ is added into the $L_{ot}$ to represent it. We get much more combinations after combining $ot$ from $L_{ot}$ with the partial result.

\paragraph{Combining:}
Algorithm 1 shows the combining algorithm for enumerating all possible
working status combinations from subnets. Dealing with a list $TR$
containing  sets of open transitions from different subnets,
the algorithm choses one of sets $L_{ot}$ and does combining on the
remaining part of $TR$ recursively. The case when a subnet is not working is
managed by adding a $null$ into the $L_{ot}$.

\begin{algorithm}
\caption{Combining}
\begin{algorithmic}[1]

%\Require The list $TR$ of the open transition sets of the subnets. 
%\Ensure The list of all the possible combinations of open transitions $L_C$.
\State Initialize an empty combination list $L_C$;
\State Extract a list of open transitions $L_{ot}$ from $TR$;
\State Add $null$ to $L_{ot}$;
\State Get the combinations $L'_C$ of the $TR$;
\For{each $ot \in L_{ot}$}
	\For{each $C' \in L'_C$}
		\State Add $ot$ into $C'$ to get a new tuple $C$;
		\State Add $C$ into $L_C$;
	\EndFor
\EndFor 
\State \Return $L_C$;

\end{algorithmic}  
\end{algorithm}

% Mixing is to match the combination of the subnets' open transitions, hole behaviors and result action with the SV (synchronization vector) to generate the possible predicates of the OT (open transition). At the same time, other parameters of the OT (global state, subnets' transitions, result action...) are also generated. Then the possible OTs of the pNet can be achieved. 
% 1. Initialize the combination through the method *combineOpenTransitions( )*.
% 2. Initialize the result list.
% 3. Get the number of subnets and holes. They are used to decide how many element of the SV should be matched with them.
% 4. For all tuples of the combination.
%      1. Invoke the *clone( )* method to generate the fresh SV.
%      2. Match the elements of the tuple with the elements of the SV. If there is a *null* matched with a *not null* action, filter this matching. Add the result expression and the predicate of the subnets into the predicate. Generate the target global state correspondingly.
%      3. Match the hole behaviors with the elements of the SV. The hole behavior could be *null*. Add the result expression into the predicate.
%      4. Match the result action with the result action of SV. Add the result expression into the predicate.
%      5. Add the predicate into the OT. Store the OT in the result list.  
% 5. Return the result list of all the possible OTs.

%We now come to the matching using combinations from the previous step, together with hole behaviors and the result action, to match with the synchronization vectors. 
%
%Algorithm 2 shows the algorithm matches combination $L_C$, hole
%behaviors $B$ with synchronization vectors $SV$. Remember variables
%inside a synchronization vector are consider local to the vector; so every time we choose a synchronization vector $sv$ from $SV$ for matching, we clone it and use the cloned one $sv'$. We use the definition declared before to construct all the possible predicates but the algorithm has not checked the correctness of these predicates. 
%
%There is one mismatch case we can easily detect here, that is its
%working status:
%At any time an action $e_1$ matching with an element $e_2$ from
%synchronization vector, the pair $(e_1, e_2)$ must be checked: if both
%$e_1$ and $e_2$ are $null$ or $not\ null$, then it continues to
%generate a new term of predicate. Otherwise, it is intuitive that the
%working status coming from the subnet $e_1$ is different from what
%asked by the synchronization vector $e_2$, then this case is filtered
%out.

\paragraph{Matching:} Algorithm 2 shows the algorithm matches
combination $L_C$, hole behaviors $B$ with synchronization vectors
$SV$ according to the definition of the predicate declared before. We
clone the $sv$ using the fresh function to generate a new $sv'$ with
fresh variables.The mismatch on working status is easy to detect so we
filter them in the algorithm through checking the matching elements
$e_1$ and $e_2$. 
Then we build the resulting open transition. Satisfiability of the
predicate is not checked at this level, but only later, at the
toplevel of the pNet.

\begin{algorithm}
\caption{Matching}
\begin{algorithmic}[1]

%\Require 
%The combination of subnets' open transitions $L_C$;
%The behaviors of the holes $B$;
%The set of synchronization vectors $SV$.
%\Ensure 
%The list of all the possible open transitions generated $L_{ot}$.
\State Initial an empty result list $L_{ot}$;
\For{each $C \in L_C$}
	\For{each $sv \in SV$}
		\State Clone the $sv$, the result is $sv'$;
		\State Generate the fresh result action $ra$;
		\State Combine $C, B, ra$ as a new vector $v$;
		\For{each $(e_1 \in v )\&\& (e_2 \in sv')$}
			\If{$(e_1\ is\ null\ \&\&\ e_2\ is\ not\ null)\ ||\ (e_1\ is\ not\ null\ \&\&\ e_2\ is\ null)$} 
			\State Skip;
			\Else ~ Add $(e_1 = e_2)$ into the $Pred$;
			\EndIf
		\EndFor
		\State Generate the result open transition $ot$ with $Pred$;
		\State Add the $ot$ into $L_{ot}$;
	\EndFor
\EndFor 
\State \Return $L_{ot}$;

\end{algorithmic} 
\end{algorithm}

\subsection{Management of state variable assignments}
%The pLTS contains some variables in each of its states (state
%variables), or just as global states of the pLTS (global variables).
%Both can be assigned in the pLTS transitions.
%The algorithm composing assignments from subnets is different from the
%generation of predicates. 
%%When applying Tr1, we just keep the
%%assignments in the Post of the open transition. And when it comes to
%%Tr2, Post is the union of the assignments propagate from the
%%subnets. 
%The assignments provide value of the variables in
%predicates. There is usually not only one assignment for a state variable as
%several incoming transitions of a state can perform different
%assignments on the same variable. The values of the
%variables are necessary for checking satisfiability of the open
%transitions together with predicates, miss of them may drop the
%unsatisfiable results. In order to avoid such mistakes, the algorithm
%manages the variables together with a list.
%
%The list we used contains triples $<v,\ S, \ AssignRH>$ set for each pLTS to record
%the information of both state and global variables of the pLTS where $v$ is the variable in pLTS, $S$ is its owner state and $AssignRH$ is a list of expressions over other variables in the right hand side of assignments of $v$. The $S$ can be $null$ in the record triple, it means the variable is global in the pLTS belonging to all the states. 
%%So it should be noted that the changing of the value of the global state variable every time a transition is conducted. $AssignRH$ is the collection of all the possible expressions coming from the the right hand side of assignments. It keeps the information of the assignments for each variable. 
%We update the list of record triples every time Tr1 is applied, adding the right side of the assignments into corresponding expression lists. So the list keeps the all the assignments in the pLTS. Then we can track all the possible value of the variables when computing open transitions. 
%The initial value of the state variable is also seen as an assignment
%for that variable. We must give a initial value to the global state
%variable or there will be some problem in the start state. However, we
%sometimes don't give initial values to the state variables (except the
%global variables) as their initial values are decided by their first
%assignments. Checking whether each variable is initialized before
%being used is a separated problem, that should be done before starting
%our algorithm.

In a pLTS, there may be several incoming transitions of some states
that assign potentially different values to a state variable.
To handle such cases, the algorithm manages the variables together
with a list for each pLTS.
The list we used contains several triples <$v$, $S$, $AssignRH$> 
where $v$ is the variable in pLTS, $S$ is its owner state and
$AssignRH$ is a list of expressions over other variables in the right
hand side of assignments of $v$.
As a pragmatic extension to the formal definition, we also manage
``global variables'', defined in all states of the pLTS.

%There are some assignment on the variables in each of states of the pLTS (state variables), or just as global states of it (global variables). There is usually not only one assignment for a state variable, miss of them may drop the unsatisfiable results. In order to avoid such mistakes,

%% \QIN{
%% The list we used contains several triples <$v$, $S$, $AssignRH$> to
%% record the information of both state and global variables of the pLTS,
%% where $v$ is the variable in pLTS, $S$ is its owner state and
%% $AssignRH$ is a list of expressions over other variables in the right
%% hand side of assignments of $v$. It means the variable is global in
%% the pLTS belonging to all the states, if $S$ is $null$. We update the
%% list of record triples every time Tr1 is applied. Then we can track
%% all the possible (symbolic) value of the variables when computing open transitions.
%% }

\subsection{Pruning the unsatisfiable results}
\label{section:pruning}
%We have used a brute-force method to generate all the possible open
%transitions in the open automaton. It obviously contains some of
%unsatisfiability in the open transitions. In
%Fig~\ref{schema:unsat-ot}, we display an unsatisfiable open transition
%from the result of the CCS running example. It shows the case that the
%top level transmits out the action from the PN2, formula $a.P$, as
%this time only PN2 is working and C1 is performing action $a$ when
%hole $P$ is also working. When matching this case against the second
%synchronization 
%vector of PN2, transmitting the behavior from hole $P$ to the higher
%level, is used while it is obviously violate the guard of the
%vector. So we can easily find the contradiction between the generated
%predicate term $ra:152:1=l \!\land\! a=:ra:152:1$ and the guard
%composed into the predicate $a \neq l$.

We use a brute-force method to generate all the possible open
transitions in the open automaton, using all possible combinations of
synchronization vectors. Naturally this builds some transitions where
the predicates express incompatible constraints. Even if having an
unsatisfiable (symbolic) transition in the open automaton would not be incorrect,
we want to check them for satisfiability, and reduce the number of
transitions and states in the automaton.
In Fig.~\ref{schema:unsat-ot}, we display an example of an
unsatisfiable open transition 
from the result of our example. It shows the case where
the controller C1 wants to move the control from $P$ to $(acc(x);Q)$,
conducting a $\delta$ transition. However, a synchronization
vector is chosen that does not match with this action, and we can
easily find the contradiction in the 
generated predicate term ``$\OTvar{:ra:11:1}=l \!\OTland\! \delta=\OTvar{:ra:11:1}$''.

Checking satisfiability requires some symbolic computation
on the action expressions and the predicates, and this 
may depend on the specific theory of the action algebra. 
We choose to use the SMT solver Z3 as the checker engine, considering
the predicates may contain the expression 
using uninterpreted functions declared by users.
The ``Modulo Theory'' part of SMT solvers is important here, so that
the solver can use specific properties of each action algebra.

\begin{figure}[t]
%$\begin{array}{@{}r@{}l@{}}
%%{s0---a-->s1},   {---:hb:15:2-->},   (:ra:15:2 = act:sva:1:3)/\(:ra:152:1 = l)/\(a = :ra:152:1)/\(l!= a)/\(:hb:15:2 = y:sva:15:2)/\(:ra:15:2 = y:sva:15:2)/\(y:sva:15:2 != l)/\(:ra:1:3 = act:sva:1:3)
%    ot  = &\openrule{\{0 \xrightarrow{a} 1\}, ~~\{\xrightarrow{:hb:15:2}\}, ~~:ra:15:2=act:sva:1:3 \\
%    \OTland:ra:152:1=l \OTland a=:ra:152:1 \OTland a \neq l \OTland :hb:15:2 = y:sva:15:2 \\
%    \OTland :ra:15:2 = y:sva:15:2  \OTland y:sva:15:2 \neq l \OTland :ra:1:3 = act:sva:1:3}
%    {\ostate{00} \xrightarrow{:ra:1:3} \ostate{10}}\\
%  \end{array}$
%\caption{One of the unsatisfiable open transitions in CCS running example}  \label{schema:unsat-ot}
%\end{figure}

%{s0---delta-->s0},   {---:hb_P:12:1-->},   (:ra:11:1=l)/\(delta=:ra:11:1)/\(var_s0=0)/\(:hb_P:12:1=a1:sva_SV0:1:1)/\(:ra:1:1=a1:sva_SV0:1:1)/\(a1:sva_SV0:1:1!=delta(x)),   {var_s0 := 1}
$\begin{array}{@{}r@{}l@{}}
    ot  = &\openrule{\{0 \xrightarrow{\delta} 0\}, ~~\{\xrightarrow{:hb:12:1}\}, ~~\OTvar{:ra:11:1}=l \OTland \delta=\OTvar{:ra:11:1} \OTland \OTvar{s0}=0 \\
     \OTland \OTvar{:hb:12:1}=\OTvar{a1:sva:1:1} \OTland \OTvar{:ra:1:1}=\OTvar{a1:sva:1:1} \OTland \OTvar{a1:sva:1:1} \neq \delta(x), ~~\{\OTvar{s0} \leftarrow 1\}}
    {\ostate{00} \xrightarrow{\OTvar{:ra:1:1}} \ostate{00}}\\
  \end{array}$
\caption{One of the unsatisfiable open transitions in LOTOS running example}  \label{schema:unsat-ot}
\end{figure}

\begin{figure}[t]
  %  \centerline{\includegraphics[width=12cm]{XFIG/SMTLIB2}}
    \centerline{\includegraphics[width=\linewidth]{XFIG/SMTLIB2}}
  \caption{The input of the Z3 solver in SMT-LIB language and the output result}  \label{schema:smt-lib}
\end{figure}

%% \begin{figure}[t]
%%   \centerline{\includegraphics[width=10cm]{XFIG/Interaction}}
%%   \caption{The architecture of the interaction between pNet API and Z3}  \label{schema:interaction}
%% \end{figure}

\paragraph{Interaction with Z3}
%\TODO{1. example input of Z3 in SMT-LIB.[DONE] 2. software architecture.[DONE]}
%There are several ways to interact with the Z3 solver. 
%It is possible to use Z3 interactively using the SMT-LIB
%(Python-based) language. 
From inside our algorithm code, we submit satisfiability requests to
Z3 using its JAVA API. In this article, to make the reading easier, we
show the Z3 code fragments using its SMT-LIB input language. As an
example in
Fig~\ref{schema:smt-lib}, we show the input for checking the (unsatisfiable)
transition of Fig.~\ref{schema:unsat-ot}.

The input to Z3 contains the declaration of the LOTOS action algebra types
and constructors, then the declaration of variables, and finally the
predicate to be checked, encoded as a set of assertions.
Generally, some $datatypes$ are declared to input the sorts, while here
we only declare one $datatypes$ named $Action$ to involve all the
constant actions. Then we input the predicate of that open transition
in two parts. The variables contained in the predicate are declared
first, then the terms of predicate are inputed as $assertions$ one by
one. The result "unsat" in the output is just what we expected.

%\TODO{edit the figure to make it smaller in height [DONE]}

%We present the architecture of the interaction with Z3 in
%Fig~\ref{schema:interaction}.
%For each OT generated at the toplevel of a pNet, the predicate must be
%submitted for satisfiability check to Z3. For that we need to
%translate into Z3 (Java-API) syntax, both the elements of the action
%algebra, and the predicates itself, including the encoding of the
%variable assignments (the Post part).
%
%The $presentation$ of the action algebra
%is declared at the beginning of the algorithm as the $sorts$ and the
%$operators$ will be used throughout the whole algorithm. It is
%translated into $datatypes$ and $functions$ then submitted to the Z3
%at the same time. Each candidate predicate will then be translated
%into $assertions$ before submitted to Z3 after the algorithm
%computing out all the possible open transitions, they act as
%intermediate results there. Together with the $datatypes$ and
%$functions$, the $assertions$ are used to check the satisfiability of
%the predicate in Z3.

To build the input submitted to Z3 for each OT,
we translate the algebra $presentation$, the predicates and the
variable assignments (the Post part) into Z3 (Java-API) syntax.
% The $presentation$ and the predicates are translated separately.

\paragraph{Translation of Action algebra presentation}

The $datatypes$ are declared according to the $sorts$ the users
declared in the $presentation$. If the return type of the $functions$
is one of the $datatypes$, we add this $function$ to the $datatypes$ declaration.
So when we declare the operator ACT we illustrated in section 4.1, it
will be just declared together with the constants like :\\
\centerline{$\OTvar{(ACT\ (action3\ Action)(data\ Int))}$}

The $functions$ can also be declared independently, it usually applies to functions returning integer or boolean, in the format like:\\
\centerline{$\OTvar{(declare\text{-}fun\ ACT\ (Action\ Int)\ Action)}$}

%Although in Fig~\ref{schema:smt-lib} we have not declared $functions$ outside of $datatypes$, there still exists $functions$ declared independently. The input and output types of the $function$ are also determined by the construct of the  $operators$ in the presentation. For example, we define a new operator to present the equation between two actions in LOTOS example $\text{EQUAL}$ in the format likes:
%
%\[EQUAL: \quad < Action , Action > \rightarrow Bool .\]
%
%It will be translated into the Z3 statement: 
%
%\[\text{(declare-fun\ EQUAL\ (Action\ Action)\ Bool) .}\]

\paragraph{Translating assignments into predicate terms}

%The assignments are also taken into account when checking the
%satisfiability of the open transitions as they represent the value of
%the variables involved in an OT. In order to do that, we employs a
%translation from assignments into term of predicates. The assignments
%required to be translated are only those belong to the variables
%contain in the start states of the open transitions. The assignments
%are represented as equations between the variables and the expressions
%on the right side of assignment. For the assignments of the same
%variable, we do not figure out which state is the precursor of current
%state, we keep all these possible equations instead, and generate the
%disjunction of these equations. The state could always move to the
%target state if there exists one of the values of its variable
%satisfies the guard.
%%\TODO{It won't matter the variable of the next state
%%except that the right hand side expression contain its variables. This
%%is a problem left to be solved in our future work. WHICH PROBLEM ?}
%After that, for the assignments of the different variables, we generate a conjunction of them. This way we obtain a new term of the predicate involving assignments of the variables.

State-variable assignments are also treated as a part of the
predicates when checking the satisfiability. Here we need conduct some
translations, such as $\{s_0 \leftarrow 1\}$ in section
~\ref{section:op-semantics}, we translate it into an equation
$s_0=1$. Then for the assignments of the same variable, we keep all
possible equations, and generate the disjunction of
these equations. Correspondingly, we generate a conjunction for the
assignments from the different variables.

\paragraph{Checking satisfiability}

% We check the satisfiability of open transitions on the top level of
% the pNet, treating these open transitions as the intermediate result
% of the algorithm. After we get the intermediate results of open
% transitions, we check them one by one.
Here we declare first all the variables in the predicates, with their
sorts.
An example of syntax is::\\
\centerline{$\OTvar{(declare-const\ |:ra:1:1|\ Action)}$}

\noindent Then each term of the
predicate is an assertion in Z3, as an example, in the former
unsatisfiable result in Fig~\ref{schema:unsat-ot}, the contradiction is occurred in terms: \\
\centerline{$\OTvar{(:ra:11:1=l)} \OTland \OTvar{(delta=:ra:11:1)}$}\\
from which we generate the following two assertions:\\
\centerline{$\OTvar{(assert\ (=\ |:ra:11:1|\ l))\ \ \ (assert\ (=\ delta\ |:ra:11:1|))}$}

\noindent The predicate submitted to the Z3 also contains the
assignments encodings.
In the case of $ot_2$, it
gives the assertion:\\
\centerline{$\OTvar{assert (or (= var\_s0 0) (= var\_s0 01)))}$}

\subsection{Result of the running example}
\label{section:full-result}
% \TODO{
% Xudong: full result of the tool execution on the running example\\\\
% Open Automata of the CCS example.

Running our algorithm for our LOTOS example yields 30 open
transitions, out of which 25 are detected unsatisfiable by Z3, so we
get only 5 results left; the resulting open automaton is shown in
Fig. ~\ref{schema:lotos-result}, we have drawn manually the open
automaton obtained as the final result generated by algorithm.

The $ot_1$ transition means that the actions of the process $P$ could
be transmitted out, until $P$ performs a $\delta(x)$ and the value of
$x$ is transmitted, through transition
$ot_2$. The other 3 open transitions shows the behaviors of the
process $Q$ while the $ot_3$ and $ot_4$ won't be performed as the
controller has already set the value $s_0$ to 1 in the previous
action.

To sum up, from the pNet encoding LOTOS expression, inputed by the user, we
generate the open transitions and prune the unsatisfiability
automatically. The final result open transitions represent all the
possible movements of the LOTOS example. They can exactly construct
the open automaton for the given example.

\begin{figure}[h]
  \centerline{\includegraphics[width=5cm]{XFIG/LOTOSresult}}

$\begin{array}{@{}r@{}l@{}}
    ot_1  = &\openrule{\{0 \xrightarrow{l}_{C_1} 0\}, ~~\{\xrightarrow{hb_{11}}_P\}, ~~[s_0=0 \!\land\! hb_{11}=a_1 \!\land\! a_1\neq \delta(x) \!\land\! v_1=hb_{11}]}
    {\ostate{00} \xrightarrow{v_1} \ostate{00}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_2  = &\openrule{\{0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1\}, ~~\{\xrightarrow{hb_{12}}_P\}, ~~[s_0=0 \!\land\! hb_{12}=\delta(x) \!\land\! v_2=\underline{\delta(x)}], ~~\{s_0 \xleftarrow{} 1\}}
    {\ostate{00} \xrightarrow{v_2} \ostate{01}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_3  = &\openrule{\{0 \xrightarrow{l}_{C_1} 0\}, ~~\{\xrightarrow{hb_{13}}_P\}, ~~[s_0=0 \!\land\! hb_{13}=a_1 \!\land\! a_1\neq \delta(x) \!\land\! v_3=a_1]}
    {\ostate{01} \xrightarrow{v_3} \ostate{01}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_4  = &\openrule{\{0 \xrightarrow{\delta}_{C_1} 0, 1 \xrightarrow{acc(x)}_{C_2} 1\}, ~~\{\xrightarrow{hb_{14}}_P\}, ~~[s_0=0 \!\land\! hb_{14}=\delta(x) \!\land\! v_4=\underline{\delta(x)}], ~~\{s_0 \xleftarrow{} 1\}}
    {\ostate{01} \xrightarrow{v_5} \ostate{01}}\\
  \end{array}$

  $\begin{array}{@{}r@{}l@{}}
    ot_5  = &\openrule{\{0 \xrightarrow{r}_{C_1} 0, 1 \xrightarrow{q}_{C_2} 1\}, ~~\{\xrightarrow{hb_{15}}_Q\}, ~~[s_0=1 \!\land\! hb_{15}=a_2 \!\land\! a_2\neq acc(x) \!\land\! v_5=a_2]}
    {\ostate{01} \xrightarrow{v_5} \ostate{01}}\\
  \end{array}$

\caption{The open automaton of the LOTOS formula}  \label{schema:lotos-result}
\end{figure}

\section{Conclusion and Discussion}
\label{section:conclusion}

In this paper we presented the algorithm for generating an open
automaton representing the semantics of an open pNet, and we described
the implementation of the
algorithm. Our implementation includes two main parts. First, we
compute all the possible open transitions. The actions in pLTS and
hole behaviors are composed in the algorithm while the generation of
predicate need to match the combinations of the subnets with the
synchronization vectors. Some of the open transitions obtained at this
intermediate step, constructed in a structural manner from all
possible combinations of possible logical predicates, may
contain predicates which do not represent any possible concrete
instantiations. 
To get rid of these useless transitions, we use a SMT solver Z3 for checking the
satisfiability of the predicate in each open transition. In
order to do that, we have done some translation on the presentation,
the predicate and the assignments of the variables which are also
considered as a part of predicates before submitting them to the
Z3 solver. We implemented our algorithm in the VerCors platform and use some
running examples, the formula from other process algebras, to test the
algorithm. The result of the running examples are shown as the open
automaton with all the parameters generated by the algorithm
automatically. The open automaton successfully presents all the
possible movements of the LOTOS expression.

Among the questions arising during this work, an important one is the
efficiency of Z3 for our needs. Z3 is famous for being very fast for
solving very large sets of assertions, and this could be important for
us. But we encountered  some problems during the implementation the
checking method using Z3. For example, if the $functions$ and the
$datatypes$ part of the algebra presentation we submit to the solver
are independent with each other, then checking the $assertions$ is a
simple work. However, $functions$ on recursive $datatypes$ make it
more complex, some special rules might be defined by user for the
induction. Also, we need more evidence that dealing with more complex
action algebras, that would involve axiomatizing their dat structures
into Z3 theories, will indeed allow us to decide validity of
predicates.
We may also try other automatic theorem provers, depending on the structure of the proofs we need.

Naturally, the next work after the generation of the open automata
will be to check the bisimulation of the pNets. In previous paper, we
have already found the FH-bisimulation, inspired by~\cite{deSimone85} as a
suitable definition. Usual approaches of bisimulation by partition
refinement does not work easily with symbolic transitions. In a first
step we may want instead to define and implement an algorithm fo
r checking that a given equivalence relation between open automata
states is indeed a FH-Bisimulation. Finding automatically a
bisimulation relation or the most general bisimulation will be more
challenging. It appears also that when comparing different pNets with
different structure, strong bisimulation will be too strong a notion,
and that we will have to define proper notions of weak equivalence or
refinements.

Before that work, we need to reconsider the result open transitions
from our algorithm. The ideal
result that we showed in the bottom part of
Fig.~\ref{schema:lotos-result} is not the one generated by the algorithm,
because it contains many intermediate
variables, such as fresh local variables of synchronization
vectors, and fresh result variables of intermediate open transitions.
Such intermediate variables come from the way predicates are
generated, and depends on the particular structure of the pNet. But
they are not significant in the resulting behavior, and should be
eliminated before comparing such behavior with another pNet.
To eliminate the intermediate variables in the predicate, we want to define
the set of the significant variables that we won't eliminate: the
local variables of pLTS transitions, the hole behaviors and the result
action at the top level.
The result of the elimination will be simpler predicates, that can be
successfully compared (e.g. by Z3) within the bisimulation checking
algorithm.

\bibliographystyle{lncs/splncs}

% \bibliography{oasis,biblio}
\bibliography{biblio}

% \newpage
% \appendix

\end{document}