### 2018-10-20 2018-Arch-pNets/arch-pNets.tex
Arch-pNet paper: Added an appendix with the summary of algebraic representations of connectors and encodings among them Added a figure with three representations of the four connectors from the Failure Monitor Added a simplified encoding (almost complete) -- to be replaced by the exact one, but committing for the record -- could be useful some other time.

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
\documentclass{llncs}
\usepackage{llncsdoc}

%\usepackage{a4wide}

\usepackage{graphicx}
\usepackage{verbatim}

\let\proof\relax   
\let\endproof\relax

\usepackage{amsmath,amsthm,amscd}
\usepackage{amssymb}    % for blackboard B
\usepackage{rotating}   % to rotate the figures
\usepackage{mathpartir} % math paragraph + inference rules
\usepackage{mathtools}  % for align, multline etc.
\usepackage{caption}
\usepackage{stmaryrd} % double brackets (\llbracket, \rrbracket)

%% \usepackage[justification=centering,belowskip=-10pt,aboveskip=0pt]{caption}
%% \setlength{\intextsep}{10pt plus 2pt minus 2pt}
%% \setlength\abovedisplayskip{0pt}

\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{hyperref}

\input{figures/macros.tex}         % Definitions of macros used in XFig figures

\usepackage{soul}
\soulregister\cite7
\soulregister\ref7
\soulregister\pageref7
\soulregister{\em}{0}
\soulregister{\mdash}{0}
\soulregister{\st}{7}
\soulregister{\bf}{0}
\soulregister{\fig}{7}
\soulregister{\ie}{0}
\soulregister{\etc}{0}
\soulregister{\eg}{0}
\soulregister{\wrt}{0}
\soulregister{\resp}{0}
\soulregister{\cf}{0}

\soulregister{\cstate}{7}
\soulregister{\tstate}{7}
\soulregister{\goesto}{7}

\usepackage[colorinlistoftodos,bordercolor=white]{todonotes}
%% \usepackage[disable,colorinlistoftodos,bordercolor=white]{todonotes}
\usepackage{macrospNets}

\newcommand{\Simon}{\\\hfill\mdash Simon}

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{#2}\Simon}}
\newcommand{\noteSBin}[2][inline,color=green!40]{\todo[#1]{{#2}\Simon}}
\newcommand{\todoSB}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Simon:} {#2}}}
\newcommand{\todoSBin}[2][inline,color=green!40]{\todo[#1]{\textbf{To-do Simon: } {#2}}}
\newcommand{\Ludo}{\\\hfill\mdash Ludo}
\newcommand{\noteLH}[2][color=orange!40, size=\tiny]{\todo[#1]{{#2}\Ludo}}
\newcommand{\noteLHin}[2][inline,color=orange!40]{\todo[#1]{{#2}\Ludo}}
\newcommand{\todoLH}[2][color=orange!40, size=\tiny]{\todo[#1]{\textbf{To-do Ludo:} {#2}}}
\newcommand{\todoLHin}[2][inline,color=orange!40]{\todo[#1]{\textbf{To-do Ludo: } {#2}}}

\newcommand{\noteIn}[2][inline,color=black!20]{\todo[#1]{{#2}}}

\newcommand{\add}[2][Added]{\todo[color=blue!20, size=\tiny]{#1}{\color{blue}#2}}
%% \newcommand{\remove}[2][Removed]{\todo[color=red!20, size=\tiny]{#1}}
\newcommand{\remove}[2][Removed]{\todo[color=red!20, size=\tiny]{#1}{\color{red}\st{#2}}}
\newcommand{\removeIn}[2][Removed]{\todo[color=red!10, inline]{{\color{red}#2}#1}}
%% \newcommand{\replace}[3][Replaced]{\todo[color=blue!20, size=\tiny]{#1}{\color{blue}#3}}
\newcommand{\replace}[3][Replaced]{\todo[color=blue!20, size=\tiny]{#1}{\color{blue}#3}{\color{red}\st{#2}}}

\newcommand{\addSB}[1]{\add[Added by Simon]{#1}}
\newcommand{\removeSB}[1]{\remove[Removed by Simon]{#1}}
\newcommand{\removeSBin}[1]{\removeIn[Removed by Simon]{#1}}
\newcommand{\replaceSB}[2]{\replace[Replaced by Simon]{#1}{#2}}

\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\newcommand{\tupleDeli}{(}
\newcommand{\tupleDelii}{)}
\newcommand{\setTupleDelims}[2][(]{
  \renewcommand{\tupleDeli}{#1}%
  \ifx#2\relax\else\renewcommand{\tupleDelii}{#2}\fi%
}
\newcommand{\tuple}[2][\ensuremath{,\allowbreak}]{%  
  \def\nextitem{\def\nextitem{#1}}% Separator
  \renewcommand*{\do}[1]{\nextitem ##1}% How to process each item
  \tupleDeli\docsvlist{#2}\tupleDelii% Process list
}

\newcommand{\pNetTuple}[2][\ensuremath{,\allowbreak}]{%
  \setTupleDelims[\mylangle]{\myrangle}%
  \tuple[#1]{#2}%
}

\newcommand{\defn}[1]{Def.~\ref{defn:#1}}
\newcommand{\defntwo}[2]{Defs~\ref{defn:#1} and \ref{defn:#2}}
\newcommand{\fig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\figs}[2]{Fig.~\ref{fig:#1} and~\ref{fig:#2}}
\newcommand{\tab}[1]{Tab.~\ref{tab:#1}}
\newcommand{\eq}[1]{(\ref{eq:#1})}
\newcommand{\res}[1]{(\ref{res:#1})}
\newcommand{\ex}[1]{Ex.~\ref{ex:#1}}
\newcommand{\exs}[2]{Ex.~\ref{ex:#1} and~\ref{ex:#2}}
\newcommand{\secn}[1]{Sect.~\ref{secn:#1}}
\newcommand{\app}[1]{App.~\ref{secn:#1}}
\newcommand{\rem}[1]{Note~\ref{rem:#1}}
\newcommand{\lem}[1]{Lem.~\ref{lem:#1}}
\newcommand{\cor}[1]{Cor.~\ref{cor:#1}}
\newcommand{\thm}[1]{Th.~\ref{thm:#1}}
\newcommand{\axs}[1]{Ax.~\ref{ax:#1}}
% \newcommand{\ax}[2]{\ref{ax:#1}.\ref{ax:#1:#2}}
\newcommand{\ax}[1]{Ax.~\ref{ax:#1}}
\newcommand{\prop}[1]{Prop.~\ref{prop:#1}}
\newcommand{\alg}[1]{Alg.~\ref{alg:#1}}
\newcommand{\asmp}[1]{Ass.~\ref{asmp:#1}}

% %%%%%%%%%%%%%%%%%%%%%

\newcommand{\cA}{\ensuremath{\mathcal{A}}}
\newcommand{\fA}{\ensuremath{\mathsf{A}}}
\newcommand{\sA}{\ensuremath{\mathbb{A}}}
\newcommand{\cB}{\ensuremath{\mathcal{B}}}
\newcommand{\fB}{\ensuremath{\mathsf{B}}}
\newcommand{\sB}{\ensuremath{\mathbb{B}}}
\newcommand{\cC}{\ensuremath{\mathcal{C}}}
\newcommand{\sC}{\ensuremath{\mathbb{C}}}
\newcommand{\cD}{\ensuremath{\mathcal{D}}}
\newcommand{\fD}{\ensuremath{\mathsf{D}}}
\newcommand{\sD}{\ensuremath{\mathbb{D}}}
\newcommand{\cE}{\ensuremath{\mathcal{E}}}
\newcommand{\sE}{\ensuremath{\mathbb{E}}}
\newcommand{\cF}{\ensuremath{\mathcal{F}}}
\newcommand{\cG}{\ensuremath{\mathcal{G}}}
\newcommand{\cH}{\ensuremath{\mathcal{H}}}
\newcommand{\cI}{\ensuremath{\mathcal{I}}}
\newcommand{\sI}{\ensuremath{\mathbb{I}}}
\newcommand{\cM}{\ensuremath{\mathcal{M}}}
\newcommand{\cN}{\ensuremath{\mathcal{N}}}
\newcommand{\sN}{\ensuremath{\mathbb{N}}}
\newcommand{\cP}{\ensuremath{\mathcal{P}}}
\newcommand{\sP}{\ensuremath{\mathbb{P}}}
\newcommand{\cQ}{\ensuremath{\mathcal{Q}}}
\newcommand{\sQ}{\ensuremath{\mathbb{Q}}}
\newcommand{\cR}{\ensuremath{\mathcal{R}}}
\newcommand{\sR}{\ensuremath{\mathbb{R}}}
\newcommand{\cS}{\ensuremath{\mathcal{S}}}
\newcommand{\sS}{\ensuremath{\mathfrak{S}}}
\newcommand{\cT}{\ensuremath{\mathcal{T}}}
\newcommand{\fT}{\ensuremath{\mathsf{T}}}
\newcommand{\sT}{\ensuremath{\mathbb{T}}}
\newcommand{\cV}{\ensuremath{\mathcal{V}}}
\newcommand{\fV}{\ensuremath{\mathsf{V}}}
\newcommand{\sV}{\ensuremath{\mathbb{V}}}
\newcommand{\sZ}{\ensuremath{\mathbb{Z}}}

\newcommand{\mdash}[1][]{---#1}
\newcommand{\ndash}{--}
\newcommand{\ie}[1][\ ]{i.e.#1}
\newcommand{\etc}[1][\ ]{etc.#1}
\newcommand{\eg}[1][\ ]{e.g.#1}
\newcommand{\cf}[1][\ ]{cf.#1}
\newcommand{\wrt}[1][\ ]{w.r.t.#1}
\newcommand{\resp}[1][\ ]{resp.#1}

\newcommand{\bydef}[1]{\ensuremath{\stackrel{\mathit{\scriptscriptstyle def}}{#1}}}
\newcommand{\suchthat}{\ensuremath{\,|\,}}
\newcommand{\rightsuchthat}{\ensuremath{\,\right|\,}}
\newcommand{\leftsuchthat}{\ensuremath{\,\left|\,}}
\newcommand{\setdef}[2]{\ensuremath{\{{#1}\,|\,{#2}\}}}
\newcommand{\bsetdef}[2]{\ensuremath{\bigl\{{#1}\,\bigl|\,{#2}\bigr.\bigr\}}}
\newcommand{\Setdef}[2]{\ensuremath{\Big\{{#1}\,\Big|\,{#2}\Big\}}}
\newcommand{\goesto}[2][]{\ensuremath{\xrightarrow[{#1}\relax]{#2}}}
\newcommand{\notgoesto}[2][]{\ensuremath{\not\xrightarrow[{#1}\relax]{\ \,#2}}}
\newcommand{\non}[1]{\ensuremath{\overline{#1}}}

\newcommand{\true} {\ensuremath{\mathtt{t\!t}}}
\newcommand{\false}{\ensuremath{\mathtt{f\!f}}}
\newcommand{\noop} {\ensuremath{\emptyset}} % \mathsf{skip}
\newcommand{\init} {\ensuremath{\mathsf{init}}}

%% \newcommand{\order}{\leqslant}
\newcommand{\order}{<}
\newcommand{\ordbool}{\ensuremath{\sB^{\order}}}
\newcommand{\data}{\ensuremath{\sD}}
\newcommand{\signature}{\ensuremath{\Sigma}}
\newcommand{\variables}{\ensuremath{\cV}}
\newcommand{\Talg}{\ensuremath{\cT_{\signature,\variables}}}
\newcommand{\actions}[1]{\ensuremath{\cA_{#1}}}
\newcommand{\exprs}[1]{\ensuremath{\sE_{#1}}}
\newcommand{\monexprs}[1]{\ensuremath{\sE^{\order}_{#1}}}
\newcommand{\boolexprs}[1]{\ensuremath{\sB_{#1}}}
\newcommand{\guards}[1]{\ensuremath{\ordbool_{#1}}}
\newcommand{\assigns}[1]{\ensuremath{\sA_{#1}}}
\newcommand{\updates}[1]{\ensuremath{\sA^{\order}_{#1}}}
\newcommand{\valuations}[1]{\ensuremath{\data^{#1}}}
\newcommand{\val}[3][]{\ensuremath{#1{\sigma}^{#2}_{#3}}}

\newcommand{\filter}[2][]{\ensuremath{\pi_{#1}({#2})}}

\newcommand{\primeit}[1]{#1'}
\newcommand{\secondit}[1]{#1''}

\newcommand{\twosynch}{%
  \mbox{\ensuremath{\bullet\!\!\!-\!\!\!-\!\!\!-\!\!\!\bullet}}}
\newcommand{\synchtrig}{%
  \mbox{\ensuremath{\bullet\!\!\!-\!\!\!-\!\!\!\blacktriangleleft}}}
\newcommand{\trigsynch}{%
  \mbox{\ensuremath{\blacktriangleright\!\!\!-\!\!\!-\!\!\!\bullet}}}
\newcommand{\twotrig}{%
  \mbox{\ensuremath{\blacktriangleright\!\!\!-\!\!\!-\!\!\!\blacktriangleleft}}}

\newcommand{\export}[1][]{\ensuremath{\varepsilon_{#1}}}
\newcommand{\valdiff}[2]{\ensuremath{#1 \triangle #2}}
\newcommand{\supp}[1]{\ensuremath{\mathrm{supp}(#1)}}
\newcommand{\semopen}[1]{\ensuremath{[{#1}]}}
\newcommand{\semclosed}[1]{\ensuremath{\llbracket{#1}\rrbracket}}
\newcommand{\reachable}[1]{\ensuremath{\mathit{reachable}({#1})}}
\newcommand{\IMextend}[2]{\ensuremath{#1 \ltimes #2}}
\newcommand{\arcomp}{\oplus}
\newcommand{\arequiv}{\equiv}
\newcommand{\Arcomp}{\Bigoplus}
\newcommand{\expmix}{\wedge}

\newcommand{\intsem}[1]{\ensuremath{\|{#1}\|}}
\newcommand{\aisem}[1]{\ensuremath{|{#1}|}}

\newtheorem*{syntax}{Syntax}
\newtheorem*{semantics}{Semantics}

\newcommand{\ai}{\ensuremath{\mathcal{AI}}}
\newcommand{\ct}{\ensuremath{\mathcal{T}}}
\newcommand{\cru}{\ensuremath{\mathcal{CR}}}
\newcommand{\ac}{\ensuremath{\mathcal{AC}}}

\newcommand{\addspace}[1]{#1 }
\newcommand{\nopri}[1][]{\ensuremath{\mathit{pNet}%
    \ifx#1\relax\else(#1)\fi}}
\newcommand{\maxprog}[1][]{\ensuremath{\mathit{pNet}_\mu%
    \ifx#1\relax\else(#1)\fi}}

\newcommand{\partition}{\cD}

\makeatletter
\newcommand{\doubletilde}[1]{{%
  \mathpalette\double@tilde{#1}%
}}
\newcommand{\double@tilde}[2]{%
  \sbox\z@{$\m@th#1\tilde{#2}$}%
  \ht\z@=.9\ht\z@
  \tilde{\box\z@}%
}
\makeatother

\newcounter{tempctr}
\newcommand{\breakenumistart}{%
  \setcounter{tempctr}{\value{enumi}}%
  \end{enumerate}%
}
\newcommand{\breakenumiend}{%
  \begin{enumerate}%
  \setcounter{enumi}{\value{tempctr}}%
}

% %%%%%%%%%%%%%%%%%%%%%

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{arrows,shapes,automata,petri}
  \tikzset{
  place/.style={
    circle,
    thick,
    draw=blue!75,
    fill=blue!20,
    minimum size=6mm
  },
  transition/.style={
    rectangle,
    thick,
    fill=black,
    minimum width=8mm,
    inner ysep=2pt
    },
    arc/.style = {
      decoration=
      {markings,mark=at position #1 with {circle;}
      },
      postaction={decorate,draw}},
  }

\let\llncssubparagraph\subparagraph
\let\subparagraph\llncssubparagraph

\begin{document}
\graphicspath{{figures/}}

\title{Architectures and Open pNets}

\author{%
Simon~Bliudze\inst{1}
\and
Ludovic~Henrio\inst{2}
\and
Eric~Madelaine\inst{3}
\and
...
}

\institute{%
  INRIA Lille -- Nord Europe, Villeneuve d'Ascq, France\\
  \email{simon.bliudze@inria.fr}
\and
\and
}

\maketitle

\begin{abstract}
  We extend the theory of architectures with a general
  mechanism for handling various types of data transfer.  We
  encode this extended model into the open pNet semantic
  model and show how this encoding can be used to verify
  that an architecture does, indeed, enforce its
  characteristic property.

\keywords{}
\end{abstract}

%****************************************************************
%****************************************************************

\section{Introduction}
\label{secn:introduction}

BIP (Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BliSif08-acp-tc}.
This approach ensures that any property, shown to hold on a given BIP
model, will also hold by construction on the generated code.

In this context, the notion of \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified
%% \footnote{%
%% %
%%   The current approach relies on taxonomies of predefined
%%   architectures.  Future work may include automatic synthesis of
%%   architectures from, \eg CTL~\cite{BaierKatoen2008} formulae.
%% %
%% }
and applied to the model, thereby potentially introducing additional
coordinator components and modifying the connectors that define
synchronisation patterns among ports of components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie when several architectures are applied, each enforcing a safety
property, the resulting system satisfies their conjunction.

The concepts and results of the paper are illustrated by a running
example extending the Failure Monitor architecture from the CubETH
nanosatellite on-board software case-study~\cite{CubETH-case-study}.

%****************************************************************
%****************************************************************

\section{Preliminaries}
\label{secn:preliminaries}

%****************************************************************

\subsection{Notations}
\label{secn:notations}

We extensively use indexed structures
over some countable indexed sets, which are equivalent to mappings over
the countable set. 
Thus, 
$a_i^{i\in I}$
%, or equivalently  $(i\mapsto a_i)^{i\in I}$
denotes a family of elements $a_i$ indexed over the
set $I$. % Such a family
% is equivalent to the mapping $(i\mapsto a_i)^{i\in I}$.
% To specify the set over which the structure is indexed,
% indexed structures are always denoted with an exponent of the form $i\in I$
% (arithmetic only appears in the indexes if necessary).
This notation defines both $I$ the set over which the family is
indexed (called \emph{range}), and $a_i$ the elements of the family.
E.g., $a^{i\in\{3\}}$ is the mapping with a single entry $a$ at index
$3$ ; abbreviated $(3\mapsto a)$ in the following.
When this is not
ambiguous, we shall use notations for sets, and typically write
``indexed set over I'', even though formally we should speak of multisets; and
write $x\in a_i^{i\in I}$ to mean $\exists i\in I.\, x=a_i$.  An empty
family is denoted $\emptyset$. We use a classical overline notation, \eg $\overline{a}$, 
for a family, where the indexing set is irrelevant. Finally, we use $\uplus$ to denote 
the disjoint union on
indexed sets.

\noteIn{states looks ok, but initial state is $s_0$ in pNet and not in archi : $s^0$.  Exponent 
position was removed from pNet because we found it complex with indexed set notation but 
could be ok.
\Ludo

I am using exponent notation for ``modifiers'', reserving\mdash
as much as possible\mdash index notation for enumeration.  I
think this can be solved, where necessary, by using parentheses,
\eg $(s^0_i)^{i \in I}$
\Simon
}

%****************************************************************

\subsection{Term algebra}
\label{secn:terms}

\todoLHin{Remove term algebra or simplify?}
\todoLHin{universal domain $\implies$ remove sort}
\todoLHin{replace assignments by $e_i$

  In fact, in the pNet action terms, you need expressions, not
  assignements.  Therefore assuming that all expressions are
  assignments will not work, so the right approach is to have
  expressions (for pNets) and assignments (for BIP) separately.
  \Simon
%
}

\noteSBin{In Archi we have $a$, whereas in pNets we have $\alpha$, where, approximately $\alpha = 
a(x_1..x_n)$, except that $a$ is port name in archi and action name in pNets.
Is this true? can we unify better? or is it sufficient?
\Ludo

$a$ is a set of ports, but otherwise, yes, this is true.  I think
this is coherent, since, in archi, we have the ``interaction
part'', \ie only ports, plus the data part, \ie variables,
guards, expressions.  In pNets, these are combined.  So, it seems
ok that notations differ\mdash ``uniformisation'' will be
achieved by encoding archi $\rightarrow$ pNets.
}

\noteLHin{I believe we should simplify this paragraph about param actions}
The pNet model in \secn{pNets} relies on a notion of parameterised actions, that are
symbolic expressions using data types and variables. As we aim at encoding the low-level 
behaviour of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on (some sort of) unification between two or more action
expressions, to express various kinds of communication or
synchronisation mechanisms.

%\def\Talg{\mathcal{T}_{\Sigma,\P}}  % Moved to the header

\noteSBin{proposed notations:

  $g$ are boolean expressions;
  $e_i$ are updates/assignments
  \Ludo

  In fact, I am not sure whether I have correctly understood this
  suggestion: do you mean, \eg using $g$ to denote the set of
  Boolean expressions or individual Boolean expressions? Using
  lower case letters to denote sets might be problematic, since
  we (at least I) use them a lot to denote instances, as in
  \[q \goesto{a, g, e} q'\,.\]
  %% In fact, I even use capital letters: in the definition of
  %% interaction model composition, I use
  %% \[\cdot \goesto{a, G, E} \cdot\]
  %% in the conclusion of the rule.  However, for this latter I can
  %% easily change the notation.

  I have set up macros for, general and Boolean expressions,
  actions, variables \etc[,] which I have substituted everywhere,
  where appropriate in the following paragraph.
}

  \begin{tabular}{
      @{}p{0.15\columnwidth}@{}
      p{0.35\columnwidth}@{}
      p{0.5\columnwidth}@{}
    }
    {\bf Output} & {\bf Command} & {\bf Definition}
    \\
    $\data$ &
    {\ttfamily\textbackslash data} &
    {\ttfamily\textbackslash sD}
    \\
    \signature &
    {\ttfamily\textbackslash signature} &
    {\ttfamily\textbackslash Sigma}
    \\
    \variables &
    {\ttfamily\textbackslash variables} &
    {\ttfamily\textbackslash cV}
    \\
    \Talg &
    {\ttfamily\textbackslash Talg} &
    {\ttfamily\textbackslash cT\_\{\textbackslash signature,\textbackslash variables\}}
    \\
    \actions{\#1} &
    {\ttfamily\textbackslash actions\{\#1\}} &
    {\ttfamily\textbackslash cA\_\{\#1\}}
    \\
    \exprs{\#1} &       
    {\ttfamily\textbackslash exprs\{\#1\}} &
    {\ttfamily\textbackslash sE\_\{\#1\}}
    \\
    \monexprs{\#1} &       
    {\ttfamily\textbackslash monexprs\{\#1\}} &
    {\ttfamily\textbackslash sE\^{}\{<\}\_\{\#1\}}
    \\
    \boolexprs{\#1} &       
    {\ttfamily\textbackslash boolexprs\{\#1\}} &
    {\ttfamily\textbackslash sB\_\{\#1\}}
    \\
    \guards{\#1} &
    {\ttfamily\textbackslash guards\{\#1\}} &
    {\ttfamily\textbackslash sB\^{}\{<\}\_\{\#1\}}
    \\
    \assigns{\#1} &
    {\ttfamily\textbackslash assigns\{\#1\}} &
    {\ttfamily\textbackslash sA\_\{\#1\}}
    \\
    \updates{\#1} &
    {\ttfamily\textbackslash updates\{\#1\}} &
    {\ttfamily\textbackslash sA\^{}\{<\}\_\{\#1\}}
    \\
    \valuations{\#1} &
    {\ttfamily\textbackslash valuations\{\#1\}} &
    {\ttfamily\textbackslash data\^{}\{\#1\}}
    \\[6pt]
    \val[]{\#2}{\#3} &
    {\ttfamily\textbackslash val[]\{\#2\}\{\#3\}} &
    {\ttfamily\#1\{\textbackslash sigma\}\^{}\{\#2\}\_\{\#3\}}
    \\[6pt]
    \val[\tilde]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash tilde]\{\#2\}\{\#3\}} &
    (optional argument usage examples)
    \\[6pt]
    \val[\primeit]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash primeit]\{\#2\}\{\#3\}} &
    \\[6pt]
    \val[\secondit]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash secondit]\{\#2\}\{\#3\}} &
  \end{tabular}
\noteSBin{commands above (tabular generates an error inside a
  note)}

Formally, we assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$ \replaceSB{, including}{and} a set of \emph{Boolean
expressions} \todoLH{$\mathcal{B} \to g$}
$\boolexprs{\variables}\subseteq\exprs{\variables}$.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$ \replaceSB{,}{and}
$\exprs{\variables}\cap\actions{\variables} = \emptyset$;
naturally action terms will use data expressions as subterms.
\addSB{In addition to the above notations, we will use
\[
\assigns{\variables} \bydef{=}
\Setdef{(x_i := e_i)^{i \in I}}{
  x_i^{i \in I}\subseteq \variables,
  e_i^{i \in I}\subseteq \exprs{\variables},
  I \text{ is a finite index set}
}
\]
to denote the set of \emph{variable assignments}.}  We use the symbol
$\noop$ to denote the empty assignment, \ie an assignment with
$I = \emptyset$.

To be able to reason about the data flow between pLTSs, we
distinguish \emph{input variables} of the form $?x$ within terms; the function
$\vars(t)$ identifies the set of variables in a term
$t\in\Talg$, and $iv(t)$ returns its input variables.
Action algebras can \replaceSB{encode naturally }{naturally encode }usual point-to-point message passing calculi (using 
$a(?x_1,\dots,?x_n)$ for inputs, $a(v_1,\dots,v_n)$ for outputs), but it also allows
for more general mechanisms, like gate negotiation in Lotos, or broadcast
communications.

Finally, we assume the existence of a universal data domain
given as a partially-ordered set $(\data, \order)$, potentially
encompassing several copies of any given data type with
different orders.  For example, we assume that $(\data,
\order)$ comprises both the unordered set of Booleans $\sB =
\tuple{\{\true, \false\}, \emptyset}$ and the naturally ordered one
$\sB^< = \tuple{\{\true, \false\}, \{\false < \true\}}$, and
similarly for integer and real numbers; as well as the set of
intervals ordered by inclusion.

For a set of variables $V \subseteq \variables$, we denote
$\valuations{V} \bydef{=} \{\val{}{}: V \rightarrow \data\}$
the set of \emph{valuations} of the variables in $V$.
Valuations extend canonically to expressions.  Thus
$\val{}{}(x)$ and $\val{}{}(e)$ denote, respectively, the
values of the variable $x$ and the expression $e$ under the
valuation $\val{}{}$.
%
\addSB{
%
It is important to observe that in practice, and in the examples of
this paper, all the variables are sorted, meaning that their values
can only range over given subsets of $\data$, such as $\sB$ or
$\sB^<$.  Hence, applied to expressions, valuations should be
interpreted as partial functions $\exprs{V} \rightharpoonup
\data$\footnote{
%

%
}
%
}

For a valuation $\val{}{} \in
\valuations{V}$ and an assignment $(x_i := e_i)^{i \in I} \in
\assigns{V}$, we denote $\val{}{}[(x_i := e_i)^{i \in I}]$ the
valuation defined by
\[
\val{}{}\bigl[(x_i := e_i)^{i \in I}\bigr](x) \bydef{=}
\begin{cases}
  \val{}{}(x), & \text{if } x \not\in x_i^{i \in I}\,,\\
  \val{}{}(e_i), & \text{if } x = x_i, \text{for some } i \in I\,.
\end{cases}
\]
For two valuations $\val{1}{}, \val{2}{} : V \rightarrow
\data$, we denote \[\valdiff{\val{1}{}}{\val{2}{}} \bydef{=}
\bsetdef{v \in V}{\val{1}{}(v) \neq \val{2}{}(v)}\] the set of
variables that are assigned different values by the two
valuations.  As usual, we write $\val{1}{} \order \val{2}{}$
iff $\val{1}{}(x) \order \val{2}{}(x)$, for all $x \in V$.  An
expression $e \in \exprs{V}$ is called \emph{monotonic} if, for
any two valuations $\val{1}{}, \val{2}{} \in \valuations{V}$,
$\val{1}{} \order \val{2}{}$ implies $\val{1}{}(e) \order
\val{2}{}(e)$.
%
Similarly, an assignment $(x_i := e_i)^{i \in I} \in \assigns{V}$ is
called monotonic if all expressions $e_i^{i \in I}$ are monotonic.
We will denote $\guards{V} \subset \boolexprs{V}$, $\monexprs{V}
\subset \exprs{V}$ and $\updates{V} \subset \assigns{V}$ the sets of
monotonic Boolean and generic expressions and assignments,
respectively.
%

%****************************************************************
%****************************************************************

\section{The theory of architectures with data}
\label{secn:archs}

%****************************************************************

\subsection{Components and composition}
\label{secn:components}

\noteSBin{%
To guarantee preservation of properties guards and expressions
have to be \emph{monotonic}.  In the case of a trivial order, \ie
when nothing is comparable, \emph{any guard or expression is
  monotonic}.  Thus, this is not really a constraint!

Do we put this in the definition of expressions and guards or in
the definitions of components and interaction models?
}

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $(Q, q^0, V, \val{0}{}, P,
  \export, \goesto{})$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    $\export : P \rightarrow 2^V$ 
    the \emph{set of variables exported by each port},
  \item $\goesto{}\, \subseteq
    Q \times 2^P \times \guards{V} \times \updates{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports
    $\emptyset \neq a \subseteq P$, a monotonic Boolean \emph{guard} $g \in
    \guards{V}$ and a monotonic \emph{update assignment} \hl{$u \in
    \updates{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, only exported variables, \ie those in the set
  $\bigcup_{p \in P}\export(p)$, belong to the component interface and
  can be used for the interactions with other components (see
  \defn{im}).  However, in the context of this paper we can simplify
  by omitting this separation.
%
  }
%
  Notations $q \goesto{a, g, u} q'$ and $q \goesto{a, g, u}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{BIPspec-ArchFailureTimerMax-v4}
  \caption{The BIP specification of the Failure Monitor architecture}
  \label{fig:schema:ArchFailure:BIP}
\end{figure}

In this paper, we will use a running example, which is a refined
version of the Failure Monitor architecture
from~\cite{CubETH-case-study}.  Although \fig{schema:ArchFailure:BIP}
shows the full definition of this architecture, we will explain its
various elements progressively, as we define the corresponding
notions.

The model in \fig{schema:ArchFailure:BIP} has two components:
%% The
%% component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
%% it only has its interface defined: $(\emptyset, \{\PortFinish,
%% \PortResume, \PortFail\}, \export[\emptyset])$, with
%% $\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
%% \PortResume, \PortFail\}$.  
%% The other two components,
%
$\NameTimer$ and $\NameCtrl$, with interfaces
\[\bigl(
\{t, z\},
\{\PortTick, \PortCancel, \PortTO\},
\export[\NameTimer]
\bigr)\,,
\qquad
\export[\NameTimer](p) =
\begin{cases}
  \{t, z\}, & \text{if } p = \PortTick,\\
  \emptyset, & \text{otherwise}
\end{cases}
\]
and
\[\bigl(
\{\mathit{zone}\},
\{\PortReset, \PortFail, \PortResume, \PortTO\},
\export[\NameCtrl]
\bigr)\,,
\quad
\export[\NameCtrl](p) =
\begin{cases}
  \{\mathit{zone}\}, & \text{if } p = \PortFail,\\
  \emptyset, & \text{otherwise}
\end{cases}
\]
%
respectively. %% , where $\export[\NameCtrl](\PortFail) =
%% \{\mathit{zone}\}$ and $\export[\NameCtrl](p) = \emptyset$, for all
%% other ports $p$; $\export[\NameTimer](\PortTick) = \{t, z\}$ and
%% $\export[\NameTimer](p) = \emptyset$, for all other ports $p$.
Variable $t$ is implicitly assumed to be an integer.  Variables $z$
and $\mathit{zone}$ are implicitly assumed to be integer intervals of
the form $[l,u]$.  For both types, we also implicitly assume the
natural ordering: increasing integers and interval inclusion.

%% are \emph{coordinators}\mdash they are specified
%% completely, including the states and transitions defining their
%% \emph{behaviour}.
Component \emph{behaviour} is defined by states ($\{\cstate{0},
\cstate{1}, \cstate{2}\}$, for $\NameCtrl$; $\{\tstate{0}\}$, for
$\NameTimer$) and transitions.
%
Transitions are labelled with the ports of the
corresponding components, Boolean guards and update assignments on local
variables.  For instance, the loop transition $\tstate{0}
\goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{0}$ in the
$\NameTimer$ component is labeled by the port $\PortTick$, it can be
fired only when the current value of the local variable $t$ is less
than $z.u$.
%% , where $z = [l, u]$ is a local variable representing an
%% interval
Upon firing, this transition increments the value of $t$ by $1$.  The
guards and update assignments of the transitions of $\NameCtrl$ are
omitted.  The default values are the constant predicate $\true$ (true) and
the empty assignment {\noop}, respectively.  The initial states
\tstate{0} and \cstate{0}, and valuations $\val{0}{\NameTimer} = \{t
\mapsto 0, z \mapsto \topinterval\}$ and $\val{0}{\NameCtrl} =
\{\mathit{zone} \mapsto [\mathrm{Min}, \mathrm{Max}]\}$ are shown by
incoming arrows $\goesto{\TimerInit[]} \tstate{0}$ and
$\goesto{\IntervalInit[]} \cstate{0}$, respectively.\footnote{%
%
  We use $\topinterval$ to denote the top interval in the inclusion
  order, \ie $(-\infty, +\infty)$.
%
} The constants $\mathrm{Min}$ and $\mathrm{Max}$ are the parameters
of the architecture.
%\noteLH{It looks like Max is in zone, I do not understand}

\todoSBin{Speak of guard monotonicity}

\begin{definition}[Component semantics]
  \label{defn:comp:semantics}
  The \emph{open semantics} of a component $B = \tuple{Q, q^0, V,
  \val{0}{}, P, \export, \goesto{}}$ is given by the LTS denoted $\semopen{B} = \tuple{S,
  s^0, \goesto{}}$, where $S = Q \times \valuations{V}$, $s^0 =
  \tuple{q^0, \val{0}{}}$ and $\goesto{}$ is the minimal transition
  relation satisfying the rule
  %
  \begin{equation}
    \label{eq:comp:semantics}
    \infer{
      q \goesto{a, g, u} q'
      \and
      \val{}{} \models g
      \and
      \val[\primeit]{}{} = \tilde{\val{}{}}[u]
      \and
      \valdiff{\val{}{}}{\tilde{\val{}{}}} \subseteq \export(a)
    }{
      (q, \val{}{}) \goesto{a,\tilde{\val{}{}}} (q', \val[\primeit]{}{})
    }
    \,.
  \end{equation}
  %
  \todoSB{Do we actually need this?}\todoLH{I agree: I do not think we need this}
  \hl{The \emph{closed semantics} of $B$ is given by the LTS denoted
  $\semclosed{B} \bydef{=}\reachable{\semopen{B}}$, comprising
  only the reachable states of $\semopen{B}$.}
\end{definition}

The use of the intermediate valuation $\tilde{\val{}{}}$ in the
conclusion and the third premise of rule \eq{comp:semantics}
allows some of the variables to get new values before the
transition is actually fired.  Thus the component is \emph{open}
to the exchange of data with its environment.  However, the
fourth premise in \eq{comp:semantics} restricts the set of
variables, which can be affected by such data transfer, to those
that are exported through the ports participating in the
interaction.

\begin{definition}[Interaction model]
  \label{defn:im}
  Consider a finite set of component interfaces $(V_i, P_i,
  \export[i])^{i \in I}$, such that all $P_i$ and all $V_i$ are
  pairwise disjoint, \ie $\forall i \neq j,\ P_i \cap P_j = V_i \cap
  V_j = \emptyset$.  Let $P = \bigcup_{i \in I} P_i$, $V = \bigcup_{i
    \in I} V_i$ and $\export : P \rightarrow 2^V$ such that, for any
  $p \in P_i$, $\export(p) = \export[i](p)$. 
%
  %% \noteSB{Updated this}\hl{and $V_a \bydef{=} \bigcup_{i \in
  %%   \supp{a}} \export[i](a \cap P_i)$ the set of component
  %% variables exported through $a$.}

  An \emph{interaction model over $(V, P, \export)$} is a set $\Gamma
  \subseteq 2^P \times \guards{V} \times \updates{V}$, such that,
  for any interaction $(a, g, u) \in \Gamma$, we have
  $g \in \guards{\export(a)}$ and $e \in \updates{\export(a)}$.\footnote{%
%
    Notice that this definition allows $(\emptyset, \true,
    \noop)$ and $(\emptyset, \false, \noop)$ to be included in
    $\Gamma$.
%
  }
  %% We call the set of ports $P$ the \emph{domain} of the
  %% interaction model.
\end{definition}

%% \noteLH{supp should go outside the definition}
When speaking of a set of components or interfaces, we will always
assume that it satisfies the disjointness assumption of \defn{im}.
%
We call the \emph{support} of a set of ports $a \subseteq P$, denoted
$\supp{a}$, either the set $\setdef{i \in I}{a \cap P_i \neq
  \emptyset}$ (for $P = \bigcup_{i=1}^n P_i$) or the set $\setdef{B
  \in \cB}{a \cap P_B \neq \emptyset}$ (for $P = \bigcup_{B \in \cB}
P_B$).  The precise meaning of this notation will always be clear from
the context.  Intuitively, for an interaction $(a, g, u)$, $\supp{a}$
denotes the set of the participating components.

\begin{definition}[Composition]
  \label{defn:im:sem}
  The \emph{composition} of a finite set of components $\cB = \tuple{Q_i,
  q^0_i, V_i, \val{0}{i}, P_i, \export[i], \goesto{}}^{i \in I}$ with
  the interaction model $\Gamma$ over $\tuple{V, P, \export}$ (as in \defn{im})
  is the component $\Gamma(\cB) = \tuple{Q, q^0, V, \val{0}{}, P,
    \export, \goesto{}}$, where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:sem}
%
    \infer{    
      (a, g, u) \in \Gamma
      \and
      a \neq \emptyset
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, u_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      g' = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      u' = u; u_i^{i \in \supp{a}}
    }{
      (q_i)^{i \in I} \goesto{a, g', u'} (q_i')^{i \in I}
    }\,.
  \end{gather}
  \todoSBin{Add the intuitive explanation}
\end{definition}

Specifying interaction models as sets of sets of ports is not
practical due to their potentially exponential size.  An algebra of
connectors was introduced in \cite{BliSif08-acp-tc} in order to
structure interactions in BIP models.

Connectors are hierarchical, tree-like structures with component ports
at the leaves.  They define sets of interactions, based on the
attributes of the connected ports, which may be either \emph{trigger}
(triangles in \fig{schema:ArchFailure:BIP}) or \emph{synchron}
(bullets in \fig{schema:ArchFailure:BIP}).
%
If all sub-connectors of a connector are synchrons, then an
interaction may be executed by the connector only if each subconnector
can contribute.
%
If at least one of the sub-connectors is a trigger, then any
interaction consisting of contributions of any set of sub-connectors,
\emph{involving at least one of the triggers}, can be executed.
The composition semantics of BIP systems consists in
firing exactly one interaction, enabled through at least one of the
top-level connectors, at each execution round.

\todoSB{Update this}
\hl{For instance, in \fig{schema:ArchFailure:BIP}, the two ports
$\NameTimer.\PortStart$ and $\NameCtrl.\PortFail$ are always
synchronised, since they belong to exactly one binary sub-connector,
where they are both synchrons.  In particular, this means that
whenever the transition \cstate{0} \goesto{\PortFail} \cstate{1} is
fired, so is the transition \tstate{0} \goesto{\PortStart,
  \TimerInit[]} \tstate{1}, initialising the timer.  The binary
connector $\NameTimer.\PortStart \twosynch
\NameCtrl.\PortFail$ is a sub-connector of a hierarchical connector,
where the port $\NameOprnd.\PortFail$ is a trigger.  Thus, the above
interaction can only happen together with $\NameOprnd.\PortFail$,
forming a ternary interaction.  On the contrary, being a trigger, the
port $\NameOprnd.\PortFail$ can fire alone, forming a singleton
interaction.}

\todoSBin{Speak of data transfer}

\todoSBin{Provide the interaction model}

\begin{definition}[Priority model]
  \label{defn:priority}
  A \emph{priority model} over a component interface $(V, P,
  \export)$\footnote{%
%
    Notice that the export function $\export$ has no impact on the
    priority model definition.
%
  } is a strict partial order on interactions: $\pi
  \subseteq \bigl(2^P \times \guards{V} \times \updates{V} \bigr)^2$.

  We write $\alpha \prec_\pi \beta$ for $(\alpha, \beta) \in \pi$ and 
  denote $\filter{\alpha} \bydef{=} \bsetdef{\beta \in (2^P \times
    \guards{V} \times \updates{V})}{\alpha \prec_\pi \beta}$ the set of
  interactions that have higher priority than $\alpha$.
\end{definition}

\begin{definition}[Application of priority]
  \label{defn:priority:sem}
  The application of a priority model $\pi$ over $(V, P, \export)$
  to a component $B = \tuple{Q, q^0, V, \val{0}{}, P, \export, \goesto{}}$
  is the component $\pi(B) = \tuple{Q, q^0, V, \val{0}{}, P, \export,
  \goesto[\pi]{}}$, where $\goesto[\pi]{}$ is the minimal transition
  relation satisfying the rule
%
  \begin{equation}
    \label{eq:priority:sem}
%
    \infer{
      \alpha = (a, g, u)
      \and
      q \goesto{a, g, u} q'
      \and
      \textstyle
      g' = g \land \bigwedge_{(b, h, w) \in \filter[q]{\alpha}} \lnot h
    }{
      q \goesto[\pi]{a, g', u} q'
    }\,, 
  \end{equation}
%
  where $\filter[q]{\alpha} \bydef{=} %% \filter{\alpha} \cap
  %% \bsetdef{\beta \in (2^P \times \guards{V} \times \updates{V})}
  \setdef{\beta \in \filter{\alpha}}{q \goesto{\beta}}$.
  \todoSBin{Add the intuitive explanation}
\end{definition}

In this paper, we will only consider the so-called \emph{maximal
  progress} priority $\mu$, where $(a, g, u) \prec_\mu (b, h, w)$ iff $a
\varsubsetneq b$ (\ie $a \subseteq b$ and $a \neq b$).  In
\secn{conclusion}, we discuss the possibility of generalising the
obtained results to other priority models.

%% Finally, \emph{priorities}\mdash defined by a strict partial order on
%% the set of possible interactions\mdash narrow the choice among the
%% enabled interactions at any given round.  The default priority is the
%% so-called \emph{maximal progress}, whereby among any two interactions
%% $a \subset b$ (as sets of ports), $b$ has higher priority than $a$.
Going back to our running example in \fig{schema:ArchFailure:BIP}, under the maximal progress priority, the
port $\NameOprnd.\PortFail$ will never fire alone in a global state,
where both $\NameTimer.\PortStart$ and $\NameCtrl.\PortFail$ are
enabled.

%% Adding priorities $\NameOprnd.\PortFinish < \NameCtrl.\PortReset$ and
%% $\NameOprnd.\PortFinish < \NameOprnd.\PortResume$, ensures that, after
%% a failure, the operation performed by $\NameOprnd$ cannot finish
%% unless it is explicitly resumed or the system is reset.

\begin{note}
  \label{rem:maxprogress}
  Below, we will assume that the maximal progress priority $\mu$ is
  applied after each composition with an interaction model.
\end{note}

%% %****************************************************************
%% \subsection{Running example: Failure Monitor architecture}
%% \label{secn:failure-monitor}

%\noteLH{Ludo ou simon: essayer de remonter la description generique en debut de section}

%****************************************************************
\subsection{Architectures}
\label{secn:archi}

\todoSBin{Sort out the following two paragraphs}

Contrary to standard BIP models, architectures comprise one or several
\emph{operand} components, whereof only the set of \emph{ports} is
given.  Here, the operand component is $\NameOprnd$ and its interface
consists of the ports $\PortFinish$, $\PortResume$ and $\PortFail$.

Application of the Failure Monitor architecture ensures that, whenever
a failure is registered in the operand component, the system will be
reset, unless a resumption is registered within $\mathrm{Max}$ time
units (more details in Sect.~\ref{section:resultOA}).

\noteLHin{should we say that always epsilon C is the epsilon of C?}

\begin{definition}[Architecture]
  \label{defn:arch}
  An \emph{architecture} is a tuple $A = \tuple{\cC, V_A, P_A,
    \export[A], \Gamma}$, where
%  
  \begin{itemize}
\item
  $\cC$ is a finite set of \emph{coordinating components}
  with pairwise disjoint sets of ports and variables, such that
  $\bigcup_{C \in \cC} P_C \subseteq P_A$ and
  $\bigcup_{C \in \cC} V_C \subseteq V_A$;
%
  ports in $P_A \setminus \bigcup_{C \in \cC} P_C$, which do not
  belong to any of the coordinating components are called
  \emph{dangling};

\item
  $\export[A] : P_A \rightarrow 2^{V_A}$ is an export function, such
  that $\export[A](p) = \export[C](p)$, for any $C \in \cC$ and $p \in
  P_C$ and $\export[A](p) \subseteq V_A \setminus \bigcup_{C \in \cC}
  V_C$ for any dangling port $p$; and

\item
  $\Gamma \subseteq 2^{P_A} \times \guards{V_A} \times \exprs{V_A}$
  is an interaction model over $(V_A, P_A, \export[A])$.
  \end{itemize}

\end{definition}

\begin{definition}[Application of an architecture]
  \label{defn:arch:application}
  Let $A = \tuple{\cC, V_A, P_A, \export[A], \Gamma}$ be an architecture and let $\cB$
  be a set of components, such that
%
  \begin{align}
    \bigcup_{B \in \cB} V_B \cap \bigcup_{C \in \cC} V_C = \emptyset\,,
    &&
    V_A \subseteq V \bydef{=} \bigcup_{B \in \cB \cup \cC} V_B\,,
    \\
    \bigcup_{B \in \cB} P_B \cap \bigcup_{C \in \cC} P_C = \emptyset\,,
    &&
    P_A \subseteq P \bydef{=} \bigcup_{B \in \cB \cup \cC} P_B\,,
  \end{align}
%
  \hl{and $\export[A](p) = V_A \cap \export[B](p)$, for any $B \in \cB$
  and $p \in P_A \cap P_B$.}
%
  The \emph{application of the architecture $A$ to the set of
  components $\cB$} is the component $ A(\cB) \bydef{=}
  \mu\bigl((\IMextend{\Gamma}{P})(\cC \cup \cB)\bigr)$, where
%
  $\IMextend{\Gamma}{P}$ is the \emph{extension} of the interaction
  model $\Gamma$ to the set of ports $P$, \ie the interaction model
  over $(V, P, \export)$, with
  \[
  \export(p) = 
  \begin{cases}
    \export[C](p), & \text{for all } C \in \cC, p \in P_C,
    \\
    \export[B](p), & \text{for all } B \in \cB, p \in P_B,
  \end{cases}
  \]
  defined by\noteLH{looks like this also depends on Pa}
%
  \noteSB{It does, indeed.  To be completely rigorous, I could replace
    this with something like $\Gamma\uparrow_{P_A}^P$ but this is
    quite heavy and the dependency is already implicit from the fact
    that $\Gamma$ is defined on $P_A$}
%
  \noteSB{Which one of \eq{im:extension} and \eq{im:extension1} is
    more intuitive?}
%
  \begin{align}
    \label{eq:im:extension}
    \IMextend{\Gamma}{P} &\bydef{=}
    \bsetdef{
      (a \cup a', g, e)
    }{
      (a, g, e) \in \Gamma, a' \subseteq P \setminus P_A
    }
    \\
    \label{eq:im:extension1}
    &= 
    \bsetdef{
      (a, g, e)
    }{
      a \subseteq P,
      (a \cap P_A, g, e) \in \Gamma
    }
  \end{align}
%
  and $\mu$ is the maximal progress priority model.
\end{definition}

\todoSBin{Define equivalence $\arequiv$?}

An architecture $A$ enforces coordination constraints on the
components in $\cB$.  The interface \hl{$(V_A, P_A, \export[A])$} of an
architecture $A$ contains all ports of the coordinating
components $\cC$ and some additional ports, which must belong to
the components in $\cB$.  In the application $A(\cB)$, the ports
belonging to $P_A$ can only participate in the interactions
defined by the interaction model $\Gamma$ of $A$.  Ports which do
not belong to $P_A$ are not restricted and can participate in any
interaction.  In particular, they can join the interactions in
$\Gamma$ (see \eq{im:extension}).  If the interface of the
architecture covers all ports of the system, \ie $P = P_A$, we
have $P\setminus P_A = \emptyset$ and the only interactions
allowed in $A(\cB)$ are those belonging to $\Gamma$.  Notice also
that the restrictions imposed by \defn{arch:application} on the
set of operand components $\cB$ ensure that an interaction in
$\IMextend{\Gamma}{P}$ can only refer to variables of the
participating components.
%
Finally, the definition of $\IMextend{\Gamma}{P}$ requires that
an interaction from $\Gamma$ be involved in every interaction
belonging to $\IMextend{\Gamma}{P}$.  To allow the ports from $P
\setminus P_A$ to be fired independently in $A(\cB)$, one must
have $(\emptyset, \true, \noop) \in \Gamma$.

\begin{definition}[Composition of architectures]
  \label{defn:arch:composition}
  Let $A_i = \tuple{\cC_i, V_{A_i}, P_{A_i}, \export[A_i], \Gamma_i}$, for $i = 1,2$,
  be two architectures.  The \emph{composition} of $A_1$ and
  $A_2$ is the architecture $A_1 \arcomp A_2 = \tuple{\cC_1 \cup \cC_2,
  V_{A_1} \cup V_{A_2}, P_{A_1} \cup P_{A_2}, \export[A_1] \cup \export[A_2], \Gamma}$, where
%
  \begin{equation}
    \label{eq:arch:composition}
    \Gamma = \bsetdef{
      (a, g^1 \land g^2, e^1 \expmix e^2) 
    }{
      (a \cap P_{A_i}, g^i, e^i) \in \Gamma_i,
      \text{ for } i = 1,2
    }
    \,.
  \end{equation}
\end{definition}

\begin{proposition}[Properties of $\arcomp$]
  \label{prop:arcomp:nice}
  Architecture composition $\arcomp$ is commutative and
  associative; it is idempotent if all coordinating components
  are deterministic; $A_{id} = \bigl(\emptyset, \emptyset,
  \emptyset, \emptyset, \{(\emptyset, \true, \noop)\}\bigr)$ is its neutral element, \ie for
  any architecture $A$, we have $A \oplus A_{id} \arequiv A$.
  Furthermore, for any component $B$, we have $A_{id}(B) = \mu(B)$.
\end{proposition}
%
\begin{proof}[Sketch of the proof]
  Commutativity and associativity follow from the corresponding
  properties of set union, Boolean conjunction and
  \todoSB{Check ``semi''}\hl{semilattice meet}.
%
\todoLH{Check when equivalence is defined}
  Suppose we have two architectures $A \arequiv A'$.  This does not
  necessarily mean that their sets of coordinating components
  coincide.  However, if all the involved coordinating components
  are deterministic, then, in any state of $(A \arcomp A')(\cB)$,
  both architectures will impose the same restrictions, enabling
  the same interactions between the coordinating and operand
  components.  Hence, we have $(A \arcomp A')(\cB) = A(\cB) =
  A'(\cB)$.  Since this holds for any set of components $\cB$, we
  conclude that $A \arcomp A' \arequiv A \arequiv A'$.
%
  The properties of $A_{id}$ follow immediately from the
  definitions of architecture application and composition.
\end{proof}

The last statement of \prop{arcomp:nice} highlights a subtle point:
$A_{id}$ is the neutral element of the $\arcomp$ operator\mdash it is
not the identity operator on components.

%****************************************************************
\subsection{Preservation of safety properties}
\label{secn:safety}

For a component $B$, we denote $S_{\semopen{B}}$,
$s^0_{\semopen{B}}$ the corresponding constituents of
$\semopen{B}$.

\begin{definition}[Safety properties]
  \label{defn:property}
  Let $B$ be a component.  A \emph{safety property} (below,
  simply \emph{property}) of $B$ is a state predicate $\Phi:
  S_{\semopen{B}} \rightarrow \sB$, such that $\bigl((q,
  \val{}{}) \models \Phi\bigr) \land (\val[\primeit]{}{} \order
  \val{}{})$ implies $(q, \val[\primeit]{}{}) \models \Phi$,
  where we write $(q, \val{}{}) \models \Phi$ iff $\Phi(q,
  \val{}{}) = \true$.  A property $\Phi$ is \emph{initial} if
  $s^0_{\semopen{B}} \models \Phi$.
  %% \todoSB{Do we need this?}\hl{;
  %% it is \emph{reachable} iff there exists a possibly empty path
  %% $s^0_{\semopen{B}} \goesto{a^1, \val{1}{}} s^1 \goesto{a^2,
  %%   \val{2}{}} \cdots \goesto{a^n, \val{n}{}} s^n$, such that
  %% $s^n \models \Phi$}.
\end{definition}

\todoSB{Cite the original paper somewhere.}
The main idea of our approach is that an architecture enforces
its characteristic property on the set of its operand components.
From this point of view, the set of coordinating components is
not relevant, neither are their states.  Thus, to talk about
properties enforced by architectures, we consider properties on
the unrestricted composition of the operand components as
formalized by the following definition.

\begin{definition}[Enforcing properties]
  \label{defn:impose}
  Let $A = \tuple{\cC, P_A, V_A, \export[A], \Gamma}$ be an architecture; let $\cB$
  be a set of components and $\Phi$ an initial property of
  their parallel composition $A_{id}(\cB)$ (see
  \prop{arcomp:nice}).  We say that \emph{$A$ enforces $\Phi$ on
    $\cB$} iff, for every state $s = (s_c, s_b)$ reachable in
  $\semopen{A(\cB)}$, with 
  $s_c \in \prod_{C \in \cC} S_{\semopen{C}}$ and
  $s_b \in \prod_{B \in \cB} S_{\semopen{B}}$,
  we have $s_b \models \Phi$.
\end{definition}

According to the above definition, when we say that an
architecture enforces some property $\Phi$, it is implicitly
assumed that $\Phi$ is initial for the coordinated components.
Below, we omit mentioning this explicitly.

\begin{definition}[Upwards compatibility]
  \label{defn:up:compat}
  \todoSB{Summarise in English and move to the appendix}
%
  Consider two architectures $A_i = \tuple{\cC_i, V_{A_i}, P_{A_i},
  \export[A_i], \Gamma_i}$, for $i = 1,2$.
%
  We say that $A_1$ and $A_2$ are \emph{upwards compatible} iff, for
  any two interactions $(a_i, g_i, e_i) \in \Gamma_i$, such that $a_1
  \cap P_{A_2} = a_2 \cap P_{A_1}$ (\ie they agree on the common ports
  of the two architectures), we have 1)~for any interaction $(b_1,
  h_1, f_1) \in \Gamma_1$, such that $a_1 \varsubsetneq b_1$, there
  exists an interaction $(b_2, h_2, f_2) \in \Gamma_2$, such that $a_2
  \subseteq b_2$, $b_1 \cap P_{A_2} = b_2 \cap P_{A_1}$ and the 
  implication  
%
  \begin{multline*}
    \left(
    \exists (q^1_C)^{C \in \supp{b_1} \cap \cC_1}:
    \exists (q^2_C)^{C \in \supp{a_2} \cap \cC_2}:
    \bigwedge_{
      \begin{multlined}
        \scriptstyle
        C \in \supp{b_1} \cap \cC_1 \cap
        \\
        \scriptstyle
        \supp{a_2} \cap \cC_2
      \end{multlined}
    }
    q^1_C = q^2_C
    \ \land
    \right.
    \\
    \begin{aligned}
      \exists \bigl(q^1_C \goesto{b_1 \cap P_C, g^{b_1}_C, f^{b_1}_C}\bigr)^{C \in \supp{b_1} \cap \cC_1}:
      & \quad \left.
      h_1 \land \bigwedge_{C \in \supp{b_1} \cap \cC_1} g^{b_1}_C
      \ \land \right.
      \\
      \exists \bigl(q^1_C \goesto{a_1 \cap P_C, g^{a_1}_C, e^{b_1}_C}\bigr)^{C \in \supp{a_1} \cap \cC_1}:
      & \quad \left.
      g_1 \land \bigwedge_{C \in \supp{a_1} \cap \cC_1} g^{a_1}_C
      \ \land \right.
      \\
      \exists \bigl(q^2_C \goesto{a_2 \cap P_C, g^{a_2}_C, e^{a_2}_C}\bigr)^{C \in \supp{a_2} \cap \cC_2}:
      & \quad
      \left.
      g_2 \land \bigwedge_{C \in \supp{a_2} \cap \cC_2} g^{a_2}_C
      \right) \implies 
    \end{aligned}
  \end{multline*}
%
  \begin{multline*}
    \left(
    \exists (q^2_C)^{C \in (\supp{b_2}\setminus\supp{a_2}) \cap \cC_2}:
            {\color{white}\bigwedge_{C \in \cC_1}}\right. % Fake bigwedge to make the parenthesis the right size
            \\
            \left.
            \exists \bigl(q^2_C \goesto{b_2 \cap P_C, g^{b_2}_C, f^{b_2}_C}\bigr)^{C \in \supp{b_2} \cap \cC_2}:
            h_2 \land
            \bigwedge_{C \in \supp{b_2} \cap \cC_2} g^{b_2}_C
            \right)
  \end{multline*}
%  
  holds for all valuations $\val{}{1} : V_{A_1} \rightarrow \data$
  and $\val{}{2} : V_{A_2} \rightarrow \data$; and 2)~symmetrically
  for any interaction $(b_2, h_2, f_2) \in \Gamma_2$, such that $a_2
  \varsubsetneq b_2$.
\end{definition}

\begin{theorem}[Preserving enforced properties]
  \label{thm:combining}
  Let $\cB$ be a set of components; let $A_i = \tuple{\cC_i, V_{A_i},
  P_{A_i}, \export[A_i], \Gamma_i}$, for $i = 1,2$, be two upwards compatible architectures
  enforcing on $\cB$ the properties $\Phi_1$ and $\Phi_2$
  respectively.  The composition $A_1 \arcomp A_2$ enforces on
  $\cB$ the property $\Phi_1 \land \Phi_2$.
\end{theorem}

%****************************************************************
%****************************************************************

\section{Open pNets}
\label{secn:pNets}
\todoLHin{rewrite after uniformisation}
pNets are tree-like structures, where the leaves are either
\emph{parameterised labelled transition systems (pLTSs)}, expressing the
behaviour of basic processes, or \emph{holes}, used as placeholders
for unknown processes, of which we only specify the set of possible
actions, this set is named the \emph{sort}.
Nodes of the tree (pNet nodes) are synchronising artifacts, using a
set of \emph{synchronisation vectors} that express the possible
synchronisation between the parameterised actions of a subset of the
sub-trees.

\subsection{The (open) pNets Core Model}
\label{section:pNets}

A pLTS is a labelled transition system with variables. Variables can be
used to characterise a state, as part of actions, guards, and
assignments. Without loss of generality and to simplify the formalisation, we suppose 
here that 
variables are local to each 
state: each state has its set of variables disjoint from the others. Transmitting 
variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
Note that we make no assumption on the finiteness of the set of states, nor
on the finite branching of the transition relation.

We first define the set of actions a pLTS can use, let $a$
range over action labels. Action terms are  of the form $\alpha=a((?x_i)^{i \in I}, 
\Expr_j^{j \in J})$, where $(?x_i)^{i \in I}$ are input variables, $\Expr_j^{j \in J})$ 
are expressions.

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of states, $\vars(s)$ denotes the set of associated variables;
\item[$\bullet$]
$s_0 \in S$ is the initial state;
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
% SB: I removed a new paragraph here, ok?
$\langle \alpha,~g,~u\rangle$,
where $\alpha$ is a parameterised action, $\alpha \in\actions{\variables}$; 
$g\in\boolexprs{\variables}$ is a guard over variables of the source state and the 
action, and $u\in\assigns{\variables}$
assigns updated value for variables in the destination state. 
If 
$s \xrightarrow{\langle \alpha,~g,~u\rangle} s'\in \to $ then 
% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! \vars(s)$, 
		$\vars(g)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
		$\vars(\codom(u))\!\subseteq\! \vars(s)\cup\vars(\alpha)\land 
		\vars(\dom(u))\!\in\!\vars(s')$. %,  and $s= s'\Rightarrow J=\emptyset$.

\end{itemize}
\end{definition}
\noteLH{is dom and codom clear?}
\noteSB{Not sure: I understand, because I know what is meant, but this is not a canonical use, so should be explained. Will be solved if we roll back to \mbox{$(?x_i := e_i)^{i \in I}$} or \mbox{$\overline{?x := e}$}.}
Now we define
pNet nodes, as constructors for hierarchical behavioural structures.
A pNet has a set of sub-pNets that can be either pNets or pLTSs, and a
set of holes, playing the role of process parameters.

A composite pNet consists of a set of sub-pNets exposing
a set of actions, each of them synchronising actions in each of
the sub-pNets. The synchronisation between global actions and
internal actions is given by  \emph{synchronisation vectors}: a
synchronisation vector synchronises one or several internal actions, and
exposes a single resulting global action.
Actions involved at the pNet level (in the synchronisation vectors) do
not have input variables, \ie they 
have the form $a(\Expr_j^{j \in J})$.

\begin{definition}[pNets]\label{def-pnets}
A pNet is a hierarchical structure where leaves are pLTSs and holes\footnote{Typing can 
be added to pNets by defining and checking their interfaces, called 
sort~\cite{HMZ-FORTE2016}}:
$\pNet\triangleq \pLTS~|~\mylangle \pNet_i^{i\in I}, J, \symb{SV}_k^{k\in 
K}\myrangle$
where
\begin{itemize}
\item[$\bullet$] $\pNet_i^{i\in I}$ is the family of sub-pNets;
%  $\pNet_i^{i\in I}$ is a family of sub-pNets where $I\in\I_\P$ is the set over which 
%sub-pNets are indexed.

\item[$\bullet$] $J$ is a set of indexes, called \emph{holes}.
$I$ and $J$ are \emph{disjoint}: $I\!\cap\! J=\emptyset$,  $I\!\cup\! J\neq\emptyset$
%\item[$\bullet$] $\Sort_j \subseteq \AlgAS$ is a set of action terms, denoting the 
%\emph{sort} of
%hole $j$.

\item[$\bullet$] $\symb{SV}_k^{k\in K}$ is a set of
  synchronisation vectors. % (\hl{$K\in\I_\P$}). 
$\forall k\!\in\! K,
  \symb{SV}_k\!=\SV{\!\alpha_{l}^{l\in I_k \uplus J_k}}{\alpha'_k}{g_k}$, where
  $\alpha'_k\in \actions{\variables}$, $I_k\subseteq I$, $J_k\subseteq J$, and 
  $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\alpha'_k$. The Boolean expression $g_k $, such that $\vars(g_k)\subseteq \bigcup_{l\in 
I_k\uplus J_k}{\vars({\alpha_l})}$, is a guard associated to the vector.

\end{itemize}
We define the set of holes and leaves of a pNet as follows:
  \begin{itemize}
\item
The set of holes $\Holes(\pNet_{i_1})$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets (we suppose those indexes 
disjoints).
%
%defined inductively; the sets of holes
%in a pNet node and its subnets are all disjoint:
%  \[\begin{array}{l}
%\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
%\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
%=J\uplus{\displaystyle \bigcup_{i\in 
%I}\Holes(\pNet_i)}\\
%\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
%\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
%\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
%\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(\pNet)= \mylangle \pNet_i \myrangle^{i \in L}$.
is said to be \emph{open}.
\end{itemize}
A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
is said to be \emph{open}.
\end{definition}

\subsection{Operational Semantics for Open pNets}
\label{section:op-semantics}

\noteLH{a pred is a g a post is a e}
The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes the actions of several holes, 
the transition occurs if some predicates hold, and can involve a 
set of state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a
	set $J$ of holes  and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{	\beta_j^{j\in J}, \Pred, \Post}
	{s \OTarrow {{\alpha}}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and $\beta_j\in\actions{\variables}$
        is an action of the hole $j$. $v$ is a variable denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states  $\beta_j$, $s$, $\alpha$. $\Post\in 
	\assigns{\variables}$\ is 
	a set of assignments that are the effects of the transition.
Open transitions are identified modulo logical equivalence on their predicate.
\end{definition}
%\begin{definition}[Open transitions]
%	\label{def:OpenTransitions}
%	An \emph{open transition} over a set \noteSB{In the architecture section, I say that 
%we omit indices on $\goesto{}$ and that they are always clear from the 
%context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
%	I}$} of LTSs, a
%	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
%	a structure of the form:	
%	\begin{mathpar}
%	\inferrule*[myfraction=\reddottedrule]
%	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
%		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
%	{s \OTarrow {v}s'}
%	\end{mathpar}
%	Where $s, s'\in\mathcal{S}$ and for all
%        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
%	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and 
%$\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort 
%$\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
%        is a transition of the hole $j$}, for any action $b_j$ in the
%        sort $\Sort_j$.  
%        of this open transition. \Pred\ is a predicate 
%	over the different variables of the
%	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
%	hold \emph{after the open transition}, they are represented as a substitution of the 
%	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
%	would like to just put $\Post_=e$}
%	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
%	variables of the open transition.
%\end{definition}

\begin{example}\emph{An open-transition.}
  \label{OT:enable-composed}
  The \texttt{EnableCompL} pNet of Fig. \ref{schema:enable-composed} has 2 controllers 
  and 2 holes. One of its possible open-transition is:

 \smallskip
 $  OT_2  = \openrule{ 
							(P\mapsto \delta(x4), Q\mapsto accept(x4))               
%                            \xrightarrow{\delta(x4)}_P ~~
%                            \xrightarrow{accept(x4)}_Q 
                      }
  %  {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}
    {A1_0 \OTarrow{\underline{\delta(x4)}} A1_1}
  $\noteSB{The underline notation did not appear until here.}
\end{example}

It is important to understand the difference between the red dotted rule and a normal 
inference rule. They correspond to two different logical levels.
 The open transition is itself a logical implication, but using a simple logic (with the 
 expressive power that is given to the predicate and necessary include logical 
 propositions and equality). We differentiate classical (black) inference rules that use 
 an expressive logic and are paper rules from open transition rules that use a simpler 
 logic, 
 will be embedded into automata transitions, and could typically be handled in a 
 mechanical way.

\begin{definition}[Open automaton]
	\label{def:open-automaton}
	An \emph{open automaton} is a structure\\ $A =
	<J,\mathcal{S},s_0,\mathcal{T}>$ where:
	\begin{itemize}
		\item[$\bullet$]   $J$ is a  set of indices,
		\item[$\bullet$]   $\mathcal{S}$ is a set of states and $s_0$ an initial state
		among $\mathcal{S}$,
		\item[$\bullet$] $\mathcal{T}$ is a set of open transitions and for each
		$t\in \mathcal{T}$ there exist  $J'$ with  $J'
		\subseteq J$, such that $t$ is an open transition over  $J'$,
		and  $\mathcal{S}$.

	\end{itemize}
\end{definition}

%
%Then the semantics of a pNet is characterized by a set of {\em open
%transitions}, where the hypotheses on process parameters are
%replaced by 1) transitions of the pLTSs at the leaves, and 2) formal
%hypotheses on the transitions of the holes. A {\em predicate} is used
%to relate the parameters and names appearing in the actions of the
%leaves and the holes involved in the rules, but also appearing in  the resulting action.

We define states of open pNets as tuples of states, where we denote tuples
in structured states as $\triangleleft\ldots\triangleright$ for distinguishing tuple 
states from other tuples.
\begin{definition}[States of open pNets]\label{def-states}
  A state of an open pNet is a tuple (not necessarily finite) of the
  states of its leaves.

  For any pNet p, let $\Leaves(p) = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be 
  the set of pLTS at its leaves,
  then $States(p) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
A pLTS being its own single leave:
  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S\}$.

The initial state is defined as:
$InitState(p) = \triangleleft {{s_i}_0}^{i\in L}  \triangleright$.
\end{definition}

%% \begin{example} \emph{State of a pNet}
%%   The states of pNet \texttt{EnableCompL} are:
%%   $\triangleleft 00 \triangleright, \triangleleft 10 \triangleright, \triangleleft 11 
%%\triangleright$
%% \end{example}

\paragraph{Predicates:}
Let
$\mylangle\overline{\pNet},\overline{\Sort},\overline{\symb{SV}}\myrangle$
be a pNet. Consider a synchronisation vector $\symb{SV}\in \overline{\symb{SV}}$. We 
define a
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\begin{multline}
\Pred_{\symb{sv}}(\SV{{(\alpha'_i)}^{i\in I}, {(\beta'_j)}^{j\in J}} 
{\alpha'} {g}, \alpha_i^{i\in I}, \beta_j^{j\in J}, \alpha)\Leftrightarrow \\
\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, \beta_j=\beta'_j \land 
\alpha=\alpha' 
\land g
\end{multline}

Somehow, this predicate entails a verification of satisfiability in the sense that if the 
predicate $\Pred_{\symb{sv}}$ is not satisfiable, then the transition associated with the 
synchronisation will not occur in the considered state.

In any other case (if the action families do not match or if there is no valuation of
variables such that the above formula can be ensured) the predicate is undefined.

This definition is not constructive but it is easy to build the predicate constructively
by brute-force unification of the sub-pNets
actions with the corresponding vector actions, possibly followed by a simplification
step.

We build the semantics of open pNets as an open automaton where LTSs are the pLTSs at
the leaves of the pNet structure, and the states are given by 
Definition~\ref{def-states}. The open transitions first
 project the global state into states of the leaves, then apply
pLTS transitions on these states, and compose them with the sort of the holes. %The pNet
%structure does not appear in the open-automaton, only the
%set of Holes and the set of Leaves.
The semantics    instantiates \emph{fresh} variables, additionally, for an action 
$\alpha$, $\fresh(\alpha)$ means all variables in $\alpha$ are fresh.

\begin{definition}[Operational semantics of open pNets]
	\label{def:operationalSemantics}
	The semantics of a pNet $p$ is an open automaton $A = 
	<Holes(p),States(p),InitState(p),
	\mathcal{T}>$ where $\mathcal{T}$ is the smallest set of open transitions		
	satisfying the rules below:
%	\begin{itemize}
%		\item $J$ is the set of holes: $Holes(p)= J$. 
		%  \item $\overline{L}^L = Leaves(p), \overline{H}^J = Holes(p)$
%		\item ${\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
%		\item $\mathcal{T}$ is the smallest set of open transitions		satisfying the 
%rules below:
%	\end{itemize}

	%% \TODO{ We should be careful here: after (re) reading "Huimin
	%% 	Lin, 'Symbolic Transition Systems with Assignements', Concur'96" I
	%% 	think handling assignments is not trivial, even for comparisons of pLTSs. }

	The rule for a pLTS  checks that the guard 
	is verified and transforms assignments into post-conditions:

	\begin{description}
		\item[{\bf Tr1:}]
	$\inferrule
		{ s \xrightarrow{\langle \alpha,~g,~u\rangle} s'\in \to  }
		{ \mylangle  S,s_0, \to \myrangle
			\models
			\openrule
			{\emptyset ,
			g,u}
			{\ostate{s} \OTarrow{\alpha} \ostate{s'}}
		}
		$
	\end{description}
%	Note that this note is greatly simplified by the fact that variables are local to 
%	thread; introducing global state variables or accepting loops to the same 
%	state would 
%	require to reason 
%	on the scope of 
%	each variables, and to introduce additional variables to handle the several occurence 
%	of the same pLTS variable in the predicates. Indeed the constraints on pLTS 
%	transitions 
%	ensure that the same variable never appears both on the left and on the right of the 
%	equations of a predicate.

	The second rule deals with pNet nodes: for each possible
	synchronisation vector (of index $k$) applicable to the rule subject, the premisses
	include one {\em open transition} for each sub-pNet involved , one possible
	{\em action} for each Hole involved, and the predicate relating these
	with the resulting action of the vector. The sub-pNets involved are split between two 
	sets, $I_2$ for subnets that are pLTSs, and $I_1$ for the others, $J$ is the set of 
	holes involved in the transition\footnote{Formally, if $SV_k \!=\! \SV{\alpha_m^{m 
	\in M}}{\alpha'}{g}$ is a synchronisation vector  of \pNet\  then $J=M\cap 
	\Holes(\pNet)$, $I_2=M\cap \Leaves(\pNet)$,  $I_1=M\setminus J \setminus 
	I_2$}.

	\begin{description}
		\item[{\bf Tr2:}]
	\end{description}

	\noindent
\begin{mathpar}
    \mprset {vskip=1ex}
\inferrule
    {
\Leaves(\mylangle \overline{\pNet}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{g} 
\\
\\     	
	\forall m\!\!\in\!\! I_1. {\pNet_m 
	\models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
  \qquad
\forall m\!\!\in\!\! I_2.		{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}} }\\\\
     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in (I_1\uplus I_2)}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,\alpha_m^{m\in (I_1\uplus I_2)},\beta_j^{j\in J},\alpha)\\ 
    		I' = \biguplus_{m\in I_1}\!\! I_m \uplus I_2
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
    }
    {\mylangle \overline{\pNet}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\openrule
    		{
    		{\beta_j}^{j\in J^\prime}, \Pred, \uplus_{m\in I_1\uplus I_2} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
\end{mathpar}

%OLD	\begin{description}
%		\item[{\bf Tr2:}]\noteSB{I do not understand why the second line $pNet_m 
%\models\dots$ is not subsumed by the first one?}
%	\end{description}
%	
%	\noindent
%    $\inferrule
%    {k\!\in\! K \\SV_k \!=\! \alpha_m^{m \in I\uplus J} \!\to\! 
%    \alpha' |g \\
%    	Leaves(\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in 
%    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
%    	\forall m\!\!\in\!\! I. 	
%    {\left(\begin{array}{l}
%	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
%    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
%    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
%    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
%    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
%		{ \pNet_m 
%    	 \models
%    	\inferrule*[myfraction=\reddottedrule]
%    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
%    	{\ostate{s_m} \OTarrow {v_m}
%    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
%\end{array}\right)}\\
%    	%\land
%    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
%     J' = \biguplus_{m\in I}\!\! J'_m \uplus J 	\\
%    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
%    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
%    		I' = \biguplus_{m\in I}\!\! I_m'
%    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
%    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
%    }
%    {\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
%    	\models
%    	{\inferrule*[myfraction=\reddottedrule]
%    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
%    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
%    		\Post_m}
%    		{\ostate{s_i^{i\in L}} \OTarrow {v}
%    			\ostate{(s_i^\prime)^{i\in L}}}
%    	}
%    }
%    $
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}

\end{definition}

%****************************************************************
%****************************************************************

\section{Encoding of architectures into open pNets}
\label{secn:encoding}

We will define the encoding of BIP architectures into pNets by
associating to each architecture \emph{and a partition of its dangling
  ports} the corresponding pNet.  Notice that the addition of a
partition of dangling ports is a matter of convenience, since one can
use a trivial partition comprising only one complete set of dangling
ports.

Let $A = \tuple{\cC, V_A, P_A, \export[A], \Gamma}$ be an
architecture.  For each $C \in \cC$, let $C = \tuple{Q_C, q^0_C, V_C,
  \val{0}{C}, P_C, \export[C], \goesto{}}$.  Finally, let $\partition
\subseteq 2^{P_A}$ be a partition of the set of dangling ports of $A$,
\ie $\bigcup_{D \in \partition} D = P_A \setminus \bigcup_{C \in \cC}
P_C$ and $D_1 \cap D_2 = \emptyset$, for any two distinct $D_1, D_2
\in \partition$.  In particular, as mentioned above, one can consider
the trivial partition $\partition \bydef{=} \{P_A \setminus \bigcup_{C
  \in \cC} P_C\}$.

Below, for the sake of clarity, we start by presenting in
\secn{enc:nopri} the encoding of $A$ into pNets without accounting for
any priority model (see \secn{components}).  Then, in
\secn{enc:maxprog}, we slightly modify this encoding to include the
implicitly assumed maximal progress priority.

%****************************************************************
\subsection{Encoding without the priority model}
\label{secn:enc:nopri}

Recall that $\Gamma$ is an interaction model over the interface
$\tuple{V_A, P_A, \export[A]}$, \ie these interface elements are
implicitly involved in the definition of $\Gamma$.

\paragraph{Encoding the coordinators} We define the
mapping $\nopri$ as $\nopri[A, \partition] \bydef{=}
\pNetTuple{(\nopri[C])^{C \in \cC}, \partition, \nopri[\Gamma]}$,
where, for each $C \in \cC$, the encoding $\nopri[C]$ is a pLTS
$\pNetTuple{S, s_0, \goesto[{\nopri[C]}]{}}$, such that
%
\begin{itemize}
\item $s_0 \not\in Q_C$ is fresh and $S = Q_C \cup \{s_0\}$,
\item $\vars(s) = V_C$, for all $s \in Q_C$, and $\vars(s_0) =
  \emptyset$,
\item the transition relation $\goesto[{\nopri[C]}]{}$ is defined as
  the union of three sets of transitions:
%
  \begin{multline}
    \label{eq:nopri:trans}
    \goesto[{\nopri[C]}]{} \bydef{=}
    \bsetdef{\tuple{s, a(\true), g, u, s'}}{
      \tuple{s, a, g, u, s'} \in \goesto{}}
    \\    
    \cup
    \bsetdef{\tuple{s, a(\false), \true, \noop, s}}{
      s \in Q_C, a \in P_C}
    \\
    \cup
    \bigl\{(s_0, \init, \true, \mathit{init}, q^0_C)\bigr\}
    \,,
  \end{multline}
%
  where $\mathit{init} = \bigl(x := \val{0}{C}(x)\bigr)^{x \in V_C}$.
\end{itemize}

Intuitively, we 1)~introduce a fresh initial state with an additional
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~define a pNet action algebra that associates, as action parameter, an
additional Boolean value to each port, 3)~create pNet transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 4)~create additional loop transitions marked by
$\false$.

The loop transitions marked by $\false$ will be used in the encoding
of connectors.  Each BIP connector can define several interactions,
\ie ports involved in a connector need not necessarily always
participate.  On the contrary, each action in a pNet synchronisation
vector must participate in the synchronisation.  To address this
difference, we use the classical~\cite{milner83-calculi} approach,
whereby non-participation of a port in an interaction is simulated
by an additional loop transition.

\Fig{enc:nopri} shows the encoding of the Failure Monitor
architecture, including the encodings of the two coordinators, \ie
$\nopri[\NameTimer]$ and $\nopri[\NameCtrl]$.  Notice that the
encoding in the figure is slightly optimised: some of the ports do not
have an associated Boolean value, nor the additional loop transitions.
\todoSB{Revise this if necessary}\hl{We will explicitate this optimisation once we have defined the
encoding of the interaction model.}

\begin{figure*}[t]
  \fbox{%
  \includegraphics[width=\textwidth]{BIPspec-ArchFailureTimerMax-v4-no-pri}}
  \caption{Encoding of the Failure Monitor architecture
    (\fig{schema:ArchFailure:BIP}) as an open pNet (without the
    priority model) \hl{\bf (Error in the picture: First two positions in the synchronisation vectors must be transposed)}}
  \label{fig:enc:nopri}
\end{figure*}

\paragraph{Encoding the interaction model}
First of all, recall that $\nopri[A, \partition ] \bydef{=}
\pNetTuple{(\nopri[C])^{C \in \cC}, \partition, \nopri[\Gamma]}$.
Thus, the holes in the resulting pNet are indexed by the elements of
the partition \partition.  For the encoding of our running example, we
take $\partition = \bigl\{\{\PortFail, \PortResume\},\{\PortAsk,
\PortReset\}\bigr\}$.  This corresponds to the intuition that the
dangling ports {\PortFail} and {\PortResume} will be provided by a 
monitored component, whereas {\PortAsk} and {\PortReset} correspond to
the actions provided by the ``environment'' (other components of the
system) that are triggered in case of a persistent failure.  Similarly
to the encoding of the coordinators, in the synchronisation vectors of
\nopri[\Gamma], we will associate Boolean values to the actions
corresponding to these ports.

The encoding of the interaction model is based on its representation
as a set of connectors. \noteSB{%
%
  A legitimate question is whether we actually need to do this,
  instead of listing explicitly the individual interactions, then
  associating a synchronisation vector to each one of them?  One of
  the factors is the encoding pNets$\rightarrow$SAT: is it better to
  have more or fewer SVs if the price is additional Boolean variables?
  The other factor might be the facility of maximal progress
  encoding\mdash TBC.  Traceability.  Anything else?
%
} Indeed, as illustrated by the Failure Monitor architecture in
\fig{schema:ArchFailure:BIP}, each connector can define several
allowed interactions, depending on its hierarchical structure and the
use of syncrhons and trigger.  \noteSB{Move this to
  \secn{components}?}For instance, the connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} is a
two-level hierarchical connector.  In the second-level subconnector
{\PortFail \trigsynch \NameCtrl.\PortFail}, the port {\PortFail} is a
trigger, whereas {\NameCtrl.\PortFail} is a synchron.  Therefore, this
subconnector allows two interactions: one with {\PortFail} alone and
another one involving both ports.  Similarly, at the top level,
{\NameTimer.\PortTick} is a trigger, whereas the second-level
subconnector is a synchron.  Hence, the overall connector allows the
interaction involving only one port {\NameTimer.\PortTick}\footnote{%
%
  Such interactions are called \emph{singleton interactions}.
%
} and any interaction involving both {\NameTimer.\PortTick} and
\emph{some} interaction of the second-level subconnector.  Combining
the two levels we can see that the entire connector defines the
following three interactions (observe that $\topinterval +
\NameTimer.t = \topinterval$ and $\NameTimer.z \cap \topinterval =
\NameTimer.z$):
%
\begin{align*}
  \bigl(\{\NameTimer.\PortTick\}&,
  \true, 
  \NameTimer.z := \NameTimer.z\bigr)
  \,,\\
  \bigl(\{\PortFail, \NameTimer.\PortTick\}&,
  \true,
  \NameTimer.z := \NameTimer.z\bigr)
  \,,\\
  \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \}&,
  \true,
  \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
  \,.\
\end{align*}

In order to encode these interactions using only one synchronisation vector, we use
the additional Boolean values associated to each port by the encoding
of coordinator components.  We observe that the three ports in this
connector form a ``causality chain'': {\NameCtrl.\PortFail} can only
participate in the interaction if the dangling port {\PortFail}
participates, which in turn can only participate if
{\NameTimer.\PortTick} does so.  These dependencies can be rewritten
as Boolean implications $\NameCtrl.\PortFail \Rightarrow \PortFail$
and $\PortFail \Rightarrow \NameTimer.\PortTick$.
%
The conjunction of these two implications can be used as a guard for
the synchronisation vector encoding connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)}.

However, notice
that, within the scope of this connector, {\NameTimer.\PortTick}
participates in all the interactions.
%
Furthermore, {\NameTimer.\PortTick} is not involved in any other
connector.  Therefore, the loop transition in \nopri[\NameTimer]
labelled by $\PortTick(\false)$ can never be taken and, therefore can
be removed from the encoding.  Furthermore, since only the transition
labelled by $\PortTick(\true)$ is ever taken, 
%
the
implication $\PortFail \Rightarrow \NameTimer.\PortTick$ is a tautology and can also be discarded.

We obtain the
synchronisation vector SV0 shown in \fig{enc:nopri}, where $b_0$ and
$b_1$ are the Boolean values associated to the actions encoding the
ports {\PortFail} and {\NameCtrl.\PortFail}, $\alpha_0$ is a fresh
action name.  The guard $b_1 \Rightarrow b_0$ encodes the causal
relation among ports {\NameCtrl.\PortFail} and {\PortFail}.  Notice
that, in our encoding, all three ports are present in the
synchronisation vector.  \Fig{enc:nopri} shows
the four synchronisation vectors SV0\ndash SV3 corresponding to the
connectors in \fig{schema:ArchFailure:BIP} and an additional vector
SV4, synchronising the {\init} transitions of the two pLTSs.

The general case encoding relies on the causal semantics of the
algebra of BIP connectors \cite{BliSif10-causal-fmsd}.  Disregarding
the variables and data transfer, the Algebra of Connectors $\ac(P)$
\cite{BliSif07-acp-emsoft} provides a syntactic notation for the BIP
connectors.  The causal semantics of the connectors, given in terms of
the Algebra of Causal Interaction Trees $\ct(P)$, explicitates the
causal dependencies discussed above, through an encoding mapping
$\tau: \ac(P) \rightarrow \ct(P)$.  Another mapping $R: \ct(P)
\rightarrow \cru(P)$ encodes causal interaction trees into systems of
causal rules, which are Boolean implications similar to the ones in
the example above.  We briefly summarise the key definitions and
results in \app{algebras}.

\begin{table}[t]
  \caption{Algebraic representations of the connectors in \fig{schema:ArchFailure:BIP} (see \app{algebras})}
  \label{tab:connectors}
  \resizebox{\columnwidth}{!}{%
  \begin{tabular}{@{}c|c|c@{}}
    \hline {\bf Connector} & {\bf Causal Interaction Tree} & {\bf
      System of Causal Rules}
    \\\hline    
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']

      \node(caa) {}; % node distance=5cm [right of=start]
      \node[node distance=1cm] (cb) [below of=caa]{\PortFail};
      \node[node distance=1.8cm] (cc) [right of=cb]{\NameCtrl.\PortFail};
      \node[node distance=1.6cm] (ca) [left of=cb]{\NameTimer.\PortTick};

      \draw [style=-*, thick, black] ($(caa.south)+(0,.1cm)$) -- ++(right:.5cm) -| (cc.north);
      \draw [style=-triangle 45 reversed, thick, black]  ($(caa.south)+(0,.1cm)$) -| (cb.north);
      \draw [style=-*, thick, black] ($(caa.south)+(0,.5cm)$) -- ++(right:.2cm) -| ($(caa.south)+(+.9,0cm)$);
      \draw [style=-triangle 45 reversed, thick, black]  ($(caa.south)+(0,.5cm)$) -| (ca.north);
    \end{tikzpicture}
%
    &
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth'
        ,initial text=
        ,every state/.style={draw=black,thick}
        ,group/.style = {draw=black,thin,rectangle, minimum width=2.5cm}
        ,port/.style = {font=\small, minimum size=5mm}
        ,legend/.style = {font=\bf}
      ]

      \node(tick){\color{red}\NameTimer.\PortTick};
      \node[node distance=1cm](x) [right of=tick]{};
      \node(fail) [below of=x] {\PortFail};
      \node[node distance=1cm](y) [right of=fail]{};      
      \node(ctrlfail) [below of=y] {\NameCtrl.\PortFail};

      \draw[style=->] (tick) -> (fail);
      \draw[style=->] (fail) -> (ctrlfail);
    \end{tikzpicture}
%
    &
%
    $
    \begin{aligned}[b]
      \NameCtrl.\PortFail &\Rightarrow
      \PortFail {\color{red}{}\land \NameTimer.\PortTick}
      \\
      {\color{red}\PortFail} &{\color{red}{}\Rightarrow \NameTimer.\PortTick}
      \\
      {\color{red}\NameTimer.\PortTick} &{\color{red}{}\Rightarrow \true}
      \\
      {\color{red}\true} &{\color{red}{}\Rightarrow \NameTimer.\PortTick}

    \end{aligned}
    $
%
    \\\hline
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']

      \node[](baa) {}; % node distance=1.8cm [right of=start]
      \node[node distance=1cm] (bb) [below of=baa]{\NameTimer.\PortCancel};
      \node[node distance=1.8cm] (bc) [right of=bb]{\NameCtrl.\PortResume};
      \node[node distance=1.6cm] (ba) [left of=bb]{\PortResume};

      \draw [style=-*, thick, black] ($(baa.south)+(0,.1cm)$) -- ++(right:.5cm) -| (bc.north);
      \draw [style=-*, thick, black]  ($(baa.south)+(0,.1cm)$) -| (bb.north);
      \draw [style=-*, thick, black] ($(baa.south)+(0,.5cm)$) -- ++(right:.2cm) -| ($(baa.south)+(+.9,0cm)$);
      \draw [style=-triangle 45 reversed, thick, black]  ($(baa.south)+(0,.5cm)$) -| (ba.north);
    \end{tikzpicture}
%
    &
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth'
        ,initial text=
        ,every state/.style={draw=black,thick}
        ,group/.style = {draw=black,thin,rectangle, minimum width=2.5cm}
        ,port/.style = {font=\small, minimum size=5mm}
        ,legend/.style = {font=\bf}
      ]

      \node(res){\color{red}\PortResume};
      \node[node distance=1cm](rest) [below of=res] {\NameTimer.\PortCancel\ \NameCtrl.\PortResume};
      \draw[style=->] (res) -> (rest);
    \end{tikzpicture}
%
    &
%
    $
    \begin{aligned}[b]
      \NameCtrl.\PortResume &\Rightarrow
      {\color{red}\PortResume \land{}} \NameTimer.\PortCancel
      \\
      \NameTimer.\PortCancel &\Rightarrow 
      {\color{red}\PortResume \land{}} \NameCtrl.\PortResume
      \\
      {\color{red}\PortResume} &{\color{red}{}\Rightarrow \true}
      \\
      {\color{red}\true} &{\color{red}{}\Rightarrow \PortResume}
    \end{aligned}
    $
%
    \\\hline
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']

      \node (start) {};
      \node(h4)[left of=start]{}; %, node distance=1.5cm

      \node[node distance=1cm] (ab) [below of=start] {\NameTimer.\PortTO};
      \node[node distance=2.1cm] (ac) [right of=ab]{\NameCtrl.\PortTO};
      \node[node distance=1.5cm] (aa) [left of=ab]{\PortAsk};

      \draw [style=-*, thick, black] ($(h4.south)+(0,.1cm)$) -- ++(right:.5cm) -| (ac.north);
      \draw [style=-*, thick, black]  ($(h4.south)+(0,.1cm)$) -| (ab.north);
      \draw [style=-*, thick, black]  ($(h4.south)+(0,.1cm)$) -| (aa.north);
    \end{tikzpicture}
%
    &
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']
      \node (start) {\color{red}\PortAsk\ \NameTimer.\PortTO\ \NameCtrl.\PortTO};
      \node()[below of=start]{};
    \end{tikzpicture}
%
    &
%
    $
    \begin{array}[b]{c@{}}
      {\color{red}
        \begin{aligned}
          \NameCtrl.\PortTO &\Rightarrow
          \PortAsk \land \NameTimer.\PortTO
          \\
          \NameTimer.\PortTO &\Rightarrow 
          \PortAsk \land \NameCtrl.\PortTO
        \end{aligned}}
      \\[11pt]
      {\color{red}
        \begin{aligned}
          \PortAsk &\Rightarrow
          \NameTimer.\PortTO \land \NameCtrl.\PortTO
          \\
          \true &\Rightarrow
          \PortAsk \land \NameTimer.\PortTO
          \land \NameCtrl.\PortTO
        \end{aligned}

      }
    \end{array}
    $
    %
    \\\hline
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']

      \node (start) {};
      \node(h4)[left of=start]{}; %, node distance=1.5cm

      \node[node distance=1cm] (ab) [below of=start] {\NameCtrl.\PortReset};
      \node[node distance=1.5cm] (aa) [left of=ab]{\PortReset};

      \draw [style=-*, thick, black]  ($(h4.south)+(0,.1cm)$) -| (ab.north);
      \draw [style=-*, thick, black]  ($(h4.south)+(0,.1cm)$) -| (aa.north);
    \end{tikzpicture}
%
    &
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']
      \node (start) {\color{red}\PortReset\ \NameCtrl.\PortReset};
      \node()[below of=start]{};
    \end{tikzpicture}
%
    &
%
    {\color{red}$
    \begin{aligned}[b]
      \NameCtrl.\PortReset &\Rightarrow \PortReset
      \\
      \PortReset &\Rightarrow \NameCtrl.\PortReset
      \\
      \true &\Rightarrow
      \PortReset \land \NameCtrl.\PortReset
    \end{aligned}
    $}
%
    \\\hline
  \end{tabular}}
\end{table}

First, consider a flat connector with $n$ trigger ports $p_1,\dots,
p_n$ (with $n > 0$) and $m$ synchron ports $q_1,\dots, q_m$
(\fig{connector}).  Similarly to the example above, for any of the
ports $q_1,\dots, q_m$ to participate in an interaction defined by
this connector, at least one of the ports $p_1,\dots, p_n$ must
participate.  This causal dependency can be represented by the Boolean
formula $q_1 \lor \dots \lor q_m \Rightarrow p_1 \lor \dots \lor p_n$.
If there are no synchrons, \ie $m = 0$, this formula simplifies to
$p_1 \lor \dots \lor p_n$.  However, if there are no triggers, \ie $n
= 0$, the connector semantics is that \emph{all} synchron ports must
participate for an interaction to take place, represented by the
Boolean formula $(q_1 = q_2) \land \dots \land (q_1 = q_m)$.

We now generalise this principle to hierarchical connectors, building
the formula bottom-up.  In a flat connector above, we have implicitly
associated a Boolean variable to each port.  In a hierarchical
connector, we associate a fresh Boolean variable to each node.
Similarly to the above example, consider a connector with $n > 0$
trigger sub-connectors and $m$ synchron sub-connectors.  We associate
to this connector a fresh Boolean variable $x$.  Assume that Boolean
variables associated to trigger sub-connectors are $x_1,\dots,x_n$,
whereas those associated to synchron sub-connectors are
$y_1,\dots,y_m$.  The causal dependencies among the interactions
defined by the sub-connectors can be encoded by the formula
%
\begin{equation}
  \label{eq:trig:encoding}
  (y_1 \lor \dots \lor y_m \Rightarrow x) \land (x = x_1 \lor \dots
  \lor x_n).
\end{equation}
%
If there are no trigger sub-connectors, \ie $n = 0$, we use the
formula
%
\begin{equation}
  \label{eq:synch:encoding}
  x = \bigl((y_1 = y_2) \land \dots \land (y_1 = y_m)\bigr)
\end{equation}
%
($x$ is equal to $\true$ if all $x_1, \dots, x_m$ are equal, it is
equal to $\false$ otherwise).

Notice that, in the flat connector example above, the formula was
essentially the same, with $p_i$ instead of $x_i$ and $q_i$ instead of
$y_i$, and without the additional variable $x$.  Indeed, $x$ is used
to encode the causal dependencies between this connector and, if
necessary, those in higher levels of the hierarchy.  The Boolean
encoding of the causal dependencies among the ports of the entire
connector is obtained by taking the conjunction of formulas obtained
for each sub-connector.

Consider again the connector {\NameTimer.\PortTick \trigsynch
  (\PortFail \trigsynch \NameCtrl.\PortFail)}.  Associating the
varaibles $b_0$, $b_1$ and $b_2$ to ports {\PortFail},
{\NameCtrl.\PortFail} and {\NameTimer.\PortTick}, respectively, and
variables $y_1$ and $y_2$ to the top- ({\NameTimer.\PortTick
  \trigsynch $\cdot$}) and second-level (\PortFail \trigsynch \NameCtrl.\PortFail) connectors in the hierarchy,
we obtain the following formula
%
\begin{equation}
  \label{eq:tff:example}
  (b_1 \Rightarrow y_2) \land (y_2 = b_0) \land (y_2 \Rightarrow y_1)
  \land (y_1 = b_2)
  \,.
\end{equation}
%
Observe now that the top-level connector, by its very nature, does not
have any higher level connectors and therefore we can eliminate the
corresponding Boolean variable.  This simplifies \eq{tff:example}, to
%
\begin{equation}
  \label{eq:tff:example:1}
  (b_1 \Rightarrow y_2) \land (y_2 = b_0) \land (y_2 \Rightarrow b_2)
  \,.
\end{equation}
%
In the general case, this corresponds to replacing, \emph{for
  top-level connectors}, equations \eq{trig:encoding} and
\eq{synch:encoding}, respectively, by
%
\begin{gather}
  \label{eq:trig:encoding:1}
  y_1 \lor \dots \lor y_m \Rightarrow x_1 \lor \dots
  \lor x_n
  \\
  \intertext{and}
  \label{eq:synch:encoding:1}
  (y_1 = y_2) \land \dots \land (y_1 = y_m)
  \,.
\end{gather}
%
Furthermore, in the context of one connector (or synchronisation
vector), some interaction must be fired.  Thus, \emph{for top-level
  connectors}, \eq{synch:encoding:1} can be further simplified to 
%
\begin{equation}
  \label{eq:synch:encoding:2}
  y_1 \land \dots \land y_m
  \,.
\end{equation}

%****************************************************************
\subsection{Encoding with the maximal progress priority model}
\label{secn:enc:maxprog}

%****************************************************************
%****************************************************************

\section{SMT encoding of open pNets}
\label{secn:smt}

%****************************************************************
%****************************************************************

\section{Case study}
\label{secn:case-study}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{FailureTimerZonePNET}
  \caption{The pNET encoding the Failure Monitor architecture}
  \label{schema:ArchFailure:BIP}
\end{figure}

%****************************************************************
%****************************************************************

\section{Related work}
\label{secn:related}

%****************************************************************
%****************************************************************

\section{Conclusion}
\label{secn:conclusion}

%****************************************************************
%****************************************************************

\bibliographystyle{abbrv}
\bibliography{biblio}

%****************************************************************
%****************************************************************
\appendix
\clearpage

%****************************************************************
%****************************************************************

\section{Proofs}
\label{secn:proofs}

\begin{lemma}
  \label{lem:onlyone}
  Let $A = \tuple{\cC, V_A, P_A, \export[A], \Gamma}$ be an architecture and denote
  by $\Gamma_\cC \bydef{=}
%
  \setdef{
    (a \cap P_\cC, \true, \noop)
  }{
    (a, g, e) \in \Gamma
  }$, with $P_\cC = \bigcup_{C \in \cC} P_C$,
%  
  the projection of $\Gamma$ onto the coordinating components of
  $A$.  Consider the architecture $A' = \tuple{\{C'\}, V_A, P_A, \export[A],
  \Gamma}$, where $C' = \Gamma_\cC(\cC)$.  For any set of
  components $\cB$, satisfying the conditions of
  \defn{arch:application}, we have
  $\semopen{A(\cB)} = \semopen{A'(\cB)}$.
\end{lemma}
%
\begin{proof}
  First of all, notice that, by \defn{im:sem},
  $V_{C'} = \bigcup_{C \in \cC} V_C$ and $P_{C'} = P_\cC = \bigcup_{C \in \cC} P_C$.
  The conditions of \defn{arch} are satisfied and $A'$ is indeed
  an architecture.  Furthermore, $\cB$ satisfies the conditions
  of \defn{arch:application} \wrt $A'$.  Hence, the component
  $A'(\cB)$ is well defined.

  Clearly the state spaces and initial states %% and interfaces
  of both LTSs $\semopen{A(\cB)}$ and $\semopen{A'(\cB)}$
  coincide.  Thus, we only have to prove that so do the
  transition relations.  Let us assume that
  $\cC = C_i^{i \in I}$ and $\cB = B_j^{j \in J}$.
  We will use $q_i, q_i'$ to denote the states of
  $C_i$ and $q_j, q_j'$ to denote the states of $B_j$,
  and similarly for the valuations of variables.

  By \defn{comp:semantics},
%
  \begin{equation}
    \label{eq:lem1:trans:sem}
    (q_i, \val{}{i})^{i \in I} (q_j, \val{}{j})^{j \in J}
    \goesto{a, \val[\tilde]{}{}}
    (q_i', \val{\prime}{i})^{i \in I}
    (q_j', \val{\prime}{j})^{j \in J}
  \end{equation}
%  
  is a transition in $\semopen{A(\cB)}$ (\resp
  $\semopen{A'(\cB)}$) iff
%
  \begin{equation}
    \label{eq:lem1:trans}
    q_i^{i \in I} q_j^{j \in J}
    \goesto{a, g', e'}
    (q_i')^{i \in I} (q_j')^{j \in J}
    \,,
  \end{equation}
%
  is a transition in $A(\cB)$ (\resp $A'(\cB)$) and
%
  \begin{mathpar}
    \val{i \in I}{i} \val{j \in J}{j}
    \models g'
    \,,
    \and
    (\val{\prime}{i})^{i \in I} (\val{\prime}{j})^{j \in J}
    = \val[\tilde]{}{}[e']
    \,,
    \and
    \valdiff{\val{i \in I}{i} \val{j \in J}{j}}{
      \val[\tilde]{}{}} \subseteq \export(a)       
  \end{mathpar}
%
  (with $\export$ defined as in \defn{arch:application}).

  By \defn{im:sem} and \defn{priority:sem}, \eq{lem1:trans} is a transition in $A(\cB)$
  iff $a \neq \emptyset$, $(a, g, e) \in \IMextend{\Gamma}{P}$,
  where $P = \bigcup_{B \in \cB \cup \cC} P_B$, and
  %
  \begin{enumerate}
  \item\label{ab:start} for $i \in I$,\\
    $a \cap P_{C_i} \neq \emptyset$ and
    $q_i \goesto{a \cap P_{C_i},\, g_{C_i},\, e_{C_i}} q_i'$
    is a transition in $C_i$, or\\
    $a \cap P_{C_i} = \emptyset$ and $q_i = q_i'$
    (for covenience, we put $g_{C_i} = \true$ and $e_{C_i} = \noop$ in item~\ref{aux1} below);
  \item for $j \in J$,\\
    $a \cap P_{B_j} \neq \emptyset$ and
    $q_j \goesto{a \cap P_{B_j},\, g_{B_j},\, e_{B_j}} q_j'$
    is a transition in $B_j$, or\\
    $a \cap P_{B_j} = \emptyset$ and $q_j = q_j'$
    (for covenience, we put $g_{B_j} = \true$ and $e_{B_j} = \noop$ in item~\ref{aux1} below);
  \item\label{ab:end}\label{aux1} $g' = g \land \bigwedge_{i \in I} g_{C_i} \land
    \bigwedge_{j \in J} g_{B_j} \land \varphi_\mu$ and
    $e' = e; e_{C_i}^{i \in I}, e_{B_j}^{j \in J}$, where    
%
    \begin{equation}
      \label{eq:only1:maxprog:guard}
      \varphi_\mu = \bigwedge_{
        \begin{array}{c}
          q_i^{i \in I} q_j^{j \in J} \goesto{b, h, f} \text{ in }
          (\IMextend{\Gamma}{P})(\cC \cup \cB)
          \\
          a \varsubsetneq b
        \end{array}
      } \lnot h
      \,.
    \end{equation}
%
  \breakenumistart

  Similarly, \eq{lem1:trans} is a transition in
  $A'(\cB)$ iff $a \neq \emptyset$,
  $(a, g, e) \in \IMextend{\Gamma}{P}$ and
%
  \breakenumiend
% 
  \item \label{a1b:start} \label{only1:coord}
    $a \cap P_{C'} \neq \emptyset$ and 
    $(q_i)^{i \in I} \goesto{a \cap P_{C'}, g_{C'}, e_{C'}} (q_i')^{i \in I}$
    is a transition in $C'$, or\\
    $a \cap P_{C'} = \emptyset$ and $q_i = q_i'$,
    for all $i \in [1,m]$
    (for covenience, we put $g_{C'} = \true$ and $e_{C'} = \noop$ in
    item~\ref{aux2} below);
  \item for $j \in J$,\\
    $a \cap P_{B_j} \neq \emptyset$ and
    $q_j \goesto{a \cap P_{B_j}, g_{B_j}, e_{B_j}} q_j'$
    is a transition in $B_j$, or\\
    $a \cap P_{B_j} = \emptyset$ and $q_j = q_j'$
    (for covenience, we put $g_{B_j} = \true$ and $e_{B_j} = \noop$ in
    item~\ref{aux2} below);
  \item\label{aux2} $g' = g \land g_{C'} \land
    \bigwedge_{j \in J} g_{B_j} \land \varphi_\mu'$ and
    $e' = e; e_{C'}, e_{B_j}^{j \in J}$ (with $\varphi_\mu'$ defined as
    in \eq{only1:maxprog:guard}, but substituting $\{C'\}$ instead of $\cC$).
%
  \breakenumistart
  Notice that, for any $i \in I$, since
  $P_{C_i} \subseteq P_{C'}$, we have
  $a \cap P_{C_i} = (a \cap P_{C'}) \cap P_{C_i}$.
  Thus, if $a \cap P_{C'} \neq \emptyset$, 
  the transition in condition~\ref{only1:coord} above is
  present in $C'$ iff
  $(a \cap P_{C'}, \true, \noop) \in \Gamma_\cC$ and
  \breakenumiend
% 
  \item for $i \in I$,\\
    $a \cap P_{C_i} \neq \emptyset$ and
    $q_i \goesto{a \cap P_{C_i}, g_{C_i}, e_{C_i}} q_i'$
    is a transition in $C_i$, or\\
    $a \cap P_{C_i} = \emptyset$ and $q_i = q_i'$,
    (for covenience, we put $g_{C_i} = \true$ and $e_{C_i} = \noop$ in
    item~\ref{aux3} below);
  \item \label{a1b:end} \label{aux3} $g_{C'} = \bigwedge_{i \in I} g_{C_i}$ and
    $e_{C'} = e_{C_i}^{i \in I}$.
  \end{enumerate}

  Consider $(a,g,e) \in \IMextend{\Gamma}{P}$.  By
  \eq{im:extension}, this is equivalent to
  $(a \cap P_A, g, e) \in \Gamma$.
  Since $P_{C'} = P_\cC \subseteq P_A$, we have
  $a \cap P_{C'} = a \cap P_\cC = (a \cap P_A) \cap P_\cC$.
  Hence, by the definition of $\Gamma_\cC$ in the lemma statement,
  this is also equivalent to
  $(a \cap P_{C'}, \true, \noop) \in \Gamma_\cC$.

  Thus, conditions \ref{ab:start}\ndash\ref{ab:end} are satisfied iff
  so are conditions \ref{a1b:start}\ndash\ref{a1b:end} (to see that
  $\varphi_\mu$ is satisfied iff so is $\varphi_\mu'$, one has to
  consider the same two sets of conditions, but this time without the
  additional conjuncts $\varphi_\mu$ and $\varphi_\mu'$ in
  items~\ref{aux1} and \ref{aux2}, since no priority model is applied
  to obtain $C'$).  This proves that the transition relations of
  $\semopen{A(\cB)}$ and $\semopen{A'(\cB)}$ coincide and, therefore,
  $\semopen{A(\cB)} = \semopen{A'(\cB)}$.
\end{proof}

\begin{note}
  \label{rem:upwards:compat}
  Notice that, since, in the statement of \lem{onlyone}, 
  %
  \begin{enumerate}
  \item the interaction models are the same for both $A$ and $A'$, and
  \item $P_{C'} = P_\cC$,
  \end{enumerate}
  %
  substituting $A'$ for $A$ preserves upwards compatibility, \ie for any
  architecture $A''$, holds the following statement: $A$ and $A''$ are
  upwards compatible if and only if so are $A'$ and $A''$.
\end{note}

\begin{lemma}
  \label{lem:onestep}
  Let $\cB$ be a set of components; let $A_i = \tuple{\{C_i\}, V_{A_i},
  P_{A_i}, \export[A_i], \Gamma_i}$, for $i = 1,2$, be two upwards compatible architectures
  with one coordinator each, applicable to $\cB$.  Denote
  $P_i = P_{C_i} \cup \bigcup_{B \in \cB} P_B$.
  Finally, let
%
  $
  s_1 s_2 s
%
  \goesto{a, \tilde{\val{}{}}}
%
  s_1' s_2' s'
  $,
%
  be a transition in $\semopen{(A_1 \arcomp A_2)(\cB)}$, with
  $a \neq \emptyset$, 
  $s, s' \in \prod_{B \in \cB} S_{\semopen{B}}$
  and
  $s_i, s_i' \in S_{\semopen{C_i}}$ ($i=1,2$).
%
  Then, for $i=1,2$, either $a \cap P_i = \emptyset$ and $s_i s=
  s_i' s'$, or there exists a transition
%
  $
  s_i s
%
  \goesto{a \cap P_i, \val[\doubletilde]{}{}}
%
  s_i' s''
  $ 
%
  in $\semopen{A_i(\cB)}$, with
  $\val[\doubletilde]{}{}$ being the restriction of
  $\val[\tilde]{}{}$ to
  $V_i = V_{C_i} \cup \bigcup_{B \in \cB} V_B$ and
  $s' \order s''$.
\end{lemma}
%
\begin{proof}
  By \defn{comp:semantics}, every semantic state consists of a
  component state and a valuation of component variables.  Thus,
  we can expand $s_1 s_2 s \goesto{a, \tilde{\val{}{}}} s_1' s_2' s'$
  as
%
  \[
  (q_1, \val{}{1}, q_2, \val{}{2}, q, \val{}{})
%
  \goesto{a, \tilde{\val{}{}}}
%
  (q_1', \val[\primeit]{}{1}, q_2', \val[\primeit]{}{1}, q', \val[\primeit]{}{})
  \,.
  \]
%
  By \eq{im:sem} and \eq{priority:sem}, this implies
  that
  $(q_1, q_2, q) \goesto {a, g', e'} (q_1', q_2', q')$
  is a transition in $(A_1 \arcomp A_2)(\cB)$, with
%
  $g' = g \land g_1 \land g_2 \land g_\cB \land \varphi_\mu$
  and
  $e' = e; (e_1, e_2, e_\cB)$,
  where $(a, g, e) \in \Gamma$ (see \eq{arch:composition} for the
  definition of $\Gamma$);
%  
  $g_1$, $g_2$ and $e_1$, $e_2$ are, respectively, the guards and
  update expressions of the underlying transitions in the coordinating
  components $C_1$ and $C_2$;
%    
  \begin{mathpar}
    g_\cB = \bigwedge_{B \in \supp{a} \cap \cB} g_B
    \,,
    \and
    e_\cB = e_B^{B \in \supp{a} \cap \cB}
    \,,
  \end{mathpar}
%
  are the compositions of, respectively, the guards and update
  expressions of the operand components;
%    
  \begin{equation}
    \label{eq:onestep:maxprog:guard}
    \varphi_\mu = \bigwedge_{
      \begin{array}{c}
        (q_1, q_2, q) \goesto{b, h', f'} \text{ in }
        \bigl(\IMextend{\Gamma}{(P_1 \cup P_2)}\bigr)
        \bigl(\{C_1, C_2\} \cup \cB\bigr)
        \\
        a \varsubsetneq b
      \end{array}
    } \lnot h'
  \end{equation}
%
  is the conjunct encoding the maximal progress priority
  \eq{priority:sem}.  As above for $g'$, for each $(b, h', f')$ in
  \eq{onestep:maxprog:guard}, $h'$ can be decomposed as $h' = h \land
  h_1 \land h_2 \land h_\cB$ (without the maximal progress conjunct,
  since no priority is applied in $\bigl(\IMextend{\Gamma}{(P_1 \cup
    P_2)}\bigr)\bigl(\{C_1, C_2\} \cup \cB\bigr)$).

  By \eq{comp:semantics}, we also have
%
  \begin{align}
    \label{eq:guard}
    (\val{}{1}, \val{}{2}, \val{}{}) &\models g'
    \,,
    \\
    \label{eq:substitution}
    (\val[\primeit]{}{1}, \val[\primeit]{}{2}, \val[\primeit]{}{})
    &= \val[\tilde]{}{}[e']
    \,,
    \\
    \label{eq:transient}
    \valdiff{(\val{}{1}, \val{}{2}, \val{}{})}{\val[\tilde]{}{}}
    &\subseteq \export(a)
    \,.
  \end{align}
%
  If $a \cap P_1 = \emptyset$, then $\supp{a} = \{C_2\}$ and, by
  \eq{transient} and \defn{im:sem}, $s_1 = s_1'$ and $s = s'$
  (recall, \defn{im}, that, for $(a,g,e) \in \Gamma$, we have $e
  \in \exprs{V_a}$).  Hereafter, assume $a \cap P_1 \neq
  \emptyset$.

  By \eq{arch:composition}, for $i=1,2$, there exist $(a \cap P_{A_i},
  g^i, e^i) \in \Gamma_i$, such that $g = g^1 \land g^2$ and $e = e^1
  \land e^2$.
  %% Similarly, for each $(b, h', f')$ in \eq{onestep:maxprog:guard},
  %% there exist $(b \cap P_{A_i}, h^i, f^i) \in \Gamma_i$, such that
  %% $h = h^1 \land h^2$ and $f = f^1 \land f^2$.
%
  We want to show that the interaction $(a \cap P_1, g^1, e^1)$ is
  enabled in the state $(q_1, \val{}{1}, q, \val{}{})$ of
  $\semopen{A_1(\cB)}$.  To this end, we proceed in two steps:
%
  \begin{enumerate}
  \item \label{onestep:proof:im} We show that $(a \cap P_1, g^1,
    e^1)$ is enabled in the state $(q_1, \val{}{i}, q, \val{}{})$ of
    $\semopen{(\IMextend{\Gamma_1}{P_1})(\cB)}$.
  \item \label{onestep:proof:maxp} We show that $(a \cap P_1, g^1,
    e^1)$ is not inhibited in $\semopen{A_1(\cB)} =
    \semopen{\mu\bigl((\IMextend{\Gamma_1}{P_1})(\cB)\bigr)}$ by the
    maximal progress priority.
  \end{enumerate}

  \paragraph*{Step \ref{onestep:proof:im}}
  %% Denoting $P_\cB = \bigcup_{B \in \supp{a} \cap \cB} P_B$, 
  By \eq{im:sem}, we have
%
  \begin{mathpar}
    q_1 \goesto{a \cap P_{C_1}, g_1, e_1} q_1'
    \and
    \text{and}
    \and
    q_B \goesto{a \cap P_B, g_B, e_B} q_B',
    \text{ for each }
    B \in \cB
    \,.
  \end{mathpar}

  By \eq{guard}, we have
  $(\val{}{1}, \val{}{}) \models g^1 \land g_1 \land g_\cB$.

  Denoting \todoSB{One tilde is enough}$\val[\doubletilde]{}{1}$ the restriction of
%
  $\val[\tilde]{}{}$ to
  $V_1 = V_{C_1} \cup \bigcup_{B \in \cB} V_B$,
%
  we have, by \eq{substitution} and the monotonicity of
  expressions,
%
  $(\val[\primeit]{}{1}, \val[\primeit]{}{}) \order
  \val[\doubletilde]{}{1}[e^1; (e_1, e_b)]$.  \todoSB{Remove this}\hl{Notice that all
  variables affected by both $e^1$ and $e^2$ must necessarily
  belong to $\bigcup_{B \in \cB} V_B$.}  Hence,
%
  $\val[\doubletilde]{}{1}[e^1; (e_1, e_b)] =
  (\val[\primeit]{}{1}, \val[\secondit]{}{})$ with 
  $\val[\primeit]{}{} \order \val[\secondit]{}{}$.

  By \eq{comp:semantics} and \eq{im:sem}, we have
  $
  (q_1, \val{}{1}, q, \val{}{})
  \goesto{a \cap P_1, \val[\doubletilde]{}{1}}
  (q_1', \val[\primeit]{}{1}, q, \val[\secondit]{}{})
  $ in $\semopen{(\IMextend{\Gamma_1}{P_1})(\cB)}$.

  Symmetrically, we have 
  $
  (q_2, \val{}{2}, q, \val{}{})
  \goesto{a \cap P_2, \val[\doubletilde]{}{2}}
  (q_2', \val[\primeit]{}{2}, q, \val[\secondit]{}{})
  $ in $\semopen{(\IMextend{\Gamma_2}{P_2})(\cB)}$.

  \paragraph*{Step \ref{onestep:proof:maxp}}
  Suppose that $a \cap P_1 \varsubsetneq b_1$, for some interaction
  $(b_1, h^1, f^1) \in \IMextend{\Gamma_1}{P_1}$ enabled in $(q_1,
  \val{}{1}, q, \val{}{})$.  Then, by upwards compatibility of $A_1$
  and $A_2$, there exists $(b_2, h^2, f^2) \in
  \IMextend{\Gamma_2}{P_2}$, enabled in $(q_2, \val{}{2}, q,
  \val{}{})$, such that $a \cap P_2 \subseteq b_2$ and $b_2 \cap P_1 =
  b_1 \cap P_2$.  By \eq{im:extension}, we have $(b_1 \cap P_{A_1},
  h^1, e^1) \in \Gamma^1$.  Furthermore,
%
  \begin{multline*}
    ((b_1 \cap P_{A_1}) \cup (b_2 \cap P_{A_2}))  \cap P_{A_1} =
    (b_1 \cap P_{A_1}) \cup (b_2 \cap P_{A_1} \cap P_{A_2}) =\\
    (b_1 \cap P_{A_1}) \cup ((b_2 \cap P_1) \cap P_{A_1} \cap P_{A_2}) =  
    (b_1 \cap P_{A_1}) \cup ((b_1 \cap P_2) \cap P_{A_1} \cap P_{A_2}) =\\
    (b_1 \cup (b_1 \cap P_2 \cap P_{A_2})) \cap P_{A_1} =
    b_1 \cap P_{A_1} 
  \end{multline*}
%
  Symmetrically, $(b_2 \cap P_{A_2}, h^2, e^2) \in \Gamma^2$ and
  $((b_1 \cap P_{A_1}) \cup (b_2 \cap P_{A_2})) \cap P_{A_2} = b_2
  \cap P_{A_2}$.  By \eq{arch:composition}, we then have $((b_1 \cap
  P_{A_1}) \cup (b_2 \cap P_{A_2}), h^1 \land h^2, f^1 \expmix f^2)
  \in \Gamma$ and, by \eq{im:extension}, $(b_1 \cup b_2, h^1 \land
  h^2, f^1 \expmix f^2) \in \IMextend{\Gamma}{(P_1 \cup P_2)}$.
%
  Since both $(b_i, h^i, f^i) \in \IMextend{\Gamma_i}{P_i}$ are
  enabled in the respective states $(q_i, \val{}{i}, q, \val{}{})$, we
  conclude that $(b_1 \cup b_2, h^1 \land h^2, f^1 \expmix f^2)$ is
  enabled in $(q_1, \val{}{1}, q_2, \val{}{2}, q, \val{}{})$.

  Since $a \cap P_1 \varsubsetneq b_1$ and $a \cap P_2 \subseteq b_2$,
  we have $a \varsubsetneq b_1 \cup b_2$.  Thus, $(b_1 \cup b_2, h^1
  \land h^2, f^1 \expmix f^2)$ contribute to $\varphi_\mu$ in
  \eq{onestep:maxprog:guard}.  Since $(b_1 \cup b_2, h^1 \land h^2,
  f^1 \expmix f^2)$ is enabled in $(q_1, \val{}{1}, q_2, \val{}{2}, q,
  \val{}{})$, we have $(\val{}{1}, \val{}{2}, \val{}{}) \models h^1
  \land h^2$ and, therefore $(\val{}{1}, \val{}{2}, \val{}{})
  \not\models \varphi_\mu$, which contradicts \eq{guard}, thereby
  proving that $(a \cap P_1, g^1, e^1)$ is not inhibited in
  $\semopen{A_1(\cB)} =
  \semopen{\mu\bigl((\IMextend{\Gamma_1}{P_1})(\cB)\bigr)}$ by the
  maximal progress priority.

  The corresponding statement for $(a \cap P_2, g^2, e^2)$ is obtained
  symmetrically.
\end{proof}

\begin{lemma}
  \label{lem:stepabove}
  Consider a component $B$ and let 
%
  $
  (q_1, \val{}{1})
%
  \goesto{a, \val[\tilde]{}{}}
%
  (q_2, \val{}{2})
  $,
%
  be a transition in $\semopen{B}$.  For any valuation
  $\val[\primeit]{}{1}$, such that
  $\val{}{1} \order \val[\primeit]{}{1}$ and
  $\valdiff{\val{}{1}}{\val[\primeit]{}{1}} \subseteq \export(a)$,
  there exists a transition
%
  $
  (q_1, \val[\primeit]{}{1})
%
  \goesto{a, \val[\tilde]{}{}}
%
  (q_2, \val{}{2})
  $
%
  in $\semopen{B}$.
\end{lemma}
%
\begin{proof}
  By \eq{comp:semantics}, we have
%
  \begin{mathpar}
    q_1 \goesto{a, g, e} q_2
    \,,
    \and
    \val{}{1} \models g
    \,,
    \and
    \val{}{2} = \val[\tilde]{}{}[e]
    \,,
    \and
    \text{and}
    \and
    \valdiff{\val{}{1}}{\val[\tilde]{}{}} \subseteq \export(a)
    \,.
  \end{mathpar}

  By monotonicity of guards, we deduce
  $\val[\primeit]{}{1} \models g$.
  For any $v \not\in \export(a)$,
  $\val[\primeit]{}{1}(v) = \val{}{1}(v) = \val[\tilde]{}{}(v)$.
  Hence, 
  $\valdiff{\val[\primeit]{}{1}}{\val[\tilde]{}{}}
  \subseteq \export(a)$.
  By \eq{comp:semantics}, we conclude 
%
  $
  (q_1, \val[\primeit]{}{1})
%
  \goesto{a, \val[\tilde]{}{}}
%
  (q_2, \val{}{2})
  $.  
\end{proof}

\begin{proof}[Proof of \thm{combining}]
  By \lem{onlyone}, we can assume without loss of generality that each of the two
  architectures has only one coordinating component.  For $i =
  1,2$, we denote $\{C_i\} = \cC_i$, $P_i = P_{C_i}
  \cup\ \bigcup_{B \in \cB} P_B$ and $V_i = V_{C_i}
  \cup\ \bigcup_{B \in \cB} V_B$.
%
  By \rem{upwards:compat}, this does not affect their upwards
  compatibility, hence neither does this affect the applicability of
  \lem{onestep} below.

  The initiality of $\Phi_1 \land \Phi_2$, is trivial: both
  $\Phi_1$ and $\Phi_2$ are initial, hence $s^0 \models \Phi_1
  \land \Phi_2$.

  Consider a path
%
  \[
  s^0_1 s^0_2 s^0
%
  \goesto{a^1, \val[\tilde]{1}{}}
%
  s^1_1 s^1_2 s^1
%
  \goesto{a^2, \val[\tilde]{2}{}}
  \cdots
  \goesto{a^k, \val[\tilde]{k}{}}
%
  s^k_1 s^k_2 s^k
  \]
%
  in $\semopen{(A_1 \arcomp A_2)(\cB)}$, where
  $s^0,\dots,s^k \in \prod_{B \in \cB} S_{\semopen{B}}$ and
  $s^0_i,\dots, s^k_i \in S_{\semopen{C_i}}$, for $i=1,2$.
  We have to show that $s^k \models \Phi_1 \land \Phi_2$.

  Assuming that $a^1 \cap P_1 \neq \emptyset$, by \lem{onestep},
  there exists $s^{1\prime} \in \prod_{B \in \cB} S_{\semopen{B}}$,
  such that
%
  \[
  s^0_1 s^0
%
  \goesto{a^1 \cap P_1, \val[\doubletilde]{1}{}}
%
  s^1_1 s^{1\prime}
  \]
%
  is a transition in $\semopen{A_1(\cB)}$ with
  $\val[\doubletilde]{1}{}$ being the restriction of
  $\val[\tilde]{1}{}$ to $V_1$ and
  $s^1 \order s^{1\prime}$.
%
  By \lem{stepabove}, 
  \[
  s^1_1 s^1_2 s^{1\prime}
%
  \goesto{a^2, \val[\tilde]{2}{}}
%
  s^2_1 s^2_2 s^2
  \]
  is a transition in $\semopen{(A_1 \arcomp A_2)(\cB)}$ and,
  therefore, 
  \[
  s^1_1 s^1_2 s^{1\prime}
%
  \goesto{a^2, \val[\tilde]{2}{}}
%
  s^2_1 s^2_2 s^2
%
  \goesto{a^3, \val[\tilde]{3}{}}
  \cdots
  \goesto{a^k, \val[\tilde]{k}{}}
%
  s^k_1 s^k_2 s^k
  \]
%  
  is a path in $\semopen{(A_1 \arcomp A_2)(\cB)}$.
%
  Notice that, if the above assumption, $a^1 \cap P_1 \neq
  \emptyset$, does not hold, then $s^0_1 = s^1_1$, $s^0 = s^1$
  and we can take $s^{1\prime} = s^1$ in this latter path.

  Repeating the entire argument $k-1$ times, starting with this
  shorter path, we obtain a path
%
  \begin{equation}
    \label{eq:path-in-1}
    s^0_1 s^0
    %
    \goesto{a^1 \cap P_1, \val[\doubletilde]{1}{}}
    %
    s^1_1 s^{1\prime}
    %
    \goesto{a^2 \cap P_1, \val[\doubletilde]{2}{}}
    %
    s^2_1 s^{2\prime}
    %
    \goesto{a^3 \cap P_1, \val[\doubletilde]{3}{}}
    \cdots
    \goesto{a^k \cap P_1, \val[\doubletilde]{k}{}}
    %
    s^k_1 s^{k\prime}
  \end{equation}
%  
  with $s^i \order s^{i\prime}$, for all $i \in [1,k]$.

  From \eq{path-in-1}, we conclude that the state $s^k_1
  s^{k\prime}$ is reachable in $\semopen{A_1(\cB)}$.  Since $A_1$
  enforces $\Phi_1$ on $\cB$, this implies that $s^{k\prime}
  \models \Phi_1$.  Since $s^k \order s^{k\prime}$, we deduce, by
  \defn{property}, that $s^k \models \Phi_1$.

  Symmetrically, $s^k \models \Phi_2$, which concludes
  the proof.
\end{proof}

%*********************************************************************
%*********************************************************************

\section{pNets additional definitions}
\label{secn:pNets:additional}

\begin{definition}[Sorts, Holes, Leaves of pNets]
  \begin{itemize}
  \item  \noteSB{I think this can be lightened up a bit, since the definitions are all 
  straightforward, whereas the notation is complex.  Maybe reformulate in words and move 
  the formal definitions in the appendix?}
 The sort of a pNet is its signature, i.e. the set of actions it can
perform. \noteSB{Is there any chance that, by writing assignments explicitly ($x := e$), 
we could eliminate the need to distinguish the input variables at all? That would 
simplify things.}\hl{In the definition of sorts, we do not need to distinguish
input variables} (that specify the dataflow within LTSs), so for
computing LTS sorts, we use a substitution operator\footnote{$\subst{y_k\gets x_k}^{k\in 
K}$ is the parallel substitution 
operation.} to remove the
\emph{input marker} of variables. Formally:
\[
\begin{array}{l}
\Sortop(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{x \gets ?x| 
x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~g,~(x_j\!:= {e}_j)^{j\in
    J}\rangle} s'\in \to \} \\
\Sortop(\mylangle \overline{\pNet}\!, %\pNet_i^{i\in I}, \Sort_j^{j\in J}
\overline{\pNet}\!,
\overline{\symb{SV}}\myrangle)
=\{\alpha' |\, \overline{\alpha}%\alpha_j^{j\in J_k}
\to\alpha'\in\set{\symb{SV}}\}
\end{array}
\]

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined 
inductively as:
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\emptyset\\%\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i\pNet_i \text{ is a }\pLTS\}}
\end{array}\]
\end{itemize}
\end{definition}

%*********************************************************************
%*********************************************************************

\section{Algebraic representations of the interaction model}
\label{secn:algebras}

In this section condensed from
\cite[Section~4]{BarBliu15-offer-scico}, we briefly recall the syntax
and semantics of the algebras used to represent BIP interaction models
\emph{disregarding variables and data transfer}.  All the algebras are
parameterised by a set $P$ of all the ports in a given system.  The
semantics of the Algebra of Interactions $\ai(P)$ is given in terms of
sets of interactions by a function $\intsem{\cdot}: \ai(P) \rightarrow
2^{2^P}$.  The corresponding equivalence relation on $\ai(P)$ is
defined as follows: two terms $x,y \in \ai(P)$ are {\em equivalent} $x
\simeq y$ iff $\intsem{x} = \intsem{y}$.  For any other algebra
$\cA(P)$ among those appearing in this section, we define its
semantics by the function $\aisem{\cdot}: \cA(P) \rightarrow \ai(P)$.
A function $\intsem{\cdot}: \cA(P) \rightarrow 2^{2^P}$ is obtained by
composing $\aisem{\cdot}: \cA(P) \rightarrow \ai(P)$ and
$\intsem{\cdot}: \ai(P) \rightarrow 2^{2^P}$.  The axiomatisation of
$\ai(P)$ given in \cite{BliSif07-acp-emsoft} is sound and complete
with respect to $\simeq$.  Hence, for other algebras, the equivalences
induced by $\intsem{\cdot}$ and $\aisem{\cdot}$ coincide.

Below, we assume that a set of ports $P$ is given, such that $0,1\not\in
P$.

For any set $X$ of propositional variables, we denote by $\sB[X]$ the
corresponding Boolean algebra generated by $X$.  For presentation
clarity, we will often omit the conjunction operator and write $a \lor
bc$ instead of $a \lor (b \land c)$.

% *********************************************************************

\subsection{Algebra of Interactions}
\label{secn:ai}

The Algebra of Interactions is used to define the interaction semantics of other algebras. The elements of this algebra can be bijectively mapped to interaction models, \ie subsets of $2^P$.

\begin{syntax}
The syntax of the {\em Algebra of Interactions}, $\ai(P)$, is defined by
the following grammar
%
\begin{equation} \label{eq:apsynt}
  \begin{array}{rc*{4}{l@{\ |\ }}l}
    x & ::= & 0 & 1 & p \in P & x\cdot x & x + x \,,\\
  \end{array}
\end{equation}
%
where `$+$' and `$\cdot$' are binary operators, respectively called
{\em union} and {\em synchronisation}.  Synchronisation binds stronger
than union.
\end{syntax}
%
As follows from the interaction semantics given below, the additive
identity element $0$ represents blocking, since it does not authorise
any interaction.  The multiplicative identity element $1$ corresponds
to the empty interaction, which represents idling (see the discussion
in \cite{BarBliu15-offer-scico}).

\begin{semantics}
The semantics of $\ai(P)$ is given by the function $\|\cdot\| : \ai(P)
\rightarrow 2^{2^P}$, defined by
%
\begin{equation} \label{eq:apsem}
  \renewcommand{\arraystretch}{1.5}
  \begin{array}{lcl}
    \lefteqn{\|0\|\ =\ \emptyset,\quad \|1\|\ =\ \{\emptyset\},\quad
      \|p\|\ =\ \Big\{\{p\}\Big\},}\\
    \|x_1 + x_2\| &=& \|x_1\| \cup \|x_2\|,\\
    \|x_1 \cdot x_2\| &=& 
    \Big\{a_1 \cup a_2\,\Big|\, a_1\in \|x_1\|, a_2\in \|x_2\| \Big\},
  \end{array}
\end{equation}
for $p\in P$, $x_1,x_2\in\ai(P)$.  Terms of $\ai(P)$ represent sets of
interactions between the ports $P$.
\end{semantics}

A sound and complete axiomatisation of $\ai(P)$ with respect to the
semantic equivalence is provided in \cite{BliSif07-acp-emsoft}.  In a
nutshell, $(\ai(P), +, \cdot, 0, 1)$ is a commutative semiring,
idempotent in both $+$ and $\cdot$.

% *********************************************************************

\subsection{Algebra of Connectors}
\label{secn:ac}

The Algebra of Connectors provides an algebraic formalisation for
structuring the interaction models.  It underlies the graphical
notation (\eg \fig{connectors}) and the syntax for connectors used in the BIP language.

\begin{syntax}
The syntax of the {\em Algebra of Connectors}, $\ac(P)$, is defined by the
following grammar
%
\begin{equation} \label{eq:acsynt}
  \renewcommand{\arraystretch}{1.5}
  \begin{array}{lcll}
    s & ::= & [0]\ |\ [1]\ |\ [p]\ |\ [x] \:\:\:\:\:\:\:\:\:(synchrons)\\
    t & ::= & [0]'\ |\ [1]'\ |\ [p]'\ |\ [x]'  \:\:\:\:\:\:   (triggers)\\
    x & ::= & s\ |\ t\ |\ x\cdot x\ |\ x + x\,, 
  \end{array}
\end{equation}
for $p\in P$, and where `$+$' is a binary operator called {\em union},
`$\cdot$' is a binary operator called {\em fusion}, and brackets
`$[\cdot]$' and `$[\cdot]'$' are unary {\em typing} operators.  Fusion
binds stronger than union.

Union has the same meaning as union in $\ai(P)$. Fusion is a
generalisation of the synchronisation in $\ai(P)$.  Typing is used to
form typed connectors: `$[\cdot]$' defines {\em synchrons} (need
synchronisation with other ports in order to interact) and
`$[\cdot]'$' defines {\em triggers} (can initiate an interaction).
\end{syntax}

In order to simplify notation, we will omit brackets on 0, 1, and
ports $p \in P$, as well as `$\cdot$' for the fusion operation.

\begin{definition}
 In a system with a set of ports $P$, connectors are elements of $\ac(P)$.
\end{definition}

The operations of the Algebra of Connectors satisfy the following
axioms.
\begin{itemize}
\item Union `$+$' is associative, commutative, idempotent and has the identity element 0.
\item Fusion `$\cdot$' is associative, commutative and has the identity element 1. It is idempotent on monomial connectors, \ie[,] for any $x \in \ac(P)$, not involving the union operation, we have $x\cdot x = x$.
\item Typing `$[\cdot]^*$' satisfies the following axioms, for $x,y,z
  \in \ac(P)$ and $[\cdot]^\alpha,[\cdot]^\beta \in \bigl\{[\cdot]', [\cdot]\bigr\}$ arbitrary typings (trigger or synchron):
  \begin{enumerate}
  \item $[0] = [0]'$,
  \item $[[x]^\alpha]^\beta = [x]^\beta$,
  \item $[x + y]^\alpha = [x]^\alpha + [y]^\alpha$,
  \item $[x]'[y]' = [x]'[y] + [x][y]'$.
  \end{enumerate}
\end{itemize}
Complete axiomatisation of $\ac(P)$ with respect to the semantic
equivalence is provided in \cite{BliSif08-acp-tc}.

\begin{semantics}
The semantics of $\ac(P)$ is given by the function $|\cdot| : \ac(P)
\rightarrow \ai(P)$ (we use the $\sum$ and $\prod$ notation for,
respectively, the union and fusion of multiple terms of $\ac(P)$):
%
\begin{align}
  \label{eq:acap}
  |[p]| & = p\,,
  & |x_1 + x_2| & = |x_1| + |x_2|\,,
  & \left|\prod_{i=1}^n [x_i]\,\right| & = \prod_{i=1}^n |x_k|\,,
  \\
  \label{eq:acaptrig}
  && \hspace{-2cm}\left|\prod_{i=1}^n [x_i]' \prod_{j=1}^m [y_j]\,\right| = 
  \sum_{i=1}^n |x_i| & \lefteqn{\left(
  \prod_{k\not=i}\Big(1 + |x_k|\Big)\ \prod_{j=1}^m \Big(1 + |y_j|\Big)\right)}\,
\end{align}
%
for $n > 0$, $m \geq 0$, $x_1,\dots,x_n,y_1,\dots,y_m \in \ac(P)$ and
$p\in P\cup \{0,1\}$.
\end{semantics}

\fig{connectors} shows four basic examples of the graphical
representation of connectors.  Triggers are denoted by triangles,
whereas synchrons are denoted by bullets.  The interaction semantics
of the four connectors is given in the sub-figure captions.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.18\textwidth}
    %% \rule{\textwidth}{0.5pt}
    \centering
    \input{figures/rdv.pspdftex}
    \caption{\centering Rendezvous~$pqr$
      $\intsem{pqr} = \{pqr\}$}
    \label{fig:connectors:rdv}
  \end{subfigure}%
  \hfill %add desired spacing between images, e. g. ~, \quad, \qquad etc.
  %(or a blank line to force the subfigure onto a new line)
  \begin{subfigure}[b]{0.24\textwidth}
    %% \rule{\textwidth}{0.5pt}
    \centering
    \input{figures/broadcast.pspdftex}
    \caption{\centering Broadcast~$p'qr$
      $\intsem{p'qr} = \{p, pq, pr, pqr\}$}
    \label{fig:connectors:bdc}
  \end{subfigure}%
  \hfill %add desired spacing between images, e. g. ~, \quad, \qquad etc.
  %(or a blank line to force the subfigure onto a new line)
  \begin{subfigure}[b]{0.24\textwidth}
    %% \rule{\textwidth}{0.5pt}
    \centering
    \input{figures/atomic-bdc.pspdftex}
    \caption{\centering Atomic broadcast~$p'[qr]$
      $\intsem{p'[qr]} = \{p, pqr\}$}
    \label{fig:connectors:atomic}
  \end{subfigure}%
  \hfill %add desired spacing between images, e. g. ~, \quad, \qquad etc.
  %(or a blank line to force the subfigure onto a new line)
  \begin{subfigure}[b]{0.23\textwidth}
    %% \rule{\textwidth}{0.5pt}
    \centering
    \input{figures/causal-chain.pspdftex}
    \caption{\centering Causal chain~$p'[q'r]$
      $\intsem{p'[q'r]} = \{p, pq, pqr\}$}
    \label{fig:connectors:causal}
  \end{subfigure}%
  \caption{Basic connector examples}
  \label{fig:connectors}
\end{figure}

The semantic equivalence $\simeq$ is not a congruence for $\ac(P)$.
We denote by $\cong$ the largest congruence relation contained in
$\simeq$.

\begin{property}[\cite{BliSif08-acp-tc}]
  \label{prop:cong}
  For any $x, y \in \ac(P)$, we have $[x]'[y]' \cong \bigl[[x]'[y]'\bigr]'$.
\end{property}

%*********************************************************************

\subsection{Algebra of Causal Interaction Trees}
\label{secn:trees}

The Algebra of Causal Interaction Trees serves as a pivot for
transformations between all other algebraic representations.  It makes
explicit the causal dependencies between ports contributing to the
interactions defined by a connector.

\begin{syntax}
The syntax of the \emph{Algebra of Causal Interaction Trees}, $\ct(P)$, is
given by
\begin{equation} \label{eq:ctsyn}
  t ::= a \,|\, a \rightarrow t \,|\, t\oplus t\,,
\end{equation}
where $a \in \ai(P)$ is an interaction, \ie $0$, $1$ or a
synchronisation of ports (without the use of the union operator), and
`$\rightarrow$' and `$\oplus$' are respectively the {\em causality}
and the {\em parallel composition} operators.  Causality binds
stronger than parallel composition.  Notice that a causal interaction
tree can have several roots.
\end{syntax}

``Atomic'' strongly synchronised interactions in the nodes of a causal
interaction tree are the building blocks for the interactions provided
by the connector.  The causality operator defines a dependency between
two interactions: $a \rightarrow b$ means that for $b$ to
participate in the overall interaction, $a$ must also participate.
The parallel composition allows one to combine interactions without
introducing dependencies: any combination of $a$
and $b$ can participate in $a \oplus b$.

The causality operator is right- (but not left-) associative: for
interactions $a_1,\dots,a_n$, we have $a_1 \rightarrow (a_2
\rightarrow (\dots \rightarrow a_n)\dots)) = a_1 \rightarrow a_2
\rightarrow \dots \rightarrow a_n$.  We call this construction a {\em
  causal chain}.

\begin{semantics}
The semantics of $\ct(P)$ is given by the function $|\cdot|: \ct(P)
\rightarrow \ai(P)$
\begin{align}
  \label{eq:ctsem}
  |a| & = a\,,
  & |a \rightarrow t| & = a\Big(1 + |t|\Big)\,,
  & |t_1 \oplus t_2| & = |t_1| + |t_2| + |t_1|\,|t_2|\,,
\end{align}
where $a\in 2^P \cup \{0,1\}$ is an interaction and $t, t_1, t_2 \in \ct(P)$.
\end{semantics}

A sound axiomatisation of $\ct(P)$ is provided in
\cite{BliSif10-causal-fmsd}.

% *********************************************************************

\subsection{Systems of Causal Rules}
\label{secn:rules}

Systems of causal rules represent an intermediate structure between
causal interaction trees and arbitrary Boolean formulas.  They
directly encode the causality information explicit in the causal
interaction trees, by transforming causality relations into dual Horn
clauses. %%  Apart from supporting connector synthesis, they provide a
%% convenient way for expressing properties to be enforced by the glue
%% operators (see \secn{example} for some examples).  Causal rules
%% have also served as basis for the macro-notation used to specify
%% the glue in Dy-BIP\mdash a dynamic flavour of BIP \cite{dybip}.

\begin{definition}
  A {\em causal rule} is a Boolean formula in $\sB[P]$ of the form $E
  \Rightarrow C$.  The \emph{effect} $E$ is either the constant
  $\true$ or a port variable $p \in P$.  The \emph{cause} $C$ is
  either a constant, $\true$ or $\false$, or a disjunction of
  interactions, \ie $\bigvee_{i=1}^n a_i$ where, for all $i\in [1,n]$,
  $a_i$ are conjunctions of positive port variables.
\end{definition}

\begin{note} \label{rem:absorption}
  Notice that $a_1 \lor a_1\,a_2 = a_1$, and therefore causal rules
  can be simplified by replacing $p \Rightarrow a_1 \lor a_1\,a_2$
  with $p \Rightarrow a_1)$.  We assume that all the causal rules are
  simplified by this absorption rule.
\end{note}

\begin{definition}
  A \emph{system of causal rules} is a set $R = \{p \Rightarrow
  x_p\}_{p\in P\cup \{\true\}}$.  An interaction $a \in 2^P$ satisfies
  the system $R$ (denoted $a \models R$), iff the characteristic
  valuation of $a$ on $P$ satisfies the formula $\bigwedge_{p\in P\cup
    \{\true\}} (p \Rightarrow x_p)$.  We denote by
  $|R|\ \bydef{=}\ \sum_{a \models R} a$ the union (in terms of the
  Algebra of Interactions) of the interactions satisfying $R$.  Thus
  we have $|\cdot| : \cru(P) \rightarrow \ai(P)$, where $\cru(P)$ is
  the set of all systems of causal rules over the set of port
  variables $P$.
\end{definition}

%*********************************************************************

\subsection{Transformations between algebraic representations of interaction models}
\label{secn:transformations}

Transformations $\ac(P)
\overset{\tau}{\underset{\sigma}{\rightleftarrows}} \ct(P)$, $\ct(P)
\overset{R}{\rightleftarrows} \cru(P)$ and $\cru(P) \rightleftarrows
\sB[P]$ were defined in \cite{BliSif10-causal-fmsd} and have been
shown to respect $\simeq$.  Below, we recall only the transformations
that are used in this paper.

We define the function $\tau : \ac(P) \rightarrow \ct(P)$, associating
a causal interaction tree with a connector.  By \prop{cong}, and the
associativity of both fusion in $\ac(P)$ and parallel composition in
$\ct(P)$, the following equations are sufficient to define $\tau$:
%
\begin{align}
  \label{eq:contree:p}
  \tau(p) &= p
  \,,\\
  \label{eq:contree:trigsynch}
  \tau\left(
  [x]'\prod_{i=1}^n [y_i]
  \right) &=
  \tau(x) \rightarrow \bigoplus_{i=1}^n \tau(y_i)
  \,,\\
  \label{eq:contree:trig}
  \tau\bigl([x_1]'[x_2]'\bigr) &= \tau(x_1) \oplus \tau(x_2)
  \,,\\
  \label{eq:contree:sync}  
  \tau\bigl([y_1][y_2]\bigr) &=
  \bigoplus_{i = 1}^{m_1}\bigoplus_{j = 1}^{m_2} a_i^1 a_j^2
  \rightarrow (t_i^1 \oplus t_j^2)
  \,,
\end{align}
%
where $x, x_1, x_2, y_1,\dots, y_n \in \ac(P)$, $p \in P \cup \{0,1\}$
and, in \eq{contree:sync}, we assume $\tau(y_k) = \bigoplus_{i =
  1}^{m_k} a_i^k \rightarrow t_i^k$, for $k = 1,2$.

%% $\sigma : \ct(P) \rightarrow \ac(P)$ is defined recursively by putting
%% %
%% \begin{align}
%%   \label{eq:treecon}
%%   \sigma(a) & = [a]\,,
%%   & \sigma(a \rightarrow t) & = [a]'\,[\sigma(t)]\,,
%%   & \sigma(t_1 \oplus t_2)  & = [\sigma(t_1)]'\,[\sigma(t_2)]'\,.
%% \end{align}

We define $R: \ct(P) \rightarrow \cru(P)$ by putting
%
\begin{equation} 
  \label{eq:trees2rules}
  R(t)\ =\ \{p \Rightarrow c_p(t)\}_{p\in P\cup\{\true\}}\,,
\end{equation}
where the function $c_p : \ct(P) \rightarrow \sB[P]$ is defined recursively
as follows.  For $a\in 2^P$ (with $p\not\in a$) and $t,t_1,t_2 \in \ct(P)$,
we put
%
\begin{align*}
  c_p(0) & = \false\,, 
  & c_{\true}(0) & = \false\,,\\
  c_p(p \rightarrow t) & = \true\,,
  & c_{\true}(1 \rightarrow t) & = \true\,,\\
  c_p(pa \rightarrow t) & = a\,,\\
  c_p(a \rightarrow t) & = a \land c_p(t)\,,
  & c_{\true}(a \rightarrow t) & = a\,,\\
  c_p(t_1 \oplus t_2) & = c_p(t_1) \lor c_p(t_2)\,, 
  & c_{\true}(t_1 \oplus t_2) & = c_{\true}(t_1) \lor c_{\true}(t_2)\,.
\end{align*}

Observe that this transformation associates to each port $p \in P$ a
causal rule $p \Rightarrow C$, where $C$ is the disjunction of all
prefixes leading from roots of $t$ to some node containing $p$,
including the ports of this node other than $p$.

\end{document}