### 2019-04-26 2018-Arch-pNets/Coordination sources/arch-pNets.tex
arch-pNets paper: Prepared the sources for submission -- removed all 'Overfull's

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
\documentclass{llncs}
\usepackage{llncsdoc}
\usepackage[T1]{fontenc}

%\usepackage{a4wide}

\usepackage{graphicx}
\usepackage{verbatim}

\let\proof\relax   
\let\endproof\relax

\usepackage{amsmath,amsthm,amscd}
\usepackage{amssymb}    % for blackboard B
\usepackage{rotating}   % to rotate the figures
\usepackage{mathpartir} % math paragraph + inference rules
\usepackage{mathtools}  % for align, multline etc.
\usepackage{caption}
\usepackage{stmaryrd} % double brackets (\llbracket, \rrbracket)
\usepackage{units}    % for nicefrac: x/y

%% \usepackage[justification=centering,belowskip=-10pt,aboveskip=0pt]{caption}
%% \setlength{\intextsep}{10pt plus 2pt minus 2pt}
%% \setlength\abovedisplayskip{0pt}

\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage{hyperref}

\input{figures/macros.tex}         % Definitions of macros used in XFig figures

\usepackage{soul}
\soulregister\cite7
\soulregister\ref7
\soulregister\pageref7
\soulregister{\em}{0}
\soulregister{\mdash}{0}
\soulregister{\st}{7}
\soulregister{\bf}{0}
\soulregister{\fig}{7}
\soulregister{\ie}{0}
\soulregister{\etc}{0}
\soulregister{\eg}{0}
\soulregister{\Eg}{0}
\soulregister{\wrt}{0}
\soulregister{\resp}{0}
\soulregister{\cf}{0}

\soulregister{\cstate}{7}
\soulregister{\tstate}{7}
\soulregister{\goesto}{7}

\usepackage[colorinlistoftodos,bordercolor=white]{todonotes}
%% \usepackage[disable,colorinlistoftodos,bordercolor=white]{todonotes}
\usepackage{macrospNets}

\newcommand{\Simon}{\\\hfill\mdash Simon}
\newcommand{\Eric}{\\\hfill\mdash Eric}

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{#2}\Simon}}
\newcommand{\noteSBin}[2][inline,color=green!40]{\todo[#1]{{#2}\Simon}}
\newcommand{\todoSB}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Simon:} {#2}}}
\newcommand{\todoSBin}[2][inline,color=green!40]{\todo[#1]{\textbf{To-do Simon: } {#2}}}
\newcommand{\Ludo}{\\\hfill\mdash Ludo}
\newcommand{\noteLH}[2][color=orange!40, size=\tiny]{\todo[#1]{{#2}\Ludo}}
\newcommand{\noteLHin}[2][inline,color=orange!40]{\todo[#1]{{#2}\Ludo}}
\newcommand{\todoLH}[2][color=orange!40, size=\tiny]{\todo[#1]{\textbf{To-do Ludo:} {#2}}}
\newcommand{\todoLHin}[2][inline,color=orange!40]{\todo[#1]{\textbf{To-do Ludo: } {#2}}}
\newcommand{\noteEM}[2][color=blue!40, size=\tiny]{\todo[#1]{{#2}\Eric}}
\newcommand{\todoEM}[2][color=blue!40, size=\tiny]{\todo[#1]{\textbf{To-do Eric:} {#2}}}
\newcommand{\noteEMin}[2][inline,color=green!40]{\todo[#1]{{#2}\Eric}}

\newcommand{\noteIn}[2][inline,color=black!20]{\todo[#1]{{#2}}}

\newcommand{\add}[2][Added]{\todo[color=blue!20, size=\tiny]{#1}{\color{blue}#2}}
%% \newcommand{\remove}[2][Removed]{\todo[color=red!20, size=\tiny]{#1}}
\newcommand{\remove}[2][Removed]{\todo[color=red!20, size=\tiny]{#1}{\color{red}\st{#2}}}
\newcommand{\removeIn}[2][Removed]{\todo[color=red!10, inline]{{\color{red}#2}#1}}
%% \newcommand{\replace}[3][Replaced]{\todo[color=blue!20, size=\tiny]{#1}{\color{blue}#3}}
\newcommand{\replace}[3][Replaced]{\todo[color=blue!20, size=\tiny]{#1}{\color{blue}#3}{\color{red}\st{#2}}}

\newcommand{\addSB}[1]{\add[Added by Simon]{#1}}
\newcommand{\removeSB}[1]{\remove[Removed by Simon]{#1}}
\newcommand{\removeSBin}[1]{\removeIn[Removed by Simon]{#1}}
\newcommand{\replaceSB}[2]{\replace[Replaced by Simon]{#1}{#2}}

\newcommand{\newCoord}[1]{{\color{blue}#1}}

\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\newcommand{\tupleDeli}{(}
\newcommand{\tupleDelii}{)}
\newcommand{\setTupleDelims}[2][(]{
  \renewcommand{\tupleDeli}{#1}%
  \ifx#2\relax\else\renewcommand{\tupleDelii}{#2}\fi%
}
\newcommand{\tuplebase}[2][\ensuremath{,\allowbreak}]{%  
  \def\nextitem{\def\nextitem{#1}}% Separator
  \renewcommand*{\do}[1]{\nextitem ##1}% How to process each item
  \tupleDeli\docsvlist{#2}\tupleDelii% Process list
}

\newcommand{\tuple}[2][\ensuremath{,\allowbreak}]{%
  \setTupleDelims[(]{)}%
  \tuplebase[#1]{#2}%
}

\newcommand{\btuple}[2][\ensuremath{,\allowbreak}]{%
  \setTupleDelims[\bigl(]{\bigr)}%
  \tuplebase[#1]{#2}%
}

\newcommand{\Tuple}[2][\ensuremath{,\allowbreak}]{%
  \setTupleDelims[\Big(]{\Big)}%
  \tuplebase[#1]{#2}%
}

\newcommand{\pNetTuple}[2][\ensuremath{,\allowbreak}]{%
  \setTupleDelims[\mylangle]{\myrangle}%
  \tuplebase[#1]{#2}%
}

\newcommand{\SVTuple}[2][\ensuremath{,\allowbreak}]{%
  \setTupleDelims[\langle]{\rangle}%
  \tuplebase[#1]{#2}%
}

\newcommand{\listset}[2][\ensuremath{,\allowbreak}]{%
  \setTupleDelims[\{]{\}}%
  \tuplebase[#1]{#2}%
}

\newcommand{\defn}[1]{Def.~\ref{defn:#1}}
\newcommand{\defntwo}[2]{Defs~\ref{defn:#1} and \ref{defn:#2}}
\newcommand{\fig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\Fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\figs}[2]{Fig.~\ref{fig:#1} and~\ref{fig:#2}}
\newcommand{\tab}[1]{Tab.~\ref{tab:#1}}
\newcommand{\eq}[1]{(\ref{eq:#1})}
\newcommand{\res}[1]{(\ref{res:#1})}
\newcommand{\ex}[1]{Ex.~\ref{ex:#1}}
\newcommand{\exs}[2]{Ex.~\ref{ex:#1} and~\ref{ex:#2}}
\newcommand{\secn}[1]{Sect.~\ref{secn:#1}}
\newcommand{\Secn}[1]{Section~\ref{secn:#1}}
\newcommand{\app}[1]{App.~\ref{secn:#1}}
\newcommand{\rem}[1]{Note~\ref{rem:#1}}
\newcommand{\lem}[1]{Lem.~\ref{lem:#1}}
\newcommand{\cor}[1]{Cor.~\ref{cor:#1}}
\newcommand{\thm}[1]{Th.~\ref{thm:#1}}
\newcommand{\Thm}[1]{Theorem~\ref{thm:#1}}
\newcommand{\axs}[1]{Ax.~\ref{ax:#1}}
% \newcommand{\ax}[2]{\ref{ax:#1}.\ref{ax:#1:#2}}
\newcommand{\ax}[1]{Ax.~\ref{ax:#1}}
\newcommand{\prop}[1]{Prop.~\ref{prop:#1}}
\newcommand{\alg}[1]{Alg.~\ref{alg:#1}}
\newcommand{\asmp}[1]{Ass.~\ref{asmp:#1}}

% %%%%%%%%%%%%%%%%%%%%%

\newcommand{\cA}{\ensuremath{\mathcal{A}}}
\newcommand{\fA}{\ensuremath{\mathsf{A}}}
\newcommand{\sA}{\ensuremath{\mathbb{A}}}
\newcommand{\cB}{\ensuremath{\mathcal{B}}}
\newcommand{\fB}{\ensuremath{\mathsf{B}}}
\newcommand{\sB}{\ensuremath{\mathbb{B}}}
\newcommand{\cC}{\ensuremath{\mathcal{C}}}
\newcommand{\sC}{\ensuremath{\mathbb{C}}}
\newcommand{\cD}{\ensuremath{\mathcal{D}}}
\newcommand{\fD}{\ensuremath{\mathsf{D}}}
\newcommand{\sD}{\ensuremath{\mathbb{D}}}
\newcommand{\cE}{\ensuremath{\mathcal{E}}}
\newcommand{\sE}{\ensuremath{\mathbb{E}}}
\newcommand{\cF}{\ensuremath{\mathcal{F}}}
\newcommand{\cG}{\ensuremath{\mathcal{G}}}
\newcommand{\cH}{\ensuremath{\mathcal{H}}}
\newcommand{\cI}{\ensuremath{\mathcal{I}}}
\newcommand{\sI}{\ensuremath{\mathbb{I}}}
\newcommand{\cM}{\ensuremath{\mathcal{M}}}
\newcommand{\cN}{\ensuremath{\mathcal{N}}}
\newcommand{\sN}{\ensuremath{\mathbb{N}}}
\newcommand{\cP}{\ensuremath{\mathcal{P}}}
\newcommand{\sP}{\ensuremath{\mathbb{P}}}
\newcommand{\cQ}{\ensuremath{\mathcal{Q}}}
\newcommand{\sQ}{\ensuremath{\mathbb{Q}}}
\newcommand{\cR}{\ensuremath{\mathcal{R}}}
\newcommand{\sR}{\ensuremath{\mathbb{R}}}
\newcommand{\cS}{\ensuremath{\mathcal{S}}}
\newcommand{\sS}{\ensuremath{\mathfrak{S}}}
\newcommand{\cT}{\ensuremath{\mathcal{T}}}
\newcommand{\fT}{\ensuremath{\mathsf{T}}}
\newcommand{\sT}{\ensuremath{\mathbb{T}}}
\newcommand{\cV}{\ensuremath{\mathcal{V}}}
\newcommand{\fV}{\ensuremath{\mathsf{V}}}
\newcommand{\sV}{\ensuremath{\mathbb{V}}}
\newcommand{\sZ}{\ensuremath{\mathbb{Z}}}

\newcommand{\mdash}[1][]{---#1}
\newcommand{\ndash}{--}
\newcommand{\ie}[1][\ ]{i.e.#1}
\newcommand{\etc}[1][\ ]{etc.#1}
\newcommand{\eg}[1][\ ]{e.g.#1}
\newcommand{\Eg}[1][\ ]{E.g.#1}
\newcommand{\cf}[1][\ ]{cf.#1}
\newcommand{\wrt}[1][\ ]{w.r.t.#1}
\newcommand{\Wrt}[1][\ ]{W.r.t.#1}
\newcommand{\resp}[1][\ ]{resp.#1}

\newcommand{\bydef}[1]{\ensuremath{\stackrel{\mathit{\scriptscriptstyle def}}{#1}}}
\newcommand{\suchthat}{\ensuremath{\,|\,}}
\newcommand{\rightsuchthat}{\ensuremath{\,\right|\,}}
\newcommand{\leftsuchthat}{\ensuremath{\,\left|\,}}
\newcommand{\setdef}[2]{\ensuremath{\{{#1}\,|\,{#2}\}}}
\newcommand{\bsetdef}[2]{\ensuremath{\bigl\{{#1}\,\bigl|\,{#2}\bigr.\bigr\}}}
\newcommand{\Setdef}[2]{\ensuremath{\Big\{{#1}\,\Big|\,{#2}\Big\}}}
\newcommand{\goesto}[2][]{\ensuremath{\xrightarrow[{#1}\relax]{#2}}}
\newcommand{\notgoesto}[2][]{\ensuremath{\not\xrightarrow[{#1}\relax]{\ \,#2}}}
\newcommand{\non}[1]{\ensuremath{\overline{#1}}}

\newcommand{\true} {\ensuremath{\mathtt{t\!t}}}
\newcommand{\false}{\ensuremath{\mathtt{f\!f}}}
\newcommand{\noop} {\ensuremath{\emptyset}} % \mathsf{skip}
\newcommand{\init} {\ensuremath{\mathsf{init}}}

\newcommand{\order}{\leqslant}
\newcommand{\ordbool}{\ensuremath{\sB^{\order}}}
\newcommand{\data}{\ensuremath{\sD}}
\newcommand{\signature}{\ensuremath{\Sigma}}
\newcommand{\variables}{\ensuremath{\cV}}
\newcommand{\Talg}{\ensuremath{\cT_{\signature,\variables}}}
\newcommand{\actions}[1]{\ensuremath{\mathit{Act}_{#1}}}
\newcommand{\exprs}[1]{\ensuremath{\sE_{#1}}}
\newcommand{\monexprs}[1]{\ensuremath{\sE^{\order}_{#1}}}
\newcommand{\boolexprs}[1]{\ensuremath{\sB_{#1}}}
\newcommand{\guards}[1]{\ensuremath{\ordbool_{#1}}}
\newcommand{\assigns}[1]{\ensuremath{\sA_{#1}}}
\newcommand{\updates}[1]{\ensuremath{\sA^{\order}_{#1}}}
\newcommand{\valuations}[1]{\ensuremath{\data^{#1}}}
\newcommand{\val}[3][]{\ensuremath{#1{\sigma}^{#2}_{#3}}}

%% CTL
\newcommand{\AG}[1][\ ]{\ensuremath{\mathtt{AG}#1}}
\newcommand{\AF}[1][\ ]{\ensuremath{\mathtt{AF}#1}}
\newcommand{\EF}[1][\ ]{\ensuremath{\mathtt{EF}#1}}
\newcommand{\EX}[1][\ ]{\ensuremath{\mathtt{EX}#1}}
\newcommand{\AW}[3][\ ]{\ensuremath{\mathtt{A}\left[#2\ \mathtt{W}\ #3\right]#1}}
\newcommand{\EU}[3][\ ]{\ensuremath{\mathtt{E}\left[#2\ \mathtt{U}\ #3\right]#1}}
\newcommand{\EW}[3][\ ]{\ensuremath{\mathtt{E}\left[#2\ \mathtt{W}\ #3\right]#1}}

\newcommand{\filter}[2][]{\ensuremath{\pi_{#1}({#2})}}

\newcommand{\primeit}[1]{#1'}
\newcommand{\secondit}[1]{#1''}

\newcommand{\twosynch}{%
  \mbox{\ensuremath{\bullet\!\!\!-\!\!\!-\!\!\!-\!\!\!\bullet}}}
\newcommand{\synchtrig}{%
  \mbox{\ensuremath{\bullet\!\!\!-\!\!\!-\!\!\!\blacktriangleleft}}}
\newcommand{\trigsynch}{%
  \mbox{\ensuremath{\blacktriangleright\!\!\!-\!\!\!-\!\!\!\bullet}}}
\newcommand{\twotrig}{%
  \mbox{\ensuremath{\blacktriangleright\!\!\!-\!\!\!-\!\!\!\blacktriangleleft}}}

\newcommand{\export}[1][]{\ensuremath{\varepsilon_{#1}}}
\newcommand{\valdiff}[2]{\ensuremath{#1 \triangle #2}}
\newcommand{\supp}[1]{\ensuremath{\mathrm{supp}(#1)}}
\newcommand{\semopen}[1]{\ensuremath{[{#1}]}}
\newcommand{\semclosed}[1]{\ensuremath{\llbracket{#1}\rrbracket}}
\newcommand{\reachable}[1]{\ensuremath{\mathit{reachable}({#1})}}
\newcommand{\IMextend}[2]{\ensuremath{#1 \ltimes #2}}
\newcommand{\arcomp}{\oplus}
\newcommand{\arequiv}{\equiv}
\newcommand{\Arcomp}{\Bigoplus}
\newcommand{\expmix}{\wedge}

\newcommand{\intsem}[1]{\ensuremath{\|{#1}\|}}
\newcommand{\aisem}[1]{\ensuremath{|{#1}|}}

\newtheorem*{syntax}{Syntax}
\newtheorem*{semantics}{Semantics}

\newcommand{\ai}{\ensuremath{\mathcal{AI}}}
\newcommand{\ct}{\ensuremath{\mathcal{T}}}
\newcommand{\cru}{\ensuremath{\mathcal{CR}}}
\newcommand{\ac}{\ensuremath{\mathcal{AC}}}

\newcommand{\addspace}[1]{#1 }
\newcommand{\nopri}[1][]{\ensuremath{\mathit{enc}%
    \ifx#1\relax\else(#1)\fi}}
\newcommand{\maxprog}[1][]{\ensuremath{\mathit{enc}_\mu%
    \ifx#1\relax\else(#1)\fi}}

\newcommand{\partition}{\cD}

\makeatletter
\newcommand{\doubletilde}[1]{{%
  \mathpalette\double@tilde{#1}%
}}
\newcommand{\double@tilde}[2]{%
  \sbox\z@{$\m@th#1\tilde{#2}$}%
  \ht\z@=.9\ht\z@
  \tilde{\box\z@}%
}
\makeatother

\newcounter{tempctr}
\newcommand{\breakenumistart}{%
  \setcounter{tempctr}{\value{enumi}}%
  \end{enumerate}%
}
\newcommand{\breakenumiend}{%
  \begin{enumerate}%
  \setcounter{enumi}{\value{tempctr}}%
}
\setlength{\topsep}{0pt}%

% %%%%%%%%%%%%%%%%%%%%%
\makeatletter
\newcommand{\verbatimfont}[1]{\renewcommand{\verbatim@font}{\ttfamily#1}}
\makeatother
% %%%%%%%%%%%%%%%%%%%%%

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{arrows,shapes,automata,petri}
  \tikzset{
  place/.style={
    circle,
    thick,
    draw=blue!75,
    fill=blue!20,
    minimum size=6mm
  },
  transition/.style={
    rectangle,
    thick,
    fill=black,
    minimum width=8mm,
    inner ysep=2pt
    },
    arc/.style = {
      decoration=
      {markings,mark=at position #1 with {circle;}
      },
      postaction={decorate,draw}},
  }

\let\llncssubparagraph\subparagraph
\let\subparagraph\llncssubparagraph

\begin{document}
\graphicspath{{figures/}}

\title{Verification of concurrent design patterns with data}

\author{%
Simon~Bliudze\inst{1}
\and
Ludovic~Henrio\inst{2}
\and
Eric~Madelaine\inst{3}
}

\institute{%
  Inria Lille -- Nord Europe, Villeneuve d'Ascq, France \\
  \email{simon.bliudze@inria.fr}
  \and
	Univ Lyon, EnsL, UCBL, CNRS, Inria,  LIP, F-69342, LYON Cedex 07, France\\ \email{ludovic.henrio@ens-lyon.fr}
  \and
    Universit\'e C\^ote d'Azur, Inria, CNRS, I3S, 06902 Sophia-Antipolis, France\\
  \email{eric.madelaine@inria.fr}
}

\maketitle

\begin{abstract}
We provide a solution for the design of safe concurrent systems by
compositional application of verified design patterns\mdash 
called \emph{architectures}\mdash to a small set of functional
components.  To this end, we extend the theory of architectures
developed previously for the BIP framework with the elements necessary
for handling data: definition and operations on data domains, syntax
and semantics of composition operators involving data transfer.
%
\newCoord{We provide a set of conditions under which} %% property preservation result holds that states that
%% %
%% under
%% monotonicity assumptions on guards and assignments and, in presence of the \emph{maximal progress} special case of
%%   priority models, additional upwards compatibility assumptions on the architectures,}
composition of
architectures preserves their characteristic safety properties.
To verify that individual architectures do enforce their associated
properties, we provide an encoding into open pNets, an intermediate
model  that supports SMT-based verification.
%% Additionally, the translation into  open pNet  \noteEM{ICI}{gives
%% us a strong insight and a comparison of the coordination mechanisms
%% provided by the two frameworks}.
The approach is illustrated by a case
study based on a previously developed BIP model of a nanosatellite
on-board software.

\keywords{Symbolic verification, composition, safety, interaction
  models.}
\end{abstract}

%****************************************************************
%****************************************************************

\section{Introduction}
\label{secn:introduction}

BIP (Behaviour-Interaction-Priority)~\cite{bip} is a framework for the
component-based design of concurrent software and systems.  In
particular, the BIP tool-set comprises compilers for generating C/C++
code, executable by linking with one of the dedicated engines, which
implement the BIP operational semantics~\cite{BliSif08-acp-tc}.
BIP ensures that any property that holds on a BIP
model will also hold on the generated code.
The notion of BIP \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified
%% \footnote{%
%% %
%%   The current approach relies on taxonomies of predefined
%%   architectures.  Future work may include automatic synthesis of
%%   architectures from, \eg CTL~\cite{BaierKatoen2008} formulae.
%% %
%% }
and applied to the model, adding
coordinator components and modifying the 
synchronisation patterns between components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie if two architectures guarantee two properties, their composition ensures the conjunction of the properties \newCoord{but~\cite{AttieBBJS16-architectures-faoc} did not consider properties depending on data}.
%
%several architectures are applied, each enforcing a safety
%property, the resulting system satisfies their conjunction.

This article goes one step further in the proof of properties and in the compositionality of architectures, but this step is a significant one: the compositional verification. To prove properties of BIP architectures
it is necessary to have a representation of the BIP architecture in a verifiable format. 
The verification problem has two unbounded parameters: 1) By nature, architectures have holes and are meant to interact with the interfaces of the component that will fill the hole; the properties must hold for all (well-typed) components that can be put inside the hole; 2) BIP interactions can transmit data, and properties might be dependent of the data, the domain of the data is generally huge or unbounded and the values of transmitted data might have a significant impact on the properties.
%% Due to the nature of the problem, 
We propose to rely on a translation of BIP architectures into \emph{open pNets}.

Parameterised Networks of synchronised automata (pNets) is a formalism for defining behavioural specification of distributed systems based on a parameterised and hierarchical model.
It inherited from the work of
Arnold on synchronisation vectors~\cite{Arnold1982}. 
It has been shown in previous work~\cite{HMZ:PDP15} that pNets can  represent the behavioural semantics
of a system 
including value-passing and many kinds of synchronisation methods, including various constructs and
languages for distributed objects. The VerCors  platform uses pNets to design
and verify  distributed software components
\cite{CM:FMCO08,HKM-FASE16}. 
There is no bound on the number of elements inside a pNets or the valuation of parameters. When restricted to finite instantiations, it becomes possible to us pNets  for finite model-checking approaches.
Closed pNets were used to encode fully defined programs or systems,
while open pNets have ``holes'', playing the role of process
parameters. Such open systems can represent composition operators or structuring architectures.
It is possible to reason, in an SMT engine, on the symbolic automaton that represents the behaviour of a pNets with holes and that communicates values~\cite{HMZ-FORTE2016QBMZ-AVOCS18}. The encoding of open pNets into Z3 that is under development is the starting point of this article. We benefit from the possibility to reason on a pNet in an SMT engine in order to prove properties on BIP architectures.

%% We propose   to encode BIP architectures
%% into open pNets and to verify, using SMT techniques,
%% the properties of the architectures. The encoding also allows us to
%% compare BIP architectures and  pNet formalism.  
%% Overall, the encoding allows us to precisely compare the coordination
%% mechanisms of the two models but also the way component composition is
%% expressed in the two formalisms.

%This also gives the opportunity
%to relate two  different visions on compositional verification.

%\noteLH{pas tres heureux la fin:
%  pas assez detaille ou trop complique -> abstraire ou detailller a
%  mon avis (Ludo). J'ai prefere abstraire/raccourcir... (Eric)}

%\paragraph{Contributions}
The main contributions of this paper are: 1) \newCoord{The addition of data to the
theory of BIP architectures, including a theorem about
preservation of data dependent properties by 
compositions.} 2) An encoding of architectures with data into  open pNets, allowing for analysis of their temporal properties using pNet's
software tools.  
The  paper is illustrated by a running
example based on the failure monitor architecture from the CubETH
nanosatellite on-board software~\cite{CubETH-case-study}.
%
This running example also relies on the \emph{maximal progress}
assumption, whereby larger interactions are preferred to smaller
ones.  Due to space limitations, we only discuss this
informally. However, proofs of the results provided in the
appendix formally account for maximal progress.

The rest of the paper is structured as follows.  In
\secn{preliminaries}, we present notations and background
material on pNets.  The theory of architectures with data is
presented in \secn{archs}.  In \secn{encoding}, we present the
encoding into open pNets and discuss verification of the running
example.  \Secn{related} discusses related work.
\Secn{conclusion} concludes the paper.

%****************************************************************
%****************************************************************

\section{General Notations and pNets Previous Results} 
\label{secn:preliminaries}

%****************************************************************

\subsubsection{Notations.}
\label{secn:notations}

We extensively use indexed structures
over some countable indexed sets, which are equivalent to mappings over
the countable set. 
Thus, 
$a_i^{i\in I}$
%, or equivalently  $(i\mapsto a_i)^{i\in I}$
denotes a family of elements $a_i$ indexed over the
set $I$. % Such a family
% is equivalent to the mapping $(i\mapsto a_i)^{i\in I}$.
% To specify the set over which the structure is indexed,
% indexed structures are always denoted with an exponent of the form $i\in I$
% (arithmetic only appears in the indexes if necessary).
This notation defines both $I$ the set over which the family is
indexed (called \emph{range}), and $a_i$ the elements of the family.
E.g., $a^{i\in\{3\}}$ is the mapping with a single entry $a$ at index
$3$ ; also abbreviated $(3\mapsto a)$.
When this is not
ambiguous, we shall use notations for sets, and typically write
``indexed set over I'', even though formally we should speak of maps; and
write $x\in a_i^{i\in I}$ to mean $\exists i\in I.\, x=a_i$.  An empty
family is denoted $\emptyset$.

%\noteIn{states looks ok, but initial state is $s_0$ in pNet and not in archi : $s^0$.  Exponent 
%position was removed from pNet because we found it complex with indexed set notation but 
%could be ok.
%\Ludo
%
%I am using exponent notation for ``modifiers'', reserving\mdash
%as much as possible\mdash index notation for enumeration.  I
%think this can be solved, where necessary, by using parentheses,
%\eg $(s^0_i)^{i \in I}$
%\Simon
%}

%****************************************************************
We assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$, $e$ ranges over  expressions; and a set of \emph{Boolean
expressions} 
$\boolexprs{\variables}\subseteq\exprs{\variables}$, $g$ (guards) ranges over Boolean expressions.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$.
We define $\assigns{\variables}$ as the set of variable assignments of the form: $(x_i := e_i)^{i \in I}$ and let 
%\[
%\assigns{\variables} \bydef{=}
%\Setdef{(x_i := e_i)^{i \in I}}{
%  x_i^{i \in I}\subseteq \variables,
%  e_i^{i \in I}\subseteq \exprs{\variables},
%  I \text{ is a finite index set}
%}
%\]
%to denote the set of \emph{variable assignments}. 
 $u$ range over sets of assignments.
%To be able to reason about the data flow between actions, we
%distinguish \emph{input variables} of the form $?x$ within terms; t
The function
$\vars(t)$ identifies the set of variables in a term.
%$t\in\Talg$, and $iv(t)$ returns its input variables.

We assume the existence of a universal data domain
given as a partially-ordered set $(\data, \order)$, potentially
encompassing several copies of any given data type with
different orders.  We assume that $(\data,
\order)$ comprises both the unordered set of Booleans $\sB =
\tuple{\{\true, \false\}, \emptyset}$ and the naturally ordered one
$\sB^\order = \tuple{\{\true, \false\}, \{\false \order \true\}}$, and
similarly for integer and real numbers; as well as the set of
intervals ordered by inclusion.
%In practice, and in the examples of
%this paper, all the variables are sorted, meaning that their values
%can only range over given subsets of $\data$, such as $\sB$ or
%$\sB^\order$.
%
When speaking of an ordered sort, \eg $\sB^\order$, we will assume that it
forms a meet-semilattice and denote by $\wedge$ the meet
operator.

For a set of variables $V \subseteq \variables$, we denote
$\valuations{V} \bydef{=} \{\val{}{}: V \rightarrow \data\}$
the set of \emph{valuations} of the variables in $V$ and let $\sigma$ range over valuations.
Valuations extend canonically to expressions, denoted $\val{}{}(e)$. We define:
%\noteLH{its the following really necessary}
%\addSB{Since we consider variables to be sorted, valuations
%  applied to expressions are interpreted as partial functions
%  $\exprs{V} \rightharpoonup \data$.}
%For a valuation $\val{}{}$ and an assignment $(x_i := e_i)^{i \in I}$, we define:
%% \vspace{-1.5ex}
\[
\val{}{}\bigl[(x_i := e_i)^{i \in I}\bigr](x) \bydef{=}
\begin{cases}
  \val{}{}(x), & \text{if } x \not\in x_i^{i \in I}\,,\\
  \val{}{}(e_i), & \text{if } x = x_i, \text{for some } i \in I\,.
\end{cases}%% \vspace{-1ex}
\]
For two valuations $\val{1}{}, \val{2}{} : V \rightarrow
\data$, we denote $\valdiff{\val{1}{}}{\val{2}{}} \bydef{=}
\bsetdef{x \in V}{\val{1}{}(x) \neq \val{2}{}(x)}$ the set of
variables that are assigned different values by the two
valuations.  As usual, we write $\val{1}{} \order \val{2}{}$
iff $\val{1}{}(x) \order \val{2}{}(x)$, for all $x \in V$.  An
expression $e$ is  \emph{monotonic} if, for
any two valuations $\val{1}{}, \val{2}{}$,
$\val{1}{} \order \val{2}{}$ implies $\val{1}{}(e) \order
\val{2}{}(e)$.
%
Similarly, an assignment $(x_i := e_i)^{i \in I} $ is
 monotonic if all expressions $e_i^{i \in I}$ are monotonic.
We  denote $\guards{V} \subset \boolexprs{V}$, $\monexprs{V}
\subset \exprs{V}$ and $\updates{V} \subset \assigns{V}$ the sets of
monotonic Boolean and generic expressions and assignments,
respectively.
%
%****************************************************************

\subsubsection{Open pNets.}
\label{secn:pNets}
This section briefly describes pNets, see~\cite{HMZ-FORTE2016} for more complete   description.
pNets are tree-like structures, where the leaves are either
\emph{parameterised labelled transition systems (pLTSs)}, expressing the
behaviour of basic processes, or \emph{holes}, used as placeholders
for unknown processes.
Nodes of the tree are synchronising artefacts using a
set of \emph{synchronisation vectors} that express the possible
synchronisation between  parameterised actions of some components.

%\paragraph{Definitions.}
%\label{section:pNets}
A pLTS is a labelled transition system with variables occurring inside states,  actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
Variables of each state are pairwise disjoint.
\newCoord{Each transition label of a pLTS consists of a parameterised action, a guard and an update assignment.  The parameters of actions are either input variables or expressions.  Input variables are bound when the action occurs; they accept any value (of the correct type), thus providing a to input data from the environment.  Expressions are computed from the values of other variables.  They allow providing aggregated values to the environment, without exposing all the underlying variables.}
We  define the set of parameterised actions a pLTS can use ($a$
ranges over action labels):
% Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
e_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $e_j^{j \in J}$ 
are expressions.

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
%\begin{itemize}
%\item
$S$ is a set of states; %$\vars(s)$ denotes the set of associated variables;
%\item
$s_0 \in S$ is the initial state;
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
%\item 
${\to} \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
% SB: I removed a new paragraph here, ok?
$\langle \alpha,~g,~u\rangle$,
where $\alpha$ is a parameterised action, $\alpha \in\actions{\variables}$; 
$g\in\boolexprs{\variables}$ is a guard over variables of the source state and the 
action, and $u\in\assigns{\variables}$
assigns updated value for variables in the destination state. 
%%% SHORTEN BY LUDO
%If 
%$s \xrightarrow{\langle \alpha,~g,~u\rangle} s'\in \to $ then 
%% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
%		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! \vars(s)$, 
%		$\vars(g)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
%		$\vars(\codom(u))\!\subseteq\! \vars(s)\cup\vars(\alpha)\land 
%		\vars(\dom(u))\!\in\!\vars(s')$. %,  and $s= s'\Rightarrow J=\emptyset$. 
%\noteLH{is dom and codom clear?}
%\noteSB{Not sure: I understand, because I know what is meant, but this is not a canonical use, so should be explained. Will be solved if we roll back to \mbox{$(?x_i := e_i)^{i \in I}$} or \mbox{$\overline{?x := e}$}.}
%\noteLH{COORDINATION: I believe we should remove the constraints}		
%\end{itemize}
\end{definition}
%Now we define
%pNets as constructors for hierarchical behavioural structures.
A pNet composes several pNets, pLTSs, and holes.
A  pNet  exposes
 global actions resulting from the synchronisation of internal actions in some sub-pNets, and some actions of the holes. \newCoord{As holes are process parameters, synchronisation with a hole has an obvious concrete meaning when a process is put inside the hole and emits the action. We  also define a semantics for open pNets with holes where open transitions express the fact that a pNet can performs a transition provided one or several holes emit some actions.}
\newCoord{This synchronisation  is specified by  \emph{synchronisation vectors} expressing the synchronous interaction between actions inside sub-pNets and holes, data transmission is expressed classically using action parameters.}
Actions involved in the synchronisation vectors do
not need to distinguish input variables, \ie they 
have the form $a(\Expr_j^{j \in J})$.

\begin{definition}[pNets]\label{defn:pnets}
A pNet is a hierarchical structure where leaves are pLTSs and holes: %\footnote{Typing can 
%be added to pNets by defining and checking their interfaces, called 
%sort~\cite{HMZ-FORTE2016}}:
$Q\triangleq \pLTS~|~\mylangle Q_i^{i\in I}, J, \symb{SV}_k^{k\in 
K}\myrangle$
where\\[-3.5ex]
\begin{itemize}
\item $Q_i^{i\in I}$ is the family of sub-pNets;
%  $Q_i^{i\in I}$ is a family of sub-pNets where $I\in\I_\P$ is the set over which 
%sub-pNets are indexed.

\item $J$ is a set of indexes, called \emph{holes}.
$I$ and $J$ are \emph{disjoint}: $I\!\cap\! J=\emptyset$,  $I\!\cup\! J\neq\emptyset$
%\item[$\bullet$] $\Sort_j \subseteq \AlgAS$ is a set of action terms, denoting the 
%\emph{sort} of
%hole $j$.

\item $\symb{SV}_k^{k\in K}$ is a set of
  synchronisation vectors. % (\hl{$K\in\I_\P$}). 
$\forall k\!\in\! K,
  \symb{SV}_k\!=\SV{\!\alpha_{l}^{l\in I_k \uplus J_k}}{\alpha'_k}{g_k}$, where
  $\alpha'_k\in \actions{\variables}$, $I_k\subseteq I$, $J_k\subseteq J$, and 
  $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action  is
$\alpha'_k$,  $g_k$ is a guard associated to the vector.

\end{itemize}
%We define the set of holes and leaves of a pNet as follows:
%  \begin{itemize}
%\item
The set of \emph{holes} $\Holes(Q)$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets (we suppose those indexes 
disjoints). A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
is said to be \emph{open}.
%
%defined inductively; the sets of holes
%in a pNet node and its subnets are all disjoint:
%  \[\begin{array}{l}
%\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
%\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
%=J\uplus{\displaystyle \bigcup_{i\in 
%I}\Holes(\pNet_i)}\\
%\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
%\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
%\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
%\end{array}\]
%\item
The set of \emph{leaves} of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(Q)= \mylangle \pLTS_i \myrangle^{i \in L}$.
%\end{itemize}

\end{definition}

%\paragraph{Operational Semantics.}
%\label{section:op-semantics}
\renewcommand{\Pred}{g}
\renewcommand{\Post}{u}
The semantics of an open pNet is expressed as  an automaton where each transition coordinates the actions of several holes, 
the transition occurs if some predicates hold, and can involve  
 state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transition]
  \label{defn:OpenTransitions}
  An \emph{open transition} over a
  set of holes $J$  and a set of states $\mathcal{S}$ is 
       a structure of the form:
\\[-2ex]       
       \begin{mathpar}
       \inferrule*[myfraction=\reddottedrule]
       {       \beta_j^{j\in J}, \Pred, \Post}
       {s \OTarrow {{\alpha}}s'}
       \end{mathpar}
       Where $s, s'\in\mathcal{S}$ and $\beta_j\in\actions{\variables}$
        is an action of the hole $j$; $\alpha$ is the resulting global action; {\Pred} is a predicate 
       over the different variables of the
       terms, labels, and states  $\beta_j$, $s$, $\alpha$. $\Post\in 
       \assigns{\variables}$\ is 
       a set of assignments that are the effects of the transition.
Open transitions are identified modulo logical equivalence on their predicate.
%%   \noindent
%%   \begin{tabular}[b]{@{}p{0.86\columnwidth}c{}@{}}
%%     of holes $J$ and a set of states $\mathcal{S}$ is a structure
%%     of the form to the right,
%% %
%%     &
%% %
%%     %% \\[-2ex]	
%%     %% \begin{mathpar}
%%     \raisebox{1mm}[0pt]{
%%       $\inferrule*[myfraction=\reddottedrule]
%%       {	\beta_j^{j\in J}, \Pred, \Post}
%%       {s \OTarrow {{\alpha}}s'}$
%%       }
%%     %% \end{mathpar}
%% %
%%     \\
%% %
%%     \multicolumn{2}{p{\columnwidth}}{%
%% %
%%     where $s, s'\in\mathcal{S}$;
%%     $\beta_j\in\actions{\variables}$ is an action of the hole
%%     $j$; $\alpha$ is the resulting global action; {\Pred} is a
%%     predicate over the different variables of the terms, labels,
%%     and states $\beta_j$, $s$, $\alpha$; and $\Post\in
%%     \assigns{\variables}$ is a set of assignments that are the
%%     effects of the transition.  Open transitions are identified
%%     modulo logical equivalence on their predicate.
%% %
%%     }
%%   \end{tabular}
\end{definition}
%\begin{definition}[Open transitions]
%	\label{def:OpenTransitions}
%	An \emph{open transition} over a set \noteSB{In the architecture section, I say that 
%we omit indices on $\goesto{}$ and that they are always clear from the 
%context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
%	I}$} of LTSs, a
%	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
%	a structure of the form:	
%	\begin{mathpar}
%	\inferrule*[myfraction=\reddottedrule]
%	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
%		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
%	{s \OTarrow {v}s'}
%	\end{mathpar}
%	Where $s, s'\in\mathcal{S}$ and for all
%        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
%	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and 
%$\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort 
%$\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
%        is a transition of the hole $j$}, for any action $b_j$ in the
%        sort $\Sort_j$.  
%        of this open transition. \Pred\ is a predicate 
%	over the different variables of the
%	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
%	hold \emph{after the open transition}, they are represented as a substitution of the 
%	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
%	would like to just put $\Post_=e$}
%	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
%	variables of the open transition.
%\end{definition}

\newCoord{The red dotted rule expresses the implication stating that if the holes perform the designated actions and the condition $g$ is verified, then the variables are modified and the state changes. This implication however  uses a simple logic with the 
 expressive power  given by the predicate algebra (it must include logical 
 propositions and equality). 
 Proposition and inference rules of the paper use a standard logic, while predicates inside the open transitions should use a more restricted logic, typically a logic that could  be handled mechanically and expressed by terms that can be encoded in a simple syntax.
Open transitions express in a symbolic way, transitions  that are not only parameterised with variables but also actions of not yet known processes.}

\begin{definition}[Open automaton]
	\label{defn:open-automaton}
	An \emph{open automaton} is a tuple $(J,\mathcal{S},s_0,\mathcal{T})$ where:
%	\begin{itemize}
%		\item  
 $J$ is a  set of indices,
%		\item 
  $\mathcal{S}$ is a set of states and $s_0$ an initial state
		among $\mathcal{S}$,
%		\item 
$\mathcal{T}$ is a set of open transitions and for each
		$t\in \mathcal{T}$ there exist  $J'$ with  $J'
		\subseteq J$, such that $t$ is an open transition over  $J'$,
		and  $\mathcal{S}$.
%		
%	\end{itemize}
\end{definition}

%
%Then the semantics of a pNet is characterized by a set of {\em open
%transitions}, where the hypotheses on process parameters are
%replaced by 1) transitions of the pLTSs at the leaves, and 2) formal
%hypotheses on the transitions of the holes. A {\em predicate} is used
%to relate the parameters and names appearing in the actions of the
%leaves and the holes involved in the rules, but also appearing in  the resulting action.

The semantics of an open pNet is an open automaton where  the states are
tuples of states of the pLTSs at the leaves, denoted $\triangleleft\ldots\triangleright$.  Each open transition between two states contains 1) the actions of the holes involved in the transition, 2) a guard  built from the synchronisation vectors coordinating the holes and the transitions involved; 3) assignments and global state change  defined by the pLTSs transitions involved; 4) a global action  defined by the synchronisation vector.

\begin{example}[An open transition]
  \label{ex:TimerOT5}
  \newCoord{The open transition
  \begin{small}
    \[
    \openrule{
      \{E\mapsto ask\},
      t \in z,
      \{t:=t+1\}
    }{
      \ostate{11} \OTarrow{ask} \ostate{11}
    }
    \]
  \end{small}%
%% \begin{small}\[ \openrule
%%   {\{B\mapsto \mathit{fail}(b)\},
%%     t<Max \land b=true ,
%%     \{t := t+1\} }
%%   {\ostate{11} \OTarrow{\alpha(t)} \ostate{14}}%% \vspace{-1.3ex}
%%   \]
%% \end{small}
%
emits a global action $ask$ defined by the synchronisation vector $\SVTuple{\mathit{timeout}_T, \mathit{timeout}_C, -, \mathit{ask}} \rightarrow \mathit{ask}$.  It requires the hole at label $E$ to fire an $\mathit{ask}$ action, with the condition $t \in z$. In this case, the global pNet loops on the state $\ostate{11}$ that has internal variables $t$ and $z$ local to the pLTS $T$ (hence not appearing in the synchronisation vector).  The variable $t$ is updated to the new value $t+1$. \Fig{FailureTimer:pNet} shows the complete pNet, whereas \fig{ArchFailure:OA} shows a complete open automaton.}
%
%It corresponds to the firing of the \replaceSB{connector
%$''tick, fail, fail''$}{interaction $\listset{\NameTimer.\PortTick, \PortFail, \NameCtrl.\PortFail}$}. It states that, if the hole $B$ performs an action $fail$, and the (complex) guard is satisfied, then the variable $z$ is set to $z1$ (as specified by the guard), then the transition below the dotted line occurs and a global action labelled $\alpha_0$ occurs. Only the second pLTS performs a transition from state $1$ to state $2$.
%The guard corresponds both to the guard of this transition and to synchronisation vector constraints.
\end{example}
%
%\begin{definition}[States of open pNets]\label{def-states}
%  A state of an open pNet is a tuple (not necessarily finite) of the
%  states of its leaves.
%
%  For any pNet p, let $\Leaves(p) = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be 
%  the set of pLTS at its leaves,
%  then $States(p) = \{\triangleleft s_i^{i\in L}
%  \triangleright| \forall i\in L. s_i \in S_i\}$.
%A pLTS being its own single leave:
%  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S\}$.
%
%The initial state is defined as:
%$InitState(p) = \triangleleft {{s_i}_0}^{i\in L}  \triangleright$.
%\end{definition}
%
%
%
%%% \begin{example} \emph{State of a pNet}
%%%   The states of pNet \texttt{EnableCompL} are:
%%%   $\triangleleft 00 \triangleright, \triangleleft 10 \triangleright, \triangleleft 11 
%%%\triangleright$
%%% \end{example}
%
%\paragraph{Predicates:}
%Let
%$\mylangle\overline{\pNet},\overline{\Sort},\overline{\symb{SV}}\myrangle$
%be a pNet. Consider a synchronisation vector $\symb{SV}\in \overline{\symb{SV}}$. We 
%define a
%predicate $\Pred$ relating
%the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
%\begin{multline}
%\Predsv(\SV{{(\alpha'_i)}^{i\in I}, {(\beta'_j)}^{j\in J}} 
%{\alpha'} {g}, \alpha_i^{i\in I}, \beta_j^{j\in J}, \alpha)\Leftrightarrow \\
%\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, \beta_j=\beta'_j \land 
%\alpha=\alpha' 
%\land g
%\end{multline}
%
%
% 
%
%Somehow, this predicate entails a verification of satisfiability in the sense that if the 
%predicate $\Pred_{\symb{sv}}$ is not satisfiable, then the transition associated with the 
%synchronisation will not occur in the considered state. 
%
%
%In any other case (if the action families do not match or if there is no valuation of
%variables such that the above formula can be ensured) the predicate is undefined.
%
%This definition is not constructive but it is easy to build the predicate constructively
%by brute-force unification of the sub-pNets
%actions with the corresponding vector actions, possibly followed by a simplification
%step.
%
%
%
%
%\begin{definition}[Operational semantics of open pNets]
%	\label{defn:operationalSemantics}
%	The semantics of a pNet $p$ is an open automaton $A = 
%	<Holes(p),States(p),InitState(p),
%	\mathcal{T}>$ where $\mathcal{T}$ is the smallest set of open transitions		
%	satisfying the rules below:
%%	\begin{itemize}
%%		\item $J$ is the set of holes: $Holes(p)= J$. 
%		%  \item $\overline{L}^L = Leaves(p), \overline{H}^J = Holes(p)$
%%		\item ${\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
%%		\item $\mathcal{T}$ is the smallest set of open transitions		satisfying the 
%%rules below:
%%	\end{itemize}
%	
%	%% \TODO{ We should be careful here: after (re) reading "Huimin
%	%% 	Lin, 'Symbolic Transition Systems with Assignements', Concur'96" I
%	%% 	think handling assignments is not trivial, even for comparisons of pLTSs. }
%
%
%	
%	The rule for a pLTS  checks that the guard 
%	is verified and transforms assignments into post-conditions:
%	
%	\begin{description}
%		\item[{\bf Tr1:}]
%	$\inferrule
%		{ s \xrightarrow{\langle \alpha,~g,~u\rangle} s'\in \to  }
%		{ \mylangle  S,s_0, \to \myrangle
%			\models
%			\openrule
%			{\emptyset ,
%			g,u}
%			{\ostate{s} \OTarrow{\alpha} \ostate{s'}}
%		}
%		$
%	\end{description}
%%	Note that this note is greatly simplified by the fact that variables are local to 
%%	thread; introducing global state variables or accepting loops to the same 
%%	state would 
%%	require to reason 
%%	on the scope of 
%%	each variables, and to introduce additional variables to handle the several occurence 
%%	of the same pLTS variable in the predicates. Indeed the constraints on pLTS 
%%	transitions 
%%	ensure that the same variable never appears both on the left and on the right of the 
%%	equations of a predicate.
%	
%	The second rule deals with pNet nodes: for each possible
%	synchronisation vector (of index $k$) applicable to the rule subject, the premisses
%	include one {\em open transition} for each sub-pNet involved , one possible
%	{\em action} for each Hole involved, and the predicate relating these
%	with the resulting action of the vector. The sub-pNets involved are split between two 
%	sets, $I_2$ for subnets that are pLTSs, and $I_1$ for the others, $J$ is the set of 
%	holes involved in the transition\footnote{Formally, if $SV_k \!=\! \SV{\alpha_m^{m 
%	\in M}}{\alpha'}{g}$ is a synchronisation vector  of \pNet\  then $J=M\cap 
%	\Holes(\pNet)$, $I_2=M\cap \Leaves(\pNet)$,  $I_1=M\setminus J \setminus 
%	I_2$}.                                                                    
%	           
%	                                                       
%	\begin{description}
%		\item[{\bf Tr2:}]
%	\end{description}
%	
%%	\noindent
%%\begin{mathpar}
%%    \mprset {vskip=1ex}
%%\inferrule
%%    {
%%\Leaves(\mylangle \overline{\pNet}, \overline{\Sort}, \symb{SV}_k^{k\in 
%%    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
%%k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{g} 
%%\\
%%\\     	
%%	\forall m\!\!\in\!\! I_1. {\pNet_m 
%%	\models\openrule
%%    	{
%%    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
%%    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
%%    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
%%  \qquad
%%\forall m\!\!\in\!\! I_2.		{ \pNet_m 
%%    	 \models
%%    	\openrule
%%    	{\emptyset, \Pred_m, \Post_m}
%%    	{\ostate{s_m} \OTarrow {\alpha_m}
%%    		\ostate{s_m'}} }\\\\
%%     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
%%    	\Pred = \bigwedge_{m\in (I_1\uplus I_2)}\!\! \Pred_m \land
%%    	\Pred_{\symb{sv}}(SV_k,\alpha_m^{m\in (I_1\uplus I_2)},\beta_j^{j\in J},\alpha)\\ 
%%    		I' = \biguplus_{m\in I_1}\!\! I_m \uplus I_2
%%    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
%%    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
%%    }
%%    {\mylangle \overline{\pNet}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
%%    	\models
%%    	{\openrule
%%    		{
%%    		{\beta_j}^{j\in J^\prime}, \Pred, \uplus_{m\in I_1\uplus I_2} 
%%    		\Post_m}
%%    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
%%    			\ostate{(s_i^\prime)^{i\in L}}}
%%    	}
%%    }
%%\end{mathpar}  
%	\noindent
%\begin{mathpar}
%    \mprset {vskip=.8ex}
%\inferrule
%    {
%\Leaves(\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in 
%    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \qquad  	
%k\!\in\! K \qquad SV_k \!=\! \SV{(\alpha'_m)^{m \in I_1\uplus I_2\uplus J}}{\alpha'}{e_b} 
%\\
%\\     	
%	\forall m\!\!\in\!\! I_1. {\pNet_m 
%	\models\openrule
%    	{
%    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
%    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
%    		\ostate{(s_i^\prime)^{i\in L_m}}} }	
%  \qquad
%\forall m\!\!\in\!\! I_2.		{ \pNet_m 
%    	 \models
%    	\openrule
%    	{\emptyset, \Pred_m, \Post_m}
%    	{\ostate{s_m} \OTarrow {\alpha_m}
%    		\ostate{s_m'}} }\\\\
%     J' = \biguplus_{m\in I_1}\!\! J_m \uplus J 	\\
%    	\Pred = \bigwedge_{m\in I_1\uplus I_2}\!\! \Pred_m \land
%    	\Predsv(SV_k,\alpha_m^{m\in I_1\uplus I_2},\beta_j^{j\in J},\alpha)\\ 
%    	\forall i\in	L\backslash \left(\biguplus_{m\in I_1}\!\! L_m \uplus I_2\right).\,s'_i=s_i \\
%    \fresh(\alpha'_m,\alpha',\beta_j,\alpha) 
%    }
%    {\mylangle {\pNet}_m^{m\in I}, \set{\Sort}, \symb{SV}_k^{k\in K}\myrangle
%    	\models
%    	{\openrule
%    		{
%    		{\beta_j}^{j\in J^\prime}, \Pred,  \biguplus_{m\in I_1\uplus I_2} 
%    		\Post_m}
%    		{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
%    			\ostate{(s_i^\prime)^{i\in L}}}
%    	}
%    }
%\end{mathpar}

%\todoLHin{add a paragraph about pnets properties and usage, here are a few points (to be completed and written up properly depending on what we need}
%
%- pnets as a model for gcm components (behaviour, distributed systems, verification)
%
%- used for finite model checking (with finite domains for parameters)
%
%- smt framework for reasoning on infinite models
%
%- a theory of pnets: composition properties , stong (and weak) bisim
%	** extract a few useful composition theorems? **

%OLD	\begin{description}
%		\item[{\bf Tr2:}]\noteSB{I do not understand why the second line $pNet_m 
%\models\dots$ is not subsumed by the first one?}
%	\end{description}
%	
%	\noindent
%    $\inferrule
%    {k\!\in\! K \\SV_k \!=\! \alpha_m^{m \in I\uplus J} \!\to\! 
%    \alpha' |g \\
%    	Leaves(\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in 
%    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
%    	\forall m\!\!\in\!\! I. 	
%    {\left(\begin{array}{l}
%	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
%    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
%    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
%    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
%    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
%		{ \pNet_m 
%    	 \models
%    	\inferrule*[myfraction=\reddottedrule]
%    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
%    	{\ostate{s_m} \OTarrow {v_m}
%    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
%\end{array}\right)}\\
%    	%\land
%    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
%     J' = \biguplus_{m\in I}\!\! J'_m \uplus J 	\\
%    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
%    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
%    		I' = \biguplus_{m\in I}\!\! I_m'
%    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
%    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
%    }
%    {\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
%    	\models
%    	{\inferrule*[myfraction=\reddottedrule]
%    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
%    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
%    		\Post_m}
%    		{\ostate{s_i^{i\in L}} \OTarrow {v}
%    			\ostate{(s_i^\prime)^{i\in L}}}
%    	}
%    }
%    $
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}

We  used pNets to define a behavioural semantics for distributed components~\cite{AmeurBoulifa2017} that allows the verification of correctness properties  by model-checking. 
More recently, a bisimulation theory has been formalised for open pNets~\cite{HMZ-FORTE2016}.

%****************************************************************
%****************************************************************
%****************************************************************

\section{The theory of architectures with data}
\label{secn:archs}

This section presents the extension of the theory of
  architectures \cite{AttieBBJS16-architectures-faoc} with data and
  briefly discusses a special case of priority models, called
  \emph{maximal progress}.
%
  \newCoord{These extensions require us to define the framework in a
    manner that would allow formulating and proving the property
    preservation result (\thm{combining} below).  In
    \cite{AttieBBJS16-architectures-faoc}, this result is obtained by
    requiring, in the definition of architecture composition, that an
    interaction among coordinated components be only possible if both
    architectures ``agree'' that it should be enabled.  With respect
    to data, the main difficulty lies in ensuring that this
    ``agreement'' extends to the transferred data values.  A trivial
    extension would allow an interaction only if the data values
    proposed by both architectures coincide.  As this
    requirement is too restrictive, we go beyond by assuming the data
    domains to be ordered and taking the \emph{meet} of the proposed
    values.  The property preservation result then holds independently
    of the proposed values, provided that guards and update
    assignments are monotonic.

    An important insight is that, although the requirement that guards
    and update assignments be monotonic appears to be a limitation, it
    is, in fact, a generalisation of the usual setting.  Indeed, the
    usual settings, where data domains are not ordered, can be
    recovered here by considering trivial partial orders with no two
    distinct elements being comparable.  In such case, \emph{all
      expressions} are trivially monotonic.

    The intuition behind the proof of the preservation of safety
    properties in \cite{AttieBBJS16-architectures-faoc} is simple.
    The composition of two architectures combines the ``constraints''
    that they impose on the possible executions of the system: as
    stated above, an interaction is only enabled if both architectures
    ``agree''.  In \cite{Baranov17-phd}, it is shown that this
    intuition extends well to priorities in the \emph{offer semantics}
    of BIP.  However, this is not the case in the classical semantics.
    In this section, we informally discuss the special case of the
    maximal progress priority models, where property preservation does
    hold in the classical semantics of BIP.
%
}

%****************************************************************

\subsubsection{Components and composition}
\label{secn:components}
\begin{definition}[Component]
  \label{defn:component}
  A component is a tuple $\tuple{Q, q^0, V, \val{0}{}, P,
  \export, \goesto{}}$, where%% \vspace{-2ex}
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports};
    $\export : P \rightarrow 2^V$ is
    the \emph{set of variables exported by each port},
  \item $\goesto{}\, \subseteq
    Q \times (2^P\setminus \{\emptyset\}) \times \guards{V} \times \updates{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports,
    %% $\emptyset \neq a \subseteq P$,
    a monotonic Boolean \emph{guard}
    %% $g \in \guards{V}$ 
    and a monotonic \emph{update assignment}.
    %% \hl{$u \in \updates{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{Only exported variables, belonging to a
  $\export(p)$, appear in the component interface  (see
  \defn{im}).  We omit here this separation between internal and exported variables.
  }
%
  Notations $q \goesto{a, g, u} q'$ and $q \goesto{a, g, u}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

In this paper, we  use  a refined
version of the Failure Monitor architecture
from~\cite{CubETH-case-study} as a running example.  Although \fig{schema:ArchFailure:BIP}
shows the full definition of this architecture, we will explain its
various elements progressively.
\fig{schema:ArchFailure:BIP} shows  components
%% The
%% component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
%% it only has its interface defined: $(\emptyset, \{\PortFinish,
%% \PortResume, \PortFail\}, \export[\emptyset])$, with
%% $\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
%% \PortResume, \PortFail\}$.  
%% The other two components,
%
$\NameTimer$(imer) and $\NameCtrl$(ontrol), with interfaces
$\btuple{
  \{t, z\},
  \{\PortTick, \PortCancel, \PortTOT\},
  \bigl\{\!\PortTick \mapsto \{t, z\}\!\bigr\}
}$,
and
$\btuple{
  \{\mathit{zone}\},
  \{\PortReset, \PortFail, \PortResume, \PortTOC\},
  \bigl\{\!\PortFail \mapsto \{\mathit{zone}\}\!\bigr\}
}$
%\noteEM{Attention, bug dans la figure 2, ou tu mets $reset\mapsto \mathit{zone}$}\noteSB{Done!}
respectively.
%
%% \[
%% \begin{array}{@{}c@{\hspace{3mm}}c@{}}
%%   \bigl(
%%   \{t, z\},
%%   \{\PortTick, \PortCancel, \PortTOT\},
%%   \export[\NameTimer]
%%   \bigr)\,,
%%   %% \qquad
%%   &
%%   \bigl(
%%   \{\mathit{zone}\},
%%   \{\PortReset, \PortFail, \PortResume, \PortTOC\},
%%   \export[\NameCtrl]
%%   \bigr)\,,
%%   \\
%%   \export[\NameTimer](p) =
%%   \begin{cases}
%%     \{t, z\}, & \text{if } p = \PortTick,\\
%%     \emptyset, & \text{otherwise,}
%%   \end{cases}
%%   %% \]
%%   %% and
%%   %% \[
%%   %% \quad
%%   &
%%   \export[\NameCtrl](p) =
%%   \begin{cases}
%%     \{\mathit{zone}\}, & \text{if } p = \PortFail,\\
%%     \emptyset, & \text{otherwise.}
%%   \end{cases}
%% \end{array}
%% \]
%
%% , where $\export[\NameCtrl](\PortFail) =
%% \{\mathit{zone}\}$ and $\export[\NameCtrl](p) = \emptyset$, for all
%% other ports $p$; $\export[\NameTimer](\PortTick) = \{t, z\}$ and
%% $\export[\NameTimer](p) = \emptyset$, for all other ports $p$.
Variable $t$ is implicitly assumed to be \newCoord{of type Integer (with trivial ordering).  Variables $z \bydef{=} [z.l, z.u]$
and $\mathit{zone} \bydef{=} [\mathit{zone}.l, \mathit{zone}.u]$ are   of type Integer Interval 
ordered by interval inclusion}.

%% are \emph{coordinators}\mdash they are specified
%% completely, including the states and transitions defining their
%% \emph{behaviour}.
Component \emph{behaviour} is defined by states %% ($\{\cstate{0},
%% \cstate{1}, \cstate{2}\}$, for $\NameCtrl$; $\{\tstate{0}\}$, for
%% $\NameTimer$)
and transitions.
The initial states
\tstate{0} and \cstate{0}, and valuations $\val{0}{\NameTimer} = \{t
\mapsto 0, z \mapsto \topinterval\}$,
$\val{0}{\NameCtrl} =
\{\mathit{zone} \mapsto [\mathrm{Min}, \mathrm{Max}]\}$ are shown by
the incoming arrows $\goesto{\TimerInit[]} \tstate{0}$ and
$\goesto{\IntervalInit[]} \cstate{0}$ where $\topinterval=(-\infty, +\infty)$.
%\footnote{%
%%
%  We use $\topinterval$ to denote the top interval in the inclusion
%  order, \ie $(-\infty, +\infty)$.
%%
%} 
The constants $\mathrm{Min}$ and $\mathrm{Max}$ are the parameters
of the architecture.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{BIPspec-ArchFailureTimerMax-v4}
  \caption{The BIP specification of the Failure Monitor architecture}
  \label{fig:schema:ArchFailure:BIP}
\end{figure}

Transitions are labelled with ports of the
corresponding components, Boolean guards and update assignments on local
variables.  \Eg[,] the loop transition $\tstate{0}
\goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{0}$. % in the
%component $\NameTimer$ is labeled by the port $\PortTick$, the associated guarg and variable update. 
% It can be
%fired only when the current value of the local variable $t$ is less
%than $z.u$.
%%% , where $z = [l, u]$ is a local variable representing an
%%% interval
%Upon firing, this transition increments the value of $t$ by $1$.  
The
guards and update assignments of the transitions of $\NameCtrl$ are
omitted. By default, an omitted guard is  $\true$ and an omitted assignment is empty {\noop}.  
%\noteLH{It looks like Max is in zone, I do not understand}
%
Clearly, all guards and update assignments are monotonic.

%% \todoSBin{Speak of guard monotonicity}

\begin{definition}[Component semantics]
  \label{defn:comp:semantics}
  The \emph{open semantics} of a component $B = \tuple{Q, q^0, V,
  \val{0}{}, P, \export, \goesto{}}$ is %% given by
  the LTS denoted $\semopen{B} = \tuple{S,
  s^0, \goesto{}}$, where $S = Q \times \valuations{V}$, $s^0 =
  \tuple{q^0, \val{0}{}}$ and $\goesto{}$ is the minimal transition
  relation satisfying the rule
  %
  \begin{equation}
    \label{eq:comp:semantics}
    \infer{
      q \goesto{a, g, u} q'
      \and
      \val{}{} \models g
      \and
      \val[\primeit]{}{} = \tilde{\val{}{}}[u]
      \and
      \valdiff{\val{}{}}{\tilde{\val{}{}}} \subseteq \export(a)
    }{
      (q, \val{}{}) \goesto{a,\tilde{\val{}{}}} (q', \val[\primeit]{}{})
    }
    \,.
  \end{equation}
%
  \newCoord{%
    The \emph{closed semantics} of $B$ is given by the LTS denoted
  $\semclosed{B}$, comprising only those transitions of $\semopen{B}$,
  where $\tilde{\val{}{}} = \val{}{}$.}%% , which we denote
  %% $(q, \val{}{}) \goesto{a} (q', \val[\primeit]{}{})$.}
\end{definition}

The use of the intermediate valuation $\tilde{\val{}{}}$ in the
conclusion and the third premise of  \eq{comp:semantics}
allows some  variables to get new values before the
transition is  fired.  Thus the component is \emph{open}
to the exchange of data with its environment.  However, the
fourth premise  states that only the variables exported through the ports participating in the interaction can be affected by the data transfer.
%restricts the set of
%variables that can be affected by such data transfer to those
%that are exported through the ports participating in the
%interaction.
%
The closed semantics excludes this possibility of data exchange.

\begin{definition}[Interaction model]
  \label{defn:im}
  For a finite set of component interfaces $(V_i, P_i,
  \export[i])^{i \in I}$, such that all $P_i$ and all $V_i$ are
  pairwise disjoint, %% \ie $\forall i \neq j,\ P_i \cap P_j = V_i \cap
  %% V_j = \emptyset$. 
  let $P = \bigcup_{i \in I} P_i$, $V = \bigcup_{i
    \in I} V_i$ and $\export : P \rightarrow 2^V$ such that, for any
  $p \in P_i$, $\export(p) = \export[i](p)$. 
%
  %% \noteSB{Updated this}\hl{and $V_a \bydef{=} \bigcup_{i \in
  %%   \supp{a}} \export[i](a \cap P_i)$ the set of component
  %% variables exported through $a$.}

  An \emph{interaction model over $(V, P, \export)$} is a set $\Gamma
  \subseteq 2^P \times \guards{V} \times \updates{V}$, such that,
  for any interaction $(a, g, u) \in \Gamma$, we have
  $g \in \guards{\export(a)}$ and $u \in \updates{\export(a)}$.\footnote{%
%
    Notice that this definition allows $(\emptyset, \true,
    \noop)$ and $(\emptyset, \false, \noop)$ to be included in
    $\Gamma$.
%
  }
  %% We call the set of ports $P$ the \emph{domain} of the
  %% interaction model.
\end{definition}

%% \noteLH{supp should go outside the definition}
We  
assume that all sets of components and interfaces satisfy the disjointness assumption above.
%
We call the \emph{support} of a set of ports $a \subseteq P$, denoted
$\supp{a}$, the set of the participating components. It is either the set $\setdef{i \in I}{a \cap P_i \neq
  \emptyset}$ (for $P = \bigcup_{i=1}^n P_i$) or the set $\setdef{B
  \in \cB}{a \cap P_B \neq \emptyset}$ (for $P = \bigcup_{B \in \cB}
P_B$).  The precise meaning of this notation will always be clear from
the context.

\begin{definition}[Composition]
  \label{defn:im:sem}
  The \emph{composition} of a finite set of components $\cB = \tuple{Q_i,
  q^0_i, V_i, \val{0}{i}, P_i, \export[i], \goesto{}}^{i \in I}$ with
  the interaction model $\Gamma$ over $\tuple{V, P, \export}$
  is the component $\Gamma(\cB) = \tuple{Q, q^0, V, \val{0}{}, P,
    \export, \goesto{}}$, where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather*}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:sem}
%
    \infer{    
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, u_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      g' = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      u' = u; u_i^{i \in \supp{a}}
      \and
      (a, g, u) \in \Gamma
      \and
      a \neq \emptyset
    }{
      (q_i)^{i \in I} \goesto{a, g', u'} (q_i')^{i \in I}
    }\,.
  \end{gather*}
  %% \todoSBin{Add the intuitive explanation}
\end{definition}

  Intuitively, an interaction can be fired if all the involved
  components are ready to fire their corresponding transitions.  The
  other components do not change their states.  Both the interaction
  guard and those of the participating transitions must be satisfied.
  The update assignment of the interaction is executed first, followed
  by those of the components.

Specifying interaction models as sets of sets of ports is not
practical due to their potentially exponential size.  An algebra of
connectors was introduced in \cite{BliSif08-acp-tc} in order to
structure interactions in BIP models.
Connectors are hierarchical, tree-like structures with component ports
at the leaves.  They define sets of interactions, based on the
attributes of the nodes, which may be either \emph{trigger}
(triangles in \fig{schema:ArchFailure:BIP}) or \emph{synchron}
(bullets in \fig{schema:ArchFailure:BIP}).
%
If all sub-connectors of a connector are synchrons, then an
interaction is allowed by the connector only if each subconnector
can contribute.
%
If at least one of the sub-connectors is a trigger, then any
interaction consisting of contributions of any set of sub-connectors
\emph{involving at least one of the triggers} is allowed.
%
The interaction model is defined as the set of all interactions
allowed by at least one of the connectors.
%% The composition semantics of BIP systems consists in
%% firing exactly one interaction, enabled through at least one of the
%% top-level connectors, at each execution round.

For instance, the connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} of \fig{schema:ArchFailure:BIP} is a
two-level hierarchical connector.  In the  subconnector
{\PortFail \trigsynch \NameCtrl.\PortFail}, the port {\PortFail} is a
trigger, whereas {\NameCtrl.\PortFail} is a synchron.  This
subconnector allows two interactions: $\{\PortFail\}$ and $\listset{\NameCtrl.\PortFail, \PortFail}$.  Similarly, at the top level,
{\NameTimer.\PortTick} is a trigger, and the 
subconnector is a synchron. 
% Hence, the overall connector allows the
%interaction involving only one port {\NameTimer.\PortTick}\footnote{%
%%
%  Such interactions are called \emph{singleton interactions}.
%%
%} and any interaction involving both {\NameTimer.\PortTick} and
%\emph{some} interaction of the second-level subconnector.  
%Combining
%the two levels we can see that t
The entire connector defines the
following three interactions (observe that $\topinterval +
\NameTimer.t = \topinterval$ and $\NameTimer.z \cap \topinterval =
\NameTimer.z$):
%
%\begin{gather*}
$  \bigl(\{\NameTimer.\PortTick\},
  \true, \emptyset\bigr)
  %% \NameTimer.z := \NameTimer.z\bigr)
 $,
$  \bigl(\{\PortFail, \NameTimer.\PortTick\},
  \true, \emptyset\bigr)$
  %% \NameTimer.z := \NameTimer.z\bigr)
  ,
$
  \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \},
  \true,
  \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
$
%  .
%\end{gather*}
%% \begin{align*}
%%   \bigl(\{\NameTimer.\PortTick\}&,
%%   \true, 
%%   \NameTimer.z := \NameTimer.z\bigr)
%%   \,,\\
%%   \bigl(\{\PortFail, \NameTimer.\PortTick\}&,
%%   \true,
%%   \NameTimer.z := \NameTimer.z\bigr)
%%   \,,\\
%%   \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \}&,
%%   \true,
%%   \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
%%   \,.
%% \end{align*}

%% \todoSB{Update this}
%% \hl{For instance, in \fig{schema:ArchFailure:BIP}, the two ports
%% $\NameTimer.\PortStart$ and $\NameCtrl.\PortFail$ are always
%% synchronised, since they belong to exactly one binary sub-connector,
%% where they are both synchrons.  In particular, this means that
%% whenever the transition \cstate{0} \goesto{\PortFail} \cstate{1} is
%% fired, so is the transition \tstate{0} \goesto{\PortStart,
%%   \TimerInit[]} \tstate{1}, initialising the timer.  The binary
%% connector $\NameTimer.\PortStart \twosynch
%% \NameCtrl.\PortFail$ is a sub-connector of a hierarchical connector,
%% where the port $\NameOprnd.\PortFail$ is a trigger.  Thus, the above
%% interaction can only happen together with $\NameOprnd.\PortFail$,
%% forming a ternary interaction.  On the contrary, being a trigger, the
%% port $\NameOprnd.\PortFail$ can fire alone, forming a singleton
%% interaction.}

%% \todoSBin{Speak of data transfer}

%% \todoSBin{Provide the interaction model}

In addition to interaction models, BIP relies on 
  \emph{priority models} that impose  a strict partial order on
  interactions.  %% Formal definitions are provided in \app{arch:appendix}.
  Intuitively, an interaction can be fired only if all
  the higher-priority interactions available in the current state are
  disabled by their respective guards.
%
In the next sections, we will implicitly assume application of the  \emph{maximal
  progress} priority $\mu$, where $(a, g, u) \prec_\mu (b, h, w)$ iff  $a \subset b$ and $a \neq b$.  For instance, the
port $\NameTimer.\PortTick$ will never fire alone if the port $\PortFail$ is also enabled.
%% In \secn{conclusion}, we discuss the generalisation of the
%% obtained results to other priority models.

%% Below, we will assume that the maximal progress priority $\mu$ is
%% applied after each composition with an interaction model.

%% Finally, \emph{priorities}\mdash defined by a strict partial order on
%% the set of possible interactions\mdash narrow the choice among the
%% enabled interactions at any given round.  The default priority is the
%% so-called \emph{maximal progress}, whereby among any two interactions
%% $a \subset b$ (as sets of ports), $b$ has higher priority than $a$.

%% In our running example, under the maximal progress priority, the
%% port $\NameTimer.\PortTick$ will never fire alone in a global state,
%% where the \noteSB{This turns out to be the first use of the word. I am tempted to pretend I did not notice.}\hl{dangling} port $\PortFail$ is also enabled.

%% Adding priorities $\NameOprnd.\PortFinish < \NameCtrl.\PortReset$ and
%% $\NameOprnd.\PortFinish < \NameOprnd.\PortResume$, ensures that, after
%% a failure, the operation performed by $\NameOprnd$ cannot finish
%% unless it is explicitly resumed or the system is reset.

%% \begin{note}
%%   \label{rem:maxprogress}
%% \end{note}

%% %****************************************************************
%% \subsection{Running example: Failure Monitor architecture}
%% \label{secn:failure-monitor}

%\noteLH{Ludo ou simon: essayer de remonter la description generique en debut de section}

%****************************************************************
\subsubsection{Architectures}
\label{secn:archi}

Architectures are partial BIP models, with \emph{dangling}
ports that serve as placeholders for the eventual connection with
 \emph{operand} components.

%% Contrary to standard BIP models, architectures comprise one or several
%% \emph{operand} components, whereof only the set of \emph{ports} is
%% given.  Here, the operand component is $\NameOprnd$ and its interface
%% consists of the ports $\PortFinish$, $\PortResume$ and $\PortFail$.

%% Application of the Failure Monitor architecture ensures that, whenever
%% a failure is registered in the operand component, the system will be
%% reset, unless a resumption is registered within $\mathrm{Max}$ time
%% units (more details in Sect.~\ref{section:resultOA}).

%% \noteLHin{should we say that always epsilon C is the epsilon of C?}

\begin{definition}[Architecture]
  \label{defn:arch}
  An \emph{architecture} is a tuple $A = \tuple{\cC, V_A, P_A,
    \export[A], \Gamma}$, where%% \vspace{-1.5ex}
  \begin{itemize}
  \item $P_A$ and $V_A$ are sets of ports and variables, respectively;
\item
  $\cC$ is a finite set of components (called \emph{coordinators}),
  %% with pairwise disjoint sets of ports and variables,
  such that
  $\bigcup_{C \in \cC} P_C \subseteq P_A$ and
  $\bigcup_{C \in \cC} V_C \subseteq V_A$;
%
  ports in $P_A \setminus \bigcup_{C \in \cC} P_C$, which do not
  belong to any of the coordinators are called
  \emph{dangling};

\item
  $\export[A] : P_A \rightarrow 2^{V_A}$ is an export function, such
  that $\export[A](p) = \export[C](p)$, for any $C \in \cC$ and $p \in
  P_C$ and $\export[A](p) \subseteq V_A \setminus \bigcup_{C \in \cC}
  V_C$ for any dangling port $p$; and

\item
  $\Gamma \subseteq 2^{P_A} \times \guards{V_A} \times \updates{V_A}$
  is an interaction model over $(V_A, P_A, \export[A])$.
  \end{itemize}

\end{definition}

\begin{definition}[Application of an architecture]
  \label{defn:arch:application}
  Let $A = \tuple{\cC, V_A, P_A, \export[A], \Gamma}$ be an architecture and let $\cB$
  be a set of components, such that
  $V_A \subseteq V \bydef{=} \bigcup_{B \in \cB \cup \cC} V_B$,
  $P_A \subseteq P \bydef{=} \bigcup_{B \in \cB \cup \cC} P_B$
%
  %% \begin{align}
  %%   \bigcup_{B \in \cB} V_B \cap \bigcup_{C \in \cC} V_C = \emptyset\,,
  %%   &&
  %%   V_A \subseteq V \bydef{=} \bigcup_{B \in \cB \cup \cC} V_B\,,
  %%   \\
  %%   \bigcup_{B \in \cB} P_B \cap \bigcup_{C \in \cC} P_C = \emptyset\,,
  %%   &&
  %%   P_A \subseteq P \bydef{=} \bigcup_{B \in \cB \cup \cC} P_B\,,
  %% \end{align}
%
  and $\export[A](p) = V_A \cap \export[B](p)$, for any $B \in \cB$
  and $p \in P_A \cap P_B$.
%
  The \emph{application of the architecture $A$ to the set of
  components $\cB$} is the component $ A(\cB) \bydef{=}
  \mu\bigl((\IMextend{\Gamma}{P})(\cC \cup \cB)\bigr)$, where
%
  $\IMextend{\Gamma}{P} \bydef{=}
  \bsetdef{
    (a, g, u)
  }{
    a \subseteq P,
    (a \cap P_A, g, u) \in \Gamma
  }$
  is
  %% is the \emph{extension} of the interaction
  %% model $\Gamma$ to the set of ports $P$, \ie
  the interaction model
  over $\tuple{V, P, \export[A] \cup  \bigcup_{B \in \cB}\export[B]}$ %, with  $\export(p) = \export[C](p) \cup  \export[B](p)$
%\vspace{-1.5ex}
%  \[
%  \export(p) = 
%  \begin{cases}
%    \export[C](p), & \text{for all } C \in \cC, p \in P_C,
%    \\
%    \export[B](p), & \text{for all } B \in \cB, p \in P_B,
%  \end{cases}
%  \]
%% defined by \noteLH{looks like this also depends on Pa}
%% %
%%   \noteSB{It does, indeed.  To be completely rigorous, I could replace
%%     this with something like $\Gamma\uparrow_{P_A}^P$ but this is
%%     quite heavy and the dependency is already implicit from the fact
%%     that $\Gamma$ is defined on $P_A$}
%% %
%%   \noteSB{Which one of \eq{im:extension} and \eq{im:extension1} is
%%     more intuitive?}
%
  %% \begin{align}
  %%   \label{eq:im:extension}
  %%   \IMextend{\Gamma}{P} &\bydef{=}
  %%   \bsetdef{
  %%     (a \cup a', g, e)
  %%   }{
  %%     (a, g, e) \in \Gamma, a' \subseteq P \setminus P_A
  %%   }
  %%   \\
  %%   \label{eq:im:extension1}
  %%   &= 
  %%   \bsetdef{
  %%     (a, g, e)
  %%   }{
  %%     a \subseteq P,
  %%     (a \cap P_A, g, e) \in \Gamma
  %%   }
  %% \end{align}
%
  and $\mu(\dots)$ denotes the application of maximal progress.
\end{definition}

%% \todoSBin{Define equivalence $\arequiv$?}

An architecture $A$ enforces coordination constraints on the
components in $\cB$.  The interface $(V_A, P_A, \export[A])$ of an
architecture $A$ contains all ports of the coordinators
$\cC$ and the dangling ports, which must belong to
the components in $\cB$.  In the application $A(\cB)$, the ports
belonging to $P_A$ can only participate in the interactions
defined by the interaction model $\Gamma$ of $A$.  Ports which do
not belong to $P_A$ are not restricted and can participate in any
interaction.  %% In particular, they can join the interactions in
%% $\Gamma$ (see \defn{arch:application}).  If the interface of the
%% architecture covers all ports of the system, \ie $P = P_A$, we
%% have $P\setminus P_A = \emptyset$ and the only interactions
%% allowed in $A(\cB)$ are those belonging to $\Gamma$.  Notice also
%% that the restrictions imposed by \defn{arch:application} on the
%% set of operand components $\cB$ ensure that an interaction in
%% $\IMextend{\Gamma}{P}$ can only refer to variables of the
%% participating components.
%
%% Finally, 
The definition of $\IMextend{\Gamma}{P}$ requires that
an interaction from $\Gamma$ be involved in every interaction
belonging to $\IMextend{\Gamma}{P}$.  To allow the ports from $P
\setminus P_A$ to be fired independently in $A(\cB)$, one must
have $(\emptyset, \true, \noop) \in \Gamma$.

In our running example, there are four dangling ports.  Intuitively, the
architecture monitors the activation of the dangling port {\PortFail}, then
waits for a period comprised between $\mathrm{Min}$ and $\mathrm{Max}$
and, unless {\PortResume} is activated, asks for a system reset
through an invocation of the dangling port {\PortAsk}.

\begin{definition}[Composition of architectures]
  \label{defn:arch:composition}
  Let $A_i = \tuple{\cC_i, V_{A_i}, P_{A_i}, \export[A_i], \Gamma_i}$, for $i = 1,2$,
  be two architectures.  The \emph{composition} of $A_1$ and
  $A_2$ is the architecture $A_1 \arcomp A_2 = \tuple{\cC_1 \cup \cC_2,
  V_{A_1} \cup V_{A_2}, P_{A_1} \cup P_{A_2}, \export[A_1] \cup \export[A_2], \Gamma}$, where%% \vspace{-1ex}
%
  \begin{equation}
    \label{eq:arch:composition}
    \Gamma = \bsetdef{
      (a, g^1 \land g^2, u^1 \expmix u^2) 
    }{
      (a \cap P_{A_i}, g^i, u^i) \in \Gamma_i,
      \text{ for } i = 1,2
    }
    \,.
  \end{equation}
\end{definition}

%\begin{proposition}[Properties of $\arcomp$]
%  \label{prop:arcomp:nice}
  %% Architecture composition
$\arcomp$ is associative and commutative.
  %%; it is idempotent if all coordinators
  %% are deterministic; $A_{id} = \bigl(\emptyset, \emptyset,
  %% \emptyset, \emptyset, \{(\emptyset, \true, \noop)\}\bigr)$ is its neutral element, \ie for
  %% any architecture $A$, we have $A \oplus A_{id} \arequiv A$.
  %% Furthermore, for any component $B$, we have $A_{id}(B) = \mu(B)$.
%\end{proposition}
%
%% \begin{proof}[Sketch of the proof]
%%   %% Commutativity and associativity 
%%   Follows from the corresponding
%%   properties of set union, Boolean conjunction and
%%   semilattice meet.
%% %
%% %% \todoLH{Check when equivalence is defined}
%% %%   Suppose we have two architectures $A \arequiv A'$.  This does not
%% %%   necessarily mean that their sets of coordinators
%% %%   coincide.  However, if all the involved coordinators
%% %%   are deterministic, then, in any state of $(A \arcomp A')(\cB)$,
%% %%   both architectures will impose the same restrictions, enabling
%% %%   the same interactions between the coordinators and operand
%% %%   components.  Hence, we have $(A \arcomp A')(\cB) = A(\cB) =
%% %%   A'(\cB)$.  Since this holds for any set of components $\cB$, we
%% %%   conclude that $A \arcomp A' \arequiv A \arequiv A'$.
%% %% %
%% %%   The properties of $A_{id}$ follow immediately from the
%% %%   definitions of architecture application and composition.
%% \end{proof}
%% The last statement of \prop{arcomp:nice} highlights a subtle point:
%% $A_{id}$ is the neutral element of the $\arcomp$ operator\mdash it is
%% not the identity operator on components.
%****************************************************************
%\subsubsection{Preservation of safety properties}
%\label{secn:safety}

\medskip
\newCoord{It is well known that, since violations of safety properties
  are characterised by finite executions, they can also be represented
  as state predicates: intuitively, a safety property corresponds to
  the predicate characterising the set of states, where this property
  is not violated.}

For a component $B$, we denote $S_{\semclosed{B}}$ and
$s^0_{\semclosed{B}}$ the corresponding constituents of
$\semclosed{B}$ (see Definition~\ref{defn:comp:semantics}).

\begin{definition}[Properties]
  \label{defn:property}
  Let $B$ be a component.  A \emph{(safety) property} 
  of $B$ is a predicate $\Phi$ on
  $S_{\semclosed{B}}$, such that $\bigl((q,
  \val{}{}) \models \Phi\bigr) \land (\val[\primeit]{}{} \order
  \val{}{})$ implies $(q, \val[\primeit]{}{}) \models \Phi$.
  %% ,
  %% where we write $(q, \val{}{}) \models \Phi$ iff $\Phi(q,
  %% \val{}{}) = \true$.
  A property $\Phi$ is \emph{initial} if
  $s^0_{\semclosed{B}} \models \Phi$.
  %% \todoSB{Do we need this?}\hl{;
  %% it is \emph{reachable} iff there exists a possibly empty path
  %% $s^0_{\semclosed{B}} \goesto{a^1, \val{1}{}} s^1 \goesto{a^2,
  %%   \val{2}{}} \cdots \goesto{a^n, \val{n}{}} s^n$, such that
  %% $s^n \models \Phi$}.
\end{definition}

\label{property:reset}
Although we define properties as state predicates, any
appropriate logic can be used to specify them.  For instance, the
property \emph{``There is always a possibility to reset the system
  after a single failure''} (\ie without
  additional failures having to occur in the meantime) enforced by the
  Failure Monitor architecture \newCoord{comprises the safety component that} can be specified using CTL as
  %  \noteSB{Removed $\lnot \PortResume$, because not needed \emph{and} could not reformulate nicely in English.}
    $\AG
\bigl(\PortFail \rightarrow \EX \EW[]{\lnot \PortFail}{
  \PortReset}\bigr)$.  %% Additional properties are provided in
%% \app{properties}.
%% \todoSB{Cite the original paper somewhere.}
An architecture enforces
its characteristic property on  its operand components.
From this point of view, the set of coordinators is
not relevant, neither are their states.  Thus, 
properties enforced by architectures only involve
the unrestricted composition of the operands:

\begin{definition}[Enforcing properties]
  \label{defn:enforce}
  Let $A = \tuple{\cC, P_A, V_A, \export[A], \Gamma}$ be an architecture; let $\cB$
  be a set of components and $\Phi$ an initial property of
  their parallel composition $\Gamma_{\|}(\cB)$,
  with $\Gamma_{\|} = \setdef{(a,\true,\emptyset)}{a \subseteq \bigcup_{B \in \cB} P_B}$.
  We say that \emph{$A$ enforces $\Phi$ on
    $\cB$} iff, for every state $s = (s_c, s_b)$ reachable in
  $\semclosed{A(\cB)}$, with 
  $s_c \in \prod_{C \in \cC} S_{\semclosed{C}}$ and
  $s_b \in \prod_{B \in \cB} S_{\semclosed{B}}$,
  we have $s_b \models \Phi$.
\end{definition}
In the following, when we say that an
architecture enforces some property $\Phi$, $\Phi$ is supposed to be initial for the coordinated components.
%
%According to the above definition, when we say that an
%architecture enforces some property $\Phi$, it is implicitly
%assumed that $\Phi$ is initial for the coordinated components.
%Below, we omit mentioning this explicitly.
In \cite{arch-pNets-RR}, we formally define
 \emph{upwards compatibility} that ensures property preservation when composing architectures. Informally, two architectures $A_1$ and $A_2$
are  upwards compatible iff, whenever their composition
involves the fusion of two interactions $a_1 = a \cap P_{A_1}$ and
$a_2 = a \cap P_{A_2}$ (see \eq{arch:composition}) and one, say
$a_1$, is inhibited in a given state by a larger interaction $b_1
\supset a_1$, there exists an interaction $b_2 \supseteq a_2$ that can
be fused with $b_1$ to form an interaction enabled in the same state.

\begin{theorem}[Preserving enforced properties]
  \label{thm:combining}
  Let $\cB$ be a set of components; let $A_i = \tuple{\cC_i, V_{A_i},
  P_{A_i}, \export[A_i], \Gamma_i}$, for $i = 1,2$, be two upwards compatible architectures
  enforcing on $\cB$ the properties $\Phi_1$ and $\Phi_2$
  respectively.  The composition $A_1 \arcomp A_2$ enforces on
  $\cB$ the property $\Phi_1 \land \Phi_2$.
\end{theorem}

%\addSB{%
%
\Thm{combining} implies that safe BIP systems can be designed
\emph{compositionally}:  it is sufficient to verify that
1)~the applied architectures do enforce their characteristic
properties and 2)~they are pairwise upwards compatible.  To
a large extent, %% item~2, \ie verification of the upwards
%% compatibility,
the latter
can be carried out syntactically by analysing the
structure of the connectors that define the interaction models.
%% maximal progress is mostly relevant for connectors involving
%% triggers; whenever a trigger port is not common to both
%% architectures, upwards \emph{in}compatibility cannot arise.
The
next section is devoted to the encoding of architectures into
pNets, addressing item~1 by symbolic verification.
%
%}

%****************************************************************
%****************************************************************

\section{Encoding of architectures into open pNets}
\label{secn:encoding}

We  define the encoding of BIP architectures into pNets by
associating to each architecture $A = \tuple{\cC, V_A, P_A,
  \export[A], \Gamma}$ with $C = \tuple{Q_C, q^0_C, V_C, \val{0}{C},
  P_C, \export[C], \goesto{}}$, for each $C \in \cC$, \emph{and a
  partition $\partition \subseteq 2^{P_A}$ of its dangling ports} (\ie
$\biguplus_{D \in \partition} D = P_A \setminus \bigcup_{C \in \cC}
P_C$), the corresponding pNet $\nopri[A, \partition]$.
%
%% and $D_1 \cap D_2 = \emptyset$, for any two distinct $D_1, D_2
%% \in \partition$.
%
%% Notice that this addition 
%% is a matter of convenience, since one can
%% use the trivial partition $\partition \bydef{=} \{P_A \setminus \bigcup_{C
%%   \in \cC} P_C\}$.
For the sake of clarity, we define the encoding  without  any priority model.  Then, we provide a brief sketch of the
modifications necessary to encode maximal
progress (implicitly assumed).
%% %****************************************************************
%% \subsection{Encoding without the priority model}
%% \label{secn:enc:nopri}
Recall that $\Gamma$ is an interaction model over the interface
$\tuple{V_A, P_A, \export[A]}$, \ie these interface elements are
implicitly involved in the definition of $\Gamma$.
%
We define $\nopri[A, \partition] \bydef{=} \pNetTuple{(\nopri[C])^{C
    \in \cC}, \partition, \nopri[\Gamma]}$, where $\nopri[C]$ and
$\nopri[\Gamma]$ are the encodings of a coordinator $C$ and the
interaction model $\Gamma$ respectively.

\newCoord{Below, we present both the encodings of coordinators and
  interaction models.  The key constraint is that we encode each
  connector by one synchronisation vector.  This is necessary to
  1)~preserve the structure of the system and 2)~allow the encoding of
  maximal progress.

  Although somewhat technical, the encoding of coordinators is, in
  fact, pretty straightforward, comprising three key ideas: 1)~we
  introduce an additional transition (hence also an additional state)
  to explicitly initialise the variables; 2)~we introduce additional
  input variables to manipulate the values provided to the coordinator
  by the rest of the system for all exported variables; and
  3)~following the classical technique~\cite{milner83-calculi}, we
  simulate the absence of action by an additional loop transition.

  The encoding of connectors (interaction models) is more involved.
  Since a connector represents a set of potential interactions, some
  ports may not participate in all of them.  To encode this
  possibility, we introduce, for each port, an additional Boolean
  variable denoting whether the port participates in the interaction
  or not and, for each connector, a predicate characterising the
  interaction pattern.  Intuitively, the semantics of \emph{flat} BIP
  connectors~\cite{BliSif08-acp-tc} depends on the synchron/trigger
  annotations of ports.  If all ports in a connector are synchrons,
  the only allowed interaction is that comprising all the ports, \ie
  they all have to ``agree to interact''.  This case corresponds
  precisely to the semantics of synchronisation vectors in pNets.  If
  a connector has at least one trigger, then the allowed interactions
  are those that comprise at least one trigger, \ie they must be
  ``initiated by a trigger''.  In a hierarchical connector, these
  principles are applied recursively.  Thus, we observe a
  ``causality'' relation among ports of a connector: participation of
  a synchron in an interaction \emph{implies} that of a trigger.
  Causal Interaction Trees and Systems of Causal Rules, proposed in
  \cite{BliSif10-causal-fmsd}, formalise this causality relation and
  provide transformations from connectors to Boolean predicates and
  back.  Since we use SMT techniques for the analysis of the resulting
  pNet, the encoding presented below is optimised to reduce the number
  of variables by treating separately the ``top-level'' triggers (\eg
  $\NameTimer.\PortTick$ in the connector {\NameTimer.\PortTick
    \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} in
  \fig{schema:ArchFailure:BIP}).
%
}

\paragraph{Encoding the coordinators.}
The encoding of a coordinator $C$ is a pLTS with: 1)~an
 initial state and an   
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~an action algebra that matches the actions of the coordinator ports but adds, an additional Boolean action parameter, and also, for each exported
variable $x$, a corresponding fresh input variable $?x'$ to allow updates during interactions, 3)~pLTS transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 4)~additional loop transitions marked by
$\false$.
%% \noteLH{enleve le 3 et integre au 2}
Formally,  $\nopri[C] \bydef{=} \pNetTuple{S, s_0, \goesto[{\nopri[C]}]{}}$, such that
%
\begin{itemize}
\item $s_0 \not\in Q_C$ is fresh and $S = Q_C \cup \{s_0\}$,
  %% \noteEM{En principe les $vars(s)$ devraient etre disjoints, mais je
  %%   veux bien laisser ca sous le tapis tant qu'on n'implemente pas le codage...}
\item $\vars(s) = V_C \cup \setdef{?x'}{x \in \export[C](p), p \in a, s\goesto{a} }$, for all $s \in Q_C$, and $\vars(s_0) =
  \emptyset$,
%% \item the transition relation $\goesto[{\nopri[C]}]{}$ is defined as
%%   the union of three sets of transitions
%%   (with $\mathit{init} \bydef{=} \bigl(x := \val{0}{C}(x)\bigr)^{x \in V_C}$):
%% %
%%   \begin{multline}
%%     \label{eq:nopri:trans}
%%     {\goesto[{\nopri[C]}]{}} \bydef{=}
%%     \bsetdef{\tuple{s, a(\true), g, u, s'}}{
%%       \tuple{s, a, g, u, s'} \in \goesto{}}
%%     \\    
%%     \cup
%%     \bsetdef{\tuple{s, a(\false), \true, \noop, s}}{
%%       s \in Q_C, a \in P_C}
%%     %% \\
%%     \cup
%%     \bigl\{(s_0, \init, \true, \mathit{init}, q^0_C)\bigr\}
%%     \,.
%%   \end{multline}

\item let  $\mathit{uinit} \bydef{=} \bigl(x := \val{0}{C}(x)\bigr)^{x \in V_C}$ and,
  %% \noteSB{Ludo, Eric, pourriez-vous v\'erifier que cela est
  %%   compr\'ehensible ? Ok pour moi (Eric) OK Ludo, rephrase un peu}
  %
  for all $s \goesto{a, g, u} s'$ with $u = (x\! :=
\! e_x)^{x
    \in V}$ (with $V \!\subseteq\! V_C$),
%  
  let $u' \bydef{=}
  \bigl(x := e_x\bigr)^{x \in V \setminus \export[C](a)}
  \cup
  \bigl(x := e_x\bigl[\nicefrac{?x'}{x}\bigr]\bigr)^{x \in V \cap \export[C](a)}$,
%
  and $\export[C]'(a) \bydef{=} \setdef{?x'}{x \in \export[C](p), p \in a}$,
%
%% \noteEM{j'ai inverse les $\export[C]'(a), \export[C](a)$ dans les
%%   actions, pour etre conforme a la notation de la Def 1 page 4. }
\begin{multline*}
    {\goesto[{\nopri[C]}]{}} \bydef{=} 
    \Setdef{\Tuple{s, a\bigl(\export[C]'(a), \export[C](a),\true\bigr), g, u', s'}}{
      s \goesto{a, g, u} s'}
    \\
    {} \cup
    \Setdef{\Tuple{s, a\bigl(\export[C]'(a), \export[C](a),\false\bigr), \true, \noop, s}}{
      s \in Q_C, \exists s' \in Q_C: s' \goesto{a}}
    \\
    {} \cup \bigl\{(s_0, \init, \true, \mathit{uinit}, q^0_C)\bigr\}.
  \end{multline*}
\end{itemize}%\vspace{-3ex}

The loop transitions marked by $\false$ will be used in the encoding
of connectors.  Each BIP connector can define several interactions,
\ie ports involved in a connector need not necessarily always
participate.  On the contrary, each action in a pNet synchronisation
vector must participate in the synchronisation.  To address this
difference, we use the classical approach
where non-participation of a port in an interaction is simulated
by an additional loop transition~\cite{milner83-calculi}.

\Fig{FailureTimer:pNet} shows the encoding of the Failure Monitor
architecture, including the encodings of the two coordinators, \ie
$\nopri[\NameTimer]$ and $\nopri[\NameCtrl]$.  Notice that the
encoding in the figure is slightly optimised: some of the ports do not
have an associated Boolean value, nor the additional loop transitions.
We will explain this optimisation after we define the
encoding of the interaction model.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{FailureTimerPNET-v4}
  \caption{The open pNet encoding the Failure Monitor architecture (\fig{schema:ArchFailure:BIP}) without the Max Progress priority model}
  \label{fig:FailureTimer:pNet}
%% %
%%   \noteSBin{The updated encoding exports more primed variables in the
%%     actions, \eg $tick(z,t,z',t')$ and $fail(zone, zone', b1)$ instead
%%     of $tick(z,t,z')$ and $fail(zone, b1)$, and SVs, \eg $\dots \land
%%     (t' = t) \land (zone' = zone)$ in the guard of SV0. Shall we
%%     update the figure to include these or, rather, explain that the
%%     figure is cleaned-out for readability?}
%
\end{figure}

%% \begin{figure*}[t]
%%   \fbox{%
%%   \includegraphics[width=\textwidth]{BIPspec-ArchFailureTimerMax-v4-no-pri}}
%%   \caption{Encoding of the Failure Monitor architecture
%%     (\fig{schema:ArchFailure:BIP}) as an open pNet (without the
%%     priority model) \hl{\bf (Error in the picture: First two positions in the synchronisation vectors must be transposed)}}
%%   \label{fig:enc:nopri}
%% \end{figure*}

\paragraph{Encoding the interaction model}
%% First of all, recall that $\nopri[A, \partition ] \bydef{=}
%% \pNetTuple{(\nopri[C])^{C \in \cC}, \partition, \nopri[\Gamma]}$.
%% Thus, 
The holes in $\nopri[A, \partition]$ are indexed by the elements of
the partition \partition.  For the encoding of our running example, we
take $\partition = \bigl\{\{\PortFail, \PortResume\},\{\PortAsk,
\PortReset\}\bigr\}$.  This corresponds to the intuition that the
dangling ports {\PortFail} and {\PortResume} will be provided by a
%\noteLH{monitored ou monitoring???}
monitored component, whereas {\PortAsk} and {\PortReset} correspond to
the actions provided by the ``environment'' (other components of the
system) that are invoked in case of a persistent failure.  As for
the encoding of the coordinators, in the synchronisation vectors of
\nopri[\Gamma], we will associate Boolean values to the actions
corresponding to these ports.

The encoding of the interaction model is based on its representation
as a set of connectors. %% \noteSB{%
%
%%   A legitimate question is whether we actually need to do this,
%%   instead of listing explicitly the individual interactions, then
%%   associating a synchronisation vector to each one of them?  One of
%%   the factors is the encoding pNets$\rightarrow$SAT: is it better to
%%   have more or fewer SVs if the price is additional Boolean variables?
%%   The other factor might be the facility of maximal progress
%%   encoding\mdash TBC.  Traceability.  Anything else?\newline
%%   \hl{UPDATE: Maximal progress encoding requires this!}
%% %
%% }
Indeed, as illustrated by the Failure Monitor architecture in
\fig{schema:ArchFailure:BIP}, each connector can define several
allowed interactions, depending on its hierarchical structure and the
use of synchrons and triggers.

%\addSB{
We encode all interactions of a connector in one synchronisation vector.  This will allow us to also encode the maximal progress priority model.
%}
We use
the additional Boolean values associated to each port by the encoding
of coordinator components.  For example, observe that the three ports in 
the connector
{\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)}
form a ``causality chain'': {\NameCtrl.\PortFail} can only
participate in an interaction if the dangling port {\PortFail}
participates, which in turn can only happen if
{\NameTimer.\PortTick} does so.  These dependencies can be rewritten
as Boolean implications $\NameCtrl.\PortFail \Rightarrow \PortFail$
and $\PortFail \Rightarrow \NameTimer.\PortTick$.
%
The conjunction of these two implications can be used as a guard for
the synchronisation vector encoding this connector.

Within the scope of this connector, the port {\NameTimer.\PortTick}
participates in all interactions.
%
Furthermore, it is not involved in any other
connector.  Hence, the loop transition in \nopri[\NameTimer]
labelled by $\PortTick(\false)$ can never be taken and, therefore, can
be removed from the encoding.  Since only the transition
labelled by $\PortTick(\true)$ is ever taken, 
%
the
implication $\PortFail \Rightarrow \NameTimer.\PortTick$ is a tautology and can also be discarded.

We obtain the
synchronisation vector SV0 shown in \fig{FailureTimer:pNet}, where $b0$ and
$b1$ are the Boolean values associated to the actions encoding the
ports {\PortFail} and {\NameCtrl.\PortFail}.
The guard $b1 \Rightarrow b0$ encodes the causal
relation between these ports.  Notice
that %% , in our encoding,
all three ports are present in the
synchronisation vector.  \Fig{FailureTimer:pNet} shows
the four synchronisation vectors SV0\ndash SV3 corresponding to the
connectors in \fig{schema:ArchFailure:BIP} and an additional vector
SV4, synchronising the {\init} transitions of the two pLTSs.

In the general case, the encoding relies on the causal semantics of the
algebra of BIP connectors \cite{BliSif10-causal-fmsd}.  Disregarding
the variables and data transfer, the Algebra of Connectors $\ac(P)$
\cite{BliSif07-acp-emsoft} provides a syntactic notation for the BIP
connectors.  The causal semantics of the connectors, given in terms of
the Algebra of Causal Interaction Trees $\ct(P)$, elicits the
causal dependencies through an encoding mapping
$\tau: \ac(P) \rightarrow \ct(P)$.  Another mapping $R: \ct(P)
\rightarrow \cru(P)$ encodes causal interaction trees into systems of
causal rules, which are Boolean implications similar to the ones in
the example above.  %% We briefly summarise the key definitions and
%% results in \app{algebras}.
The $\ac(P)$, $\ct(P)$ and $\cru(P)$
representations of the connectors in \fig{schema:ArchFailure:BIP} are
shown in \tab{connectors} (elements shown in red can be removed for
simplification as described in the example above).

\begin{table}[t]
  \caption{Algebraic representations of the connectors in \fig{schema:ArchFailure:BIP}}
  \label{tab:connectors}
  \resizebox{\columnwidth}{!}{%
  \begin{tabular}{@{}c|c|c@{}}
    \hline {\bf Connector} & {\bf Causal Interaction Tree} & {\bf
      System of Causal Rules}
    \\\hline    
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']

      \node(caa) {}; % node distance=5cm [right of=start]
      \node[node distance=1cm] (cb) [below of=caa]{\PortFail};
      \node[node distance=1.8cm] (cc) [right of=cb]{\NameCtrl.\PortFail};
      \node[node distance=1.6cm] (ca) [left of=cb]{\NameTimer.\PortTick};

      \draw [style=-*, thick, black] ($(caa.south)+(0,.1cm)$) -- ++(right:.5cm) -| (cc.north);
      \draw [style=-triangle 45 reversed, thick, black]  ($(caa.south)+(0,.1cm)$) -| (cb.north);
      \draw [style=-*, thick, black] ($(caa.south)+(0,.5cm)$) -- ++(right:.2cm) -| ($(caa.south)+(+.9,0cm)$);
      \draw [style=-triangle 45 reversed, thick, black]  ($(caa.south)+(0,.5cm)$) -| (ca.north);
    \end{tikzpicture}
%
    &
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth'
        ,initial text=
        ,every state/.style={draw=black,thick}
        ,group/.style = {draw=black,thin,rectangle, minimum width=2.5cm}
        ,port/.style = {font=\small, minimum size=5mm}
        ,legend/.style = {font=\bf}
      ]

      \node(tick){\color{red}\NameTimer.\PortTick};
      \node[node distance=1cm](x) [right of=tick]{};
      \node(fail) [below of=x] {\PortFail};
      \node[node distance=1cm](y) [right of=fail]{};      
      \node(ctrlfail) [below of=y] {\NameCtrl.\PortFail};

      \draw[style=->] (tick) -> (fail);
      \draw[style=->] (fail) -> (ctrlfail);
    \end{tikzpicture}
%
    &
%
    $
    \begin{aligned}[b]
      \NameCtrl.\PortFail &\Rightarrow
      \PortFail {\color{red}{}\land \NameTimer.\PortTick}
      \\
      {\color{red}\PortFail} &{\color{red}{}\Rightarrow \NameTimer.\PortTick}
      \\
      {\color{red}\NameTimer.\PortTick} &{\color{red}{}\Rightarrow \true}
      \\
      {\color{red}\true} &{\color{red}{}\Rightarrow \NameTimer.\PortTick}

    \end{aligned}
    $
%
    \\\hline
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']

      \node[](baa) {}; % node distance=1.8cm [right of=start]
      \node[node distance=1cm] (bb) [below of=baa]{\NameTimer.\PortCancel};
      \node[node distance=1.8cm] (bc) [right of=bb]{\NameCtrl.\PortResume};
      \node[node distance=1.6cm] (ba) [left of=bb]{\PortResume};

      \draw [style=-*, thick, black] ($(baa.south)+(0,.1cm)$) -- ++(right:.5cm) -| (bc.north);
      \draw [style=-*, thick, black]  ($(baa.south)+(0,.1cm)$) -| (bb.north);
      \draw [style=-*, thick, black] ($(baa.south)+(0,.5cm)$) -- ++(right:.2cm) -| ($(baa.south)+(+.9,0cm)$);
      \draw [style=-triangle 45 reversed, thick, black]  ($(baa.south)+(0,.5cm)$) -| (ba.north);
    \end{tikzpicture}
%
    &
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth'
        ,initial text=
        ,every state/.style={draw=black,thick}
        ,group/.style = {draw=black,thin,rectangle, minimum width=2.5cm}
        ,port/.style = {font=\small, minimum size=5mm}
        ,legend/.style = {font=\bf}
      ]

      \node(res){\color{red}\PortResume};
      \node[node distance=1cm](rest) [below of=res] {\NameTimer.\PortCancel\ \NameCtrl.\PortResume};
      \draw[style=->] (res) -> (rest);
    \end{tikzpicture}
%
    &
%
    $
    \begin{aligned}[b]
      \NameCtrl.\PortResume &\Rightarrow
      {\color{red}\PortResume \land{}} \NameTimer.\PortCancel
      \\
      \NameTimer.\PortCancel &\Rightarrow 
      {\color{red}\PortResume \land{}} \NameCtrl.\PortResume
      \\
      {\color{red}\PortResume} &{\color{red}{}\Rightarrow \true}
      \\
      {\color{red}\true} &{\color{red}{}\Rightarrow \PortResume}
    \end{aligned}
    $
%
    \\\hline
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']

      \node (start) {};
      \node(h4)[left of=start]{}; %, node distance=1.5cm

      \node[node distance=1cm] (ab) [below of=start] {\NameTimer.\PortTOT};
      \node[node distance=2.1cm] (ac) [right of=ab]{\NameCtrl.\PortTOC};
      \node[node distance=1.5cm] (aa) [left of=ab]{\PortAsk};

      \draw [style=-*, thick, black] ($(h4.south)+(0,.1cm)$) -- ++(right:.5cm) -| (ac.north);
      \draw [style=-*, thick, black]  ($(h4.south)+(0,.1cm)$) -| (ab.north);
      \draw [style=-*, thick, black]  ($(h4.south)+(0,.1cm)$) -| (aa.north);
    \end{tikzpicture}
%
    &
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']
      \node (start) {\color{red}\PortAsk\ \NameTimer.\PortTOT\ \NameCtrl.\PortTOC};
      \node()[below of=start]{};
    \end{tikzpicture}
%
    &
%
    $
    \begin{array}[b]{c@{}}
      {\color{red}
        \begin{aligned}
          \NameCtrl.\PortTOC &\Rightarrow
          \PortAsk \land \NameTimer.\PortTOT
          \\
          \NameTimer.\PortTOT &\Rightarrow 
          \PortAsk \land \NameCtrl.\PortTOC
        \end{aligned}}
      \\[11pt]
      {\color{red}
        \begin{aligned}
          \PortAsk &\Rightarrow
          \NameTimer.\PortTOT \land \NameCtrl.\PortTOC
          \\
          \true &\Rightarrow
          \PortAsk \land \NameTimer.\PortTOT
          \land \NameCtrl.\PortTOC
        \end{aligned}

      }
    \end{array}
    $
    %
    \\\hline
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']

      \node (start) {};
      \node(h4)[left of=start]{}; %, node distance=1.5cm

      \node[node distance=1cm] (ab) [below of=start] {\NameCtrl.\PortReset};
      \node[node distance=1.5cm] (aa) [left of=ab]{\PortReset};

      \draw [style=-*, thick, black]  ($(h4.south)+(0,.1cm)$) -| (ab.north);
      \draw [style=-*, thick, black]  ($(h4.south)+(0,.1cm)$) -| (aa.north);
    \end{tikzpicture}
%
    &
%
    \begin{tikzpicture}[shorten >=1pt,node distance=.7cm,>=stealth']
      \node (start) {\color{red}\PortReset\ \NameCtrl.\PortReset};
      \node()[below of=start]{};
    \end{tikzpicture}
%
    &
%
    {\color{red}$
    \begin{aligned}[b]
      \NameCtrl.\PortReset &\Rightarrow \PortReset
      \\
      \PortReset &\Rightarrow \NameCtrl.\PortReset
      \\
      \true &\Rightarrow
      \PortReset \land \NameCtrl.\PortReset
    \end{aligned}
    $}
%
    \\\hline
  \end{tabular}}
\end{table}

% \noteEM{Dans le paragraph precedent, tu resume le cas des connecteurs sans data, dans les prochains, tu formalise avec data, mais pour le lecteur ``casual'', le changement n'est pas evident... }

Now we lift this encoding to the data-sensitive case. Below, we assume that, as in \fig{schema:ArchFailure:BIP}, the
interaction model is defined by a set of connectors, annotated with
Boolean guards and with update assignments.  In particular, we assume
that the guards and update assignments are well-defined for any
interaction allowed by the connector.  For example, the choice
$\NameCtrl.\PortFail\ ?\ \NameCtrl.\mathit{zone} : \topinterval$ in
the update assignment
$
\NameTimer.\mathit{z} := \NameTimer.\mathit{z}
\cap \bigl(
(
\NameCtrl.\PortFail\ ?\ \NameCtrl.\mathit{zone} : \topinterval
)
+ \NameTimer.\mathit{t} \bigr)
$
associated to the connector {\NameTimer.\PortTick \trigsynch
  (\PortFail \trigsynch \NameCtrl.\PortFail)} in
\fig{schema:ArchFailure:BIP} ensures that the assignment is
well-defined independently of whether {\NameCtrl.\PortFail}
participates or not.  Let us denote by $\gamma \subset \ac(P_A)
\times \guards{V_A} \times \assigns{V_A}$ the set of connectors in the
architecture $A$ and by $P_x$ the set of ports involved in the
connector $x \in \gamma$.  Then,
%\addSB{
the interaction model defined by $\gamma$ is 
  $\Gamma = \setdef{(a,g,u)}{(x,g,u) \in \gamma, a \in \intsem{x}}$
and the set of synchronisation vectors encoding $\Gamma$ is
%}
$\nopri[\Gamma] \bydef{=}
\setdef{\nopri[x,g,u]}{(x,g,u) \in \gamma}$, with %% $\nopri[x,g,u]$ is the  defined by
%
\begin{multline}
  \label{eq:nopri:connector}
  \nopri[x,g,u] \bydef{=}
  %% \Tuple{p\bigl(\export[A](p),{{\color{blue}\export[A]'(p),}}b_p\bigr)}^{p \in P_x}
  \Tuple{\Setdef{p\bigl(\export[A](p),\export[A]'(p),b_p\bigr)}{p \in P_x \cap P}}^{P \in P_C^{C \in \cC} \cup \partition}
  \rightarrow
  \alpha \bigl(b_p^{p \in P_x}\bigr)
  \\
  \left[
    \bigwedge R(\tau(x))\bigl[\nicefrac{b_p}{p}\bigr]
    \land
    \bigwedge_{(x := e_x) \in u} (?x' = e_x)
    \land
    \bigwedge_{x \in \export[A](x), x \not\in u} (?x' = x)
    \right]
  \,,
\end{multline}
%
with $b_p^{p \in P_x}$ fresh Boolean variables, $\alpha$ a fresh name, $\tau :
\ac(P_A) \rightarrow \ct(P_A)$ and $R : \ct(P_A) \rightarrow
\cru(P_A)$ the two mappings \cite{BliSif10-causal-fmsd} discussed above and
illustrated in \tab{connectors}, $\bigl[\nicefrac{b_p}{p}\bigr]$ is
the substitution that replaces in the expression that precedes it
all occurrences of all $p$ by corresponding variables $b_p$.

%\addSB{%
%
For the sake of clarity, we simplify the case study encoding in
\fig{FailureTimer:pNet} by reusing the port names of the original
architecture instead of fresh names $\alpha$.  This is made
possible by the fact that each synchronisation vector involves at
most one action of interest (see the properties in
\cite{arch-pNets-RR}).
%% In the general encoding, the global actions of the
%% synchronisation vectors should be fresh action names.
%% However,
%% for the properties considered in our case study, we only need to
%% exhibit the actions of the holes.  Furthermore, each
%% synchronisation vector involves at most one such action.
%% This
%% allows us to reuse them as the global actions of the
%% synchronisation vectors, simplifying the translation of
%% properties to be verified (\cf \app{properties}).
%
%}\noteLH{ce
  %% paragraphe sonne vraiment note de rapport de recherche, tu
  %% arrives a lui donner un peu d'abstraction/de recul?}\noteSB{Est-ce que c'est mieux comme \c{c}a ? OUI -- Ludo}

\newCoord{In the following theorem, we claim an \emph{isomorphism}
  between the open automaton semantics of a pNet encoding a BIP
  architecture and the LTS semantics of this architecture applied to a
  set of simple components.  We omit the formal definition of this
  isomorphism relation.  However, noting that open automata are,
  essentially, symbolic representations of automata with data, we can
  summarise it as follows: a transition belongs to the LTS iff a
  corresponding open transition belongs to the open automaton and the
  source and target data values of the LTS transition satisfy the
  predicate and implement the assignments of the open transition.
%
}

\begin{theorem}
  The open automaton $\semopen{\nopri[A,\partition]}$ corresponding to
  $\nopri[A,\partition]$ is
  isomorphic to the LTS $\semclosed{\Gamma\bigl(\cC_A, (B_D)^{D \in
      \partition}\bigr)}$ (see \defn{comp:semantics}), with, for each
  $D \in \partition$, the component $B_D \bydef{=} \tuple{\{q\}, q,
    V_D, \val{0}{D}, D, \export[D], \goesto{}}$, with
  a fresh state $q$ and
%
  \begin{align*}
    V_D &= \bigcup_{p \in D} \export[A](p)
    \,,
    &\val{0}{D}(v) &= \bot, \text{ for all $v \in V_D$}
    \,,\\
    \goesto{} &= \bsetdef{(q, p, \true, \noop, q)}{p \in D}
    \,,
    &\export[D](p) &= \export[A](p), \text{ for all $p \in D$}
    \,.
  \end{align*}
%
\end{theorem}
%
%% \begin{proof}[Sketch of a proof]
%%   \noteSB{This theorem is very probably correct: the key question is
%%   whether isomorphism is, indeed, the relation to use. In absence of
%%   silent transitions, must be the case.}
%%   The theorem follows from the correctness of $\tau$ and $R$ shown in
%%   \cite{BliSif10-causal-fmsd}.
%% \end{proof}

%% \todoSB{This repeats things that were already said before.  Will reduce.}
%% Observe now that this encoding can be optimised to remove unnecessary
%% Boolean variables and loop transitions marked by \false.  For
%% instance, consider the ports shown in red in the second column of
%% \tab{connectors}.  All these ports only appear in the roots of the
%% corresponding causal interaction trees and, furthermore, each of these
%% trees only has one root (see \app{trees}).  In such cases, as in the
%% {\NameTimer.\PortTick \trigsynch (\PortFail \trigsynch
%%   \NameCtrl.\PortFail)} example discussed above, the Boolean variables
%% and the \false-marked loop transitions associated to these ports can
%% be removed.  Furthermore, the rules associated to these ports in the
%% corresponding systems of causal rules (see the rules in red in the
%% third column of \tab{connectors}) become tautological and can also be
%% removed from the corresponding guards in \eq{nopri:connector}.

%% %****************************************************************
%% \subsection{Encoding with the maximal progress priority model}
%% \label{secn:enc:maxprog}

\paragraph{Encoding of the maximal progress}
We only present the key idea, which consists in
introducing an additional Boolean variable, for each port, for which we
have introduced one in $\nopri[A, \partition]$.  Intuitively, 
the Boolean
variables introduced for the encoding of interaction models
determine whether it is the original transition
labeled by the port that is executed (\true), or rather the corresponding
self-loop, introduced by the encoding (\false).  The new variables
determine whether \emph{there is an original transition labelled by
  $p$ that could be executed from the same state}, \ie[,] with $b_p$
the variable introduced above for the encoding without maximal
progress, $q \goesto{p(\true, b_p)} q'$ iff $\exists q'': q
\goesto{p(b_p)} q''$ with $b_p = \true$ in $\nopri[A, \partition]$.
In the SV guard, we have to check whether \emph{all ports
  leading to $p$ in the causal interaction tree $\tau(x)$ can be fired} (see the second column of
\tab{connectors} for examples).  If so, then $p$ must be fired, \ie
$p(\false)$ must be blocked.
%% This is pretty much straightforward, but
%% care must be taken when doing this in the general case, since causal
%% interaction trees can get pretty tricky due to $\oplus$ (see
%% \app{trees}).}
%****************************************************************
%****************************************************************

%% \subsection{SMT encoding of open pNets}
% \label{secn:smt}

%% \section{Case study}
%% \label{secn:case-study}

%% %****************************************************************

%% \subsection{Verification of the characteristic properties}
%% Eric: je pense que ce soutitre est inutile...

%% \label{secn:arch:verif}

%% \begin{figure}[t]
%%   \centering
%%   \includegraphics[width=\columnwidth]{TimerOpenAut}
%%   \caption{The open Automaton of the Failure Monitor Architecture}
%%   \label{schema:ArchFailure:BIP}
%% \end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{TimerOADetailed}
  \caption{The open Automaton of the Failure Monitor Architecture}
  \label{fig:ArchFailure:OA}
\end{figure}

%\bigskip

\paragraph{Practical experiments.}
The above encoding provides a mechanism for the symbolic verification
of architectures using our existing tool
\cite{HMZ-FORTE2016QBMZ-AVOCS18} to compute the open automaton
semantics of an open pNet. This tool computes open transitions from
pLTS behaviours and synchronisation vectors of the pNet, then uses an
SMT engine to check satisfiability of their predicates, minimising the
size of the resulting automaton.
%% The encoding into the SMT language involves the definition of the
%% action and data domains and operators, and the axiomatisation of
%% domain-dependant properties, to allow the SMT engine to be as precise
%% as possible (even if it cannot be complete).
%% The resulting open automaton will later be subject to (symbolic)
%% model-checking and equivalence checking algorithms, themselves
%% requiring automatic proof assistance. \todoEM{trop flou pour le
%%   moment, mais je n'ose pas dire des trucs trop precis... meme si on a
%%   deja une implem de la FH-bisim, elle n'est pas validee!}

In \fig{ArchFailure:OA} we show the full open automaton
obtained from the pNet in \fig{FailureTimer:pNet}.  Due to space limitations, we do not show the
details of the open transitions, but only the assignments of state
variables, some useful parts of the predicates, and the resulting
action; full details can be found in 
\cite{arch-pNets-RR}.
%
%\noteSB{Removed ``As an example...'', since that \ex{TimerOT3} was
%  simplified and does not correspond to the real OT.}
%% As an example, $OT_3$ is the one
%% explained in \ex{TimerOT3} (page~\pageref{ex:TimerOT3}).
This automaton has 12 transitions, including those encoding various
possible firing of some interactions, \eg $OT_2$ and $OT_3$ for
{\PortFail}.  \newCoord{Notice, however, that the global actions of these open transitions have an additional Boolean parameter.  In this context, model checking should be understood after application of the encoding, namely here the original {\PortFail} event must be an effective ``fail'' of the hole ``B'', that is a $\mathit{fail}(\mathit{true})$ action in the open automaton.}

\newCoord{Model-checking of open-automaton is out of the scope of this paper,
though the resulting automaton here is small enough to observe the
kind of properties we can prove. It is clear that our encoding allows
to test specific values of state-variables in formulas, like e.g.:
  \begin{equation}
    \label{eq:prop:z:top}
    \AW[\quad]{(z = \topinterval)}{\PortFail} \land\quad
    \AG \bigl(
    \PortReset \rightarrow \AW[]{(z = \topinterval)}{\PortFail}
    \bigr).
  \end{equation}
that says that \emph{as long as no failure has occured, the $z$ variable of the
$\NameTimer$ component has the value $\topinterval$}.

But we can also (as long as we get a proper axiomatisation of our data
operators in the SMT engine) handle more involved data properties,
like here the fact that \emph{a reset can only be requested within the
specified delay after a failure:}
  \begin{multline}
    \label{eq:prop:to:delay}
    \forall T_0, T_1 \in \sZ,\
    \AG \Big((\PortFail \land \NameTimer.t = T_0) \rightarrow
    \\
    \AW[]{\bigl((\PortAsk \land \NameTimer.t = T_1) \rightarrow
      (T_1 - T_0 \in \NameCtrl.\mathit{zone})\bigr)
    }{
      (\PortReset \lor \PortResume)
    }\Big),
  \end{multline}
}

Last, a detailed study shows that the safety property stated
on page~\pageref{property:reset} does not hold, because of the
{\PortFail} loop on state $11$.  This is because we did not
use the maximal progress assumption here.  If we do, we get the
corrected behaviour %% in \fig{ArchFailure:OA-MaxProgress}
%% (\app{full-results}), in which
where
$OT_7$ disappears, and $OT_2$ and
$OT_{10}$ are restricted to $b0=\false$. This one verifies all
the properties listed above.%% in \app{properties}.

%% \noteEMin{on peut encore bouger deux lignes de l'annexe D.2, pour
%%   mentionner ici une propriete de liveness, mais ca va dependre de la
%%   place, c'est deja un peu long ici...}

%% %****************************************************************

%% \subsection{Composition of architectures}
%% \label{secn:arch:composition}

%% We now consider a system comprising two components $B_1$ and $B_2$
%% with the corresponding interfaces $\tuple{V_i, P_i,
%%   \export[i]}_{i=1,2}$, such that $\listset{\PortFail_i,
%%   \PortResume_i} \subseteq P_i$, for $i = 1,2$.

%% Assuming that the associated system requirements comprise the
%% capability of monitoring the failure events corresponding to ports
%% $\PortFail_1$ and $\PortFail_2$ of the two components, we would like
%% to apply two instances of the Failure Monitor architecture.  However,
%% we impose an additional constraint: only one timer coordinator is
%% allowed, for example because it must eventually be implemented by a
%% hardware component.

%% We consider two architecture instances $A_i = \tuple{\cC_i, V_{A_i},
%%   P_{A_i}, \export[A_i], \Gamma_i}$, for $i = 1,2$, with all their
%% constituting elements cloned from the Failure Monitor architecture of
%% the previous examples while substituting each entity $X$ by $X^i$.
%% For instance, the initialisation of the $\mathit{zone}$ variable is
%% transformed into $\IntervalInit[1]$ and $\IntervalInit[2]$,
%% respectively.  The only exception to this cloning procedure are the
%% sets of coordinators: $\cC^i = \listset{\NameTimer,
%%   \NameCtrl^i}$, \ie the $\NameTimer$ component is shared, while
%% components $\NameCtrl^1$ and $\NameCtrl^2$ are obtained by making two
%% copies of $\NameCtrl$ as above.

%% In order to apply both architectures, we must verify that they are
%% upwards compatible and compute the combined interaction model $\Gamma$
%% in $A_1 \arcomp A_2 = \tuple{\cC_1 \cup \cC_2, V_{A_1} \cup V_{A_2},
%%   P_{A_1} \cup P_{A_2}, \export[A_1] \cup \export[A_2], \Gamma}$
%% (see \defn{arch:composition}).

%% \todoSBin{Complete this section}

%****************************************************************
%****************************************************************

\section{Related work}
\label{secn:related}
%% Aside from previous work on pNets and BIP architectures mentioned in \secn{introduction}. We  review here the closest related works outside the scope of pNets and BIP architectures, with a particular focus on compositional and parameterised verification.

%%A few tracks, pas r\'edig\'es...

%%% compositional verification
%approaches by typing interfaces/protocols == Simon

%% infinite/parameterised model checking? == ERIC (listes ci-dessous)  + Simon

The design methodology based on BIP architectures
is inspired by the notion of design patterns introduced
in~\cite{gamma94-patterns}.  It is radically different
insofar as BIP architectures posses formal semantics; their
composition is well defined and preserves their characteristic
properties.  This is a relatively novel trend with few comparable
works, whereof the most relevant is a theory formalising common
architectural styles, such as \emph{publisher-subscriber} or
\emph{blackboard}, proposed in \cite{Marmsoler14,Marmsoler18}.

Although direct verification of BIP models is possible
\cite{Lacra15,DFinder,DFinder2,atva15,wang2015verification}, none of
these previous works address
compositional verification of parameterised BIP systems with data
and maximal progress priorities achieved in the present paper.

From a broader perspective,
basic research on behaviour models and verification algorithms for
data-sensitive systems started in the nineties, with the seminal work
of Hennessy, Lin, and their colleagues on value-passing systems with
assignments
\cite{HennessyL1995:353,CONCUR::Lin1996,APSEC::Lin2001}. Later, many
different works addressing various classes of infinite-state systems
and/or parameterised topologies have been published, using
combinations of approaches, often including predicate abstraction and
SMT satisfiability (\eg \cite{DBLP:journals/jsat/AlbertiGPRR12,BruniEtAl-Tiles-Concur2000,CimattiEtAl-NUXMV-CAV2014,ChampionEtAl-Kind2-CAV2016,DBLP:conf/cade/GhilardiNRZ08}). With
respect to these, we use symbolic representations not only to get a
finite representation of infinite spaces, but also to express the
(data-sensitive) synchronisations with the environment, making our
models suitable for compositional verification.
Among these works, several have shown the capacity of the SMT engines
(either Z3 or Yikes) as servers for solving verification conditions of
the algorithms, for large case-studies
(\eg \cite{DBLP:journals/corr/abs-1806-11459,DBLP:journals/corr/CimattiGMT13}). %% Also

%% SMT on an application not using component structure. is there a strong Z3 example == Eric

As compositional proof of safety is difficult, some approaches rely on theorem proving to ensure the safety of component operations. Coqots and Pycots~\cite{BCDLM:CBSE2014} even manage to prove the safety of reconfiguration procedures which are known to be highly difficult to verify, and massively parameterised. The approach relies on a high expertise and significant efforts from the user. Here, we rely on automatic verification thanks to the SMT solver but we cannot prove the safety of the reconfiguration procedure.

In~\cite{DDJO:JLAMP2012}, the authors propose a compositional proof system for distributed objects that is suitable for implementation within the KeY framework~\cite{BHS:Key2007} and uses a Hoare logic approach.
%The approach relies on history invariant for each object, separating local reasoning for proving object invariant from global reasoning that relies on well-formedness of traces. 
Compared to this  approach, we do not  deal with complex history-based specifications and use interaction specification and SMT-reasoning instead of Hoare logic. 
%Also the correctness of the composition is given by the properties of architectures which allows for more flexibility than well-formedness of traces.

%****************************************************************
%****************************************************************

\section{Conclusion}
\label{secn:conclusion}

BIP architectures are composition tools that enforce safety properties; the composition of architectures entails the composition of the associated properties.
We have extended architectures with data-sensitive interactions, and proved that this extension still guarantees the preservation of safety properties by architecture composition, under reasonable assumptions. This extends the original compositional methodology offered by BIP architectures. Then we use pNets as a semantic formalism to encode architectures with data. pNet is a low level coordination model for open systems, in which composition preserves bisimulation equivalences. pNet is equipped with tools computing its behavioural semantics in terms of symbolic automata, allowing model-checking and equivalence checking with algorithms relying on SMT engines.
As a result, we obtain automatic and compositional guarantees of safety properties with data where compositionality is given by the BIP architectures, and pNet tools provide automatic verification of the properties of each architecture.

The translational approach allows us to benefit from the methods
and tools developed separately in BIP and pNets communities,
avoiding the additional effort of designing the corresponding
tools in both contexts from scratch.

\newCoord{The presented work opens a number of avenues for future
  work, among which the most immediate ones consist in 1)~developing
  tools that would implement the discussed encoding and verification
  techniques; 2)~studying the preservation of liveness properties by
  architecture composition under assumptions similar to those
  discussed in \cite{AttieBBJS16-architectures-faoc} and
  3)~generalisation to priority models other than maximal progress.}
%The automatic checking of properties given by the pNet framework
%
%Discussion…:
%-	What did we learn about comparison of the coordination models ? Worth mentioning architectures are flat, so we don’t use the power of hierschical pNets (as providinh=g encapsulation mechanism) ?
%-	What about liveness, fairness ?
%-	Extension to other kind of priorities ?
%-	Shall we make a full proof of correctness of the encoding?
%-	Tooling = implementation of the encoding from BIP arch editing tools + tracing back verification diagnoses to the BIP level.
%
%
%What have we done.
%
%Discussion:
%
%- what did we lern about comparison of the coordination models ?
%
%- extension to other priority features
%
%- proof of correctness ?

%****************************************************************
%****************************************************************

\bibliographystyle{abbrv}
\bibliography{biblio,biblioFromPnets}

%****************************************************************
%****************************************************************
\end{document}