### 2020-07-24 2016-Weak-FH-Bisim/Coverletter.tex
first vesrsion

\documentclass[10pt]{article}
\usepackage[latin1]{inputenc} 
\usepackage{alltt,url}
%% Old ttbox definition without frame:
\newenvironment{ttbox}{\begin{alltt}\small\tt}%
                      {\end{alltt}}

\thispagestyle{empty}
\addtolength{\textheight}{2cm}
\begin{document}
\section*{Cover Letter }
\subsection*{Submission to Logical Methods in Computer Science}
\medskip
\begin{ttbox}
Title: Compositional equivalences based on Open pNets
Authors: R. Ameur-Boulifa, L. Henrio, and E. Madelaine
\end{ttbox}
\bigskip
%\subsubsection*{Letter to Reviewers and Editors}
Dear reviewers and editors,

\medskip
This article defines bisimulation relations  for the comparison of systems specified as parameterised Networks of synchronised automata; it formalises the work we did in the last 4 years theory for the operators composing concurrent processes. This theory is based on the definition of operators as open processes. By defining the operational semantics of open processes, using open transitions that include symbolic hypotheses on the behaviour of the holes.

\smallskip

We think this article fits with the scope and criteria of ``Logical Methods in Computer Science'' for the following reasons. First, to our knowledge this is the first paper we propose a theoretical framework for analysing the behaviour of open systems modelled as pNets, i.e., open terms of suitable process calculi.
The  semantics of pNets is given by means of a symbolic open transition system. Behavioural equivalences for pNets, like strong and weak bisimilarities, are defined over such a open transition systems.

\smallskip

Concerning prior publications Forte'16, we have already defined a theory for the operators composing concurrent processes. The theory is based on the definition of operators as open pNets. However, the study of compositionability was only partial, and in particular the proof that bisimulation is an equivalence that is one new contribution of this article.....

\smallskip

Considering the length of the submitted material, we suggest that we
leave only the two first appendices in the final version of the paper (leading to a
paper of 73 pages) because they are crucial to understand the paper.
The other appendices will be available as a research report, we keep
them in the submission to ease the work of the reviewers. %Of course, depending on the opinion of the reviewers, we will also move more or less material in the body of the paper, in the appendices, or to the research report.

\begin{flushright}
  Best Regards,\\
  Rab\'ea Ameur-Boulifa, Ludovic Henrio, and Eric Madelaine
\end{flushright}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: