### 2021-04-23 2021-Refinement/Traffic_Lights_Spec.tex
Composisiton example fully integrated

\begin{tikzpicture}

\node[state,initial] (v1) at (150:3.5) {R};
\node[state] (v2) at (30:3.5) {G};
\node[state] (v3) at (270:2) {Y};

\draw (v1) edge[dir=220] node[below] {\OTd{\act{tick}}{R \mapsto \act{tick}}{}} (v1);
\draw (v1) edge[dir=150] node[left] {\OTd{\act{onRed}}{}{}} (v1);
\draw (v1) edge[dir=80] node[above] {\OTd{\tau}{C \mapsto \theta\mpar{x}, R \mapsto \act{set}\mpar{x}}{}} (v1);
\draw (v1) edge node {\OTd{\act{TurnGreen}}{R \mapsto \act{over}\mpar{x}, C \mapsto \delta\mpar{x}}{}} (v2);
\draw (v2) edge[dir=100] node[above] {\OTd{\act{tick}}{R \mapsto \act{tick}}{}} (v2);
\draw (v2) edge[dir=30] node[right] {\OTd{\act{onGreen}}{}{}} (v2);
\draw (v2) edge[dir=320] node[xshift=1cm,below] {\OTd{\tau}{C \mapsto \theta\mpar{x}, R \mapsto \act{set}\mpar{x}}{}} (v2);
\draw (v2) edge node[pos=0.6] {\OTd{\act{TurnYellow}}{R \mapsto \act{over}\mpar{x}, C \mapsto \delta\mpar{x}}{}} (v3);
\draw (v3) edge[dir=340] node {\OTd{\act{tick}}{R \mapsto \act{tick}}{}} (v3);
\draw (v3) edge[dir=270] node {\OTd{\act{onYellow}}{}{}} (v3);
\draw (v3) edge[dir=200] node {\OTd{\tau}{C \mapsto \theta\mpar{x}, R \mapsto \act{set}\mpar{x}}{}} (v3);
\draw (v3) edge node[pos=0.4] {\OTd{\act{TurnRed}}{R \mapsto \act{over}\mpar{x}, C \mapsto \delta\mpar{x}}{}} (v1);

\end{tikzpicture}