### 2021-04-24 2021-Refinement/Composition_nolock.tex

\begin{tikzpicture}

\node[state] (v2) at (45:3) {0C};
\node[state,initial] (v1) at (135:3) {0S};
\node[state] (v4) at (225:3) {1S};
\node[state] (v3) at (315:3) {1C};

\draw (v1) edge[bend left,looseness=0] node {\OTd{\act{set}\mpar{x}}{}{c \gets 0, t \gets x}} (v2);
\draw (v2) edge[bend left,looseness=0] node {\OTd{\tau}{}{}} (v3);
\draw (v3) edge[bend left,looseness=0] node[near start] {\OTd{\act{tick}}{}{c \gets c + 1}} (v2);
\draw (v2) edge[bend left,looseness=0] node {\OTd{\act{over}\mpar{c}}{}[c \geq t]{}} (v1);
\draw (v1) edge[bend left,looseness=0] node {\OTd{\tau}{}{}} (v4);
\draw (v4) edge[bend left,looseness=0] node {\OTd{\act{tick}}{}{}} (v1);

\end{tikzpicture}