### 2021-08-31 2016-Weak-FH-Bisim/2021-09-letter2Reviewers.tex
squelette de lettre de reponse

\documentclass{article}
\usepackage{a4wide}
\title{Reply to reviewers\\
LMCS \#\,6784 Manuscript: \\``Compositional equivalences based on Open pNets'' }

\begin{document}

\maketitle
\noindent
Dear Editor, dear reviewers,

Please find enclosed the revision of our submission to LMCS. We improved our article according to the comments made by the
anonymous reviewers.

We would first like to thank the reviewers for their analysis of our
paper, and the useful advices they gave us.

We detail below the major changes we made in this revision. After
this, we will review in the next pages the comments of the reviewers
and explain how we addressed them.  \emph{Comments that could be
  addressed trivially and exactly as suggested by the reviewers are
  omitted}.  \bigskip

In this new version, you will find the major improvements listed below:
\begin{itemize}
\item 
\end{itemize}

\bigskip
\noindent
Best regards,
\newline
\noindent
Rab\'ea Ameur-Boulifa, Ludovic Henrio, Eric Madelaine.

\newpage
As explained above, we quote below and reply to text requiring changes that we did not 
address in a straightforward manner. Major points of the reviews have not been omitted in 
any case.

\section*{Annex 1: response to first reviewers' comments}

\section*{Annex 2: response to second reviewers' comments}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: